### 2015-01-14 Manuscript/ohyou.tex
Converted version of original manuscript to LaTeX, minus figures

% Preamble
\documentclass[a4paper,12pt]{article}

\usepackage[osf]{mathpazo} % palatino
\usepackage{ms}            % load the template
\usepackage[round]{natbib} % author-year citations
\usepackage{graphicx}
\usepackage{parskip} 
\usepackage{enumerate}     
\pagenumbering{arabic}

% Title page information

\title{A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies}
\author{
  Natalie Cooper$^{1,2,*,\dag}$, Gavin H. Thomas$^{3,\dag}$, Chris Venditti$^{4}$\\ Andrew Meade$^{4}$ and Rob P. Freckleton$^{3}$\\
}
\date{}
\affiliation{\noindent{\footnotesize

  $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
  $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
  $^4$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
  $^\dag$These authors contributed equally.
}}

\vfill
%\paragraph{Word-count:} X words

\runninghead{Limits of comparative analyses}
\keywords{Bayesian method - comparative methods - delta - kappa - lambda - OU - phylogeny - stabilizing selection - selection} %meant to be after abstract

% End of preamble

\begin{document}
\modulolinenumbers[1]   % Line numbering on every line

\mstitlepage
\parindent = 1.5em
\addtolength{\parskip}{.3em}

\section{Abstract 100-200}
Phylogenetic comparative methods are increasingly used to give new insight into variation, causes and consequences of trait variation among species. The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. However, the parameters of these models have been hypothesised to be biased and only asymptotically behave in a statistically predictable way as datasets become large. This does not seem to be widely appreciated. We show that a commonly used model in evolutionary biology (the Ornstein-Uhlenbeck model) is biased over a wide range of conditions. Many studies fitting this model use datasets that are small and prone to substantial biases. Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model.

\newpage
\raggedright
\doublespacing
\setlength{\parindent}{1cm}

\section{Introduction}
\label{section:introduction} 
The outcomes of evolutionary processes acting over millions of years can be seen in the distributions and covariances of species traits at the tips of phylogenetic trees. Such processes can be modeled using phylogenetic comparative methods (PCMs). These approaches have been used, for example, to model niche conservatism \citep{Wiens:2010aa}, infer potential rates of species responses to climate change \citep{Quintero:2013aa}, test the role of ecological niche as a driver of morphological evolution \citep{pienaar2013macroevolution}, compare rates of evolution of species traits \citep{Claramunt:2012ab} and test for constraints in adaptive radiations \citep{blackburn2013adaptive}. PCMs are now widely recognized as powerful tools with which to model observed patterns in species traits and potentially to infer the evolutionary processes that drive them \citep[e.g.,][]{freckleton2009seven,Nunn:2011aa,o2012evolutionary,pennell2013integrative}. Despite these advances, the properties of some of the most widely used PCMs are poorly understood leading to the potential for inappropriate use and misinterpretation of results.

The majority of PCMs use an explicit evolutionary model to characterize trait evolution \citep{Freckleton:2011aa}. A model can be fit using statistical methods (typically maximum likelihood) and model parameters interrogated to test ideas from evolutionary theory. Most model-based methods for characterizing trait evolution are based on the Brownian constant variance model \citep[for exceptions see][]{price1997correlated,harvey2000comparative,freckleton2006detecting}. The Brownian model, first applied in a phylogenetic context by \citet{cavalli1967} and to across-species data by \citet{felsenstein1973maximum}, is a simple model of trait evolution in which trait variance accrues as a linear function of time, and makes the prediction that traits of closely-related species are more similar than those of distantly-related ones. The Brownian model has been modified in various ways to account for a suite of ecological and evolutionary processes (e.g., Grafen 1989; Hansen 1997; Pagel 1997, 1999). Most of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. These modified Brownian models have become very popular because they fit better, have attractive biological interpretations, and are implemented in freely available packages \citep[e.g., R;][]{R-Core-Team:2014aa}.

The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes. $\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood. It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$. Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously. \citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989). They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes). This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.

One of the most commonly used Brownian-like models is the Ornstein Uhlenbeck (OU) model. This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. The popularity of the OU model has grown exponentially in recent years (Fig. 1), in part because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. OU models have become particularly prevalent in ecological studies: 652 papers containing the phrases “Ornstein Uhlenbeck” and “ecology” were published between 2011 and 2013 (Google Scholar search 30th January 2014). Additionally, although they are pattern based, fit to an OU model is now being used as evidence for the action of ecological, or ecologically driven, processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.

Uses of the Ornstein-Uhlenbeck can be divided into three broad categories that are near equally represented in the literature: (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses. Here, we focus on the first of these because the principles apply to a range of other macroevolutionary models.

Here we investigate the OU model because it does not appear to have been subject to a rigorous critique. The $\alpha$ parameter of the OU model is very similar to Grafen’s $\rho$ (along with other parameters in the literature) in that it is in effect a non-linear transformation of the phylogeny. Consequently, one might imagine that it shares some of the same statistical properties. Results from \citet{ho2013asymptotic}, indicate that there is the potential for revealing strong biases in $\alpha$ , but the implications for practice (e.g. likelihood of rejection of alternative models) is unexplored other than to note that tree shape may play an influence. We present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in comparative analyses. We also show that very small amounts of measurement error in data can have profound effects on the performance of models. Many of our findings are applicable to other models of evolution (see Discussion and Supplemental Figs. S6-S11), but we focus here on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the α parameter.

\section{Materials and Methods}
\subsection{Model Outline} 
\label{section:models} 
According to the Brownian model, a trait X evolves at random at a rate $\sigma$:

  \begin{equation}
    dX(t) = \sigma dW(t)
    \label{equation:BMrate} 
  \end{equation}

where $W(t)$ is a white noise function and is a random variate drawn from a normal distribution with mean $0$ and variance $\sigma^2$. This model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant. The model has two parameters, $\sigma$ and the state of the root at time zero, $X(0)$. The Brownian model predicts after a time $T$ the variance in trait value $X_i$ for species $i$ is:

  \begin{equation}
    var(X_i) = \sigma^2 T
    \label{equation:BMvar} 
  \end{equation}

and the covariance in traits for species $i$ and $j$ is:

  \begin{equation}
    cov(X_i,X_j) = \sigma^2 t_{ij}
    \label{equation:BMcov} 
  \end{equation}

where $t_{ij}$ is the shared evolutionary pathway for species $i$ and $j$, i.e. the time at which they last shared a common ancestor.  Equations \ref{equation:BMvar} and \ref{equation:BMcov} encapsulate the simplicity of the Brownian model, namely it predicts that variances accrue as a linear function of time.

The Ornstein-Uhlenbeck (OU) model describes a mean-reverting process and has the following form, adding an extra term to the Brownian model:

  \begin{equation}
    dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
    \label{equation:OUrate} 
  \end{equation}

The parameter $\mu$ is a long-term mean, and it is assumed that species evolve around this value. $\sigma$ is the strength of evolutionary force that returns traits back towards the mean if they evolve away. The OU model was introduced to population genetics by \citet{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape. The process operating in comparative data is analogous, although clearly is not stabilizing selection despite being sometimes referred to as such (see Discussion). This model has two parameters in addition to those of the Brownian model, $\alpha$ and $\mu$.
The OU model predicts that after a time $T$ for a species $i$, the variance in trait value $X_i$ is:

  \begin{equation}
    var(X_i) = 1 - e^{-2 \alpha T}
    \label{equation:OUvar} 
  \end{equation}

And for a pair of species i and j, the covariance in traits is:

  \begin{equation}
    cov(X_i, X_j) = e^{-2 \alpha (T - t_{ij})} (1 - e^{-2\alpha t_{ij}})
    \label{equation:OUcov} 
  \end{equation}

The variances and covariances predicted by equations \ref{equation:OUvar} and \ref{equation:OUcov} are more complex than those predicted by the Brownian model. In the light of the results below, some properties of this model are worth highlighting:

\begin{enumerate}[(i)]

  \item If $\alpha$ is small then evolution is approximately Brownian: If $\alpha$ is small then $1 - e^{-2\alpha T} \approx 2\alpha T$, i.e. traits accrue variance as if evolving according to a Brownian process. The question is obviously what is a small value of $\alpha$? Figure 2 provides some guidance on this issue, and see point (iv) below.\\

  \item If species $i$ and $j$ diverged recently, evolution is approximately Brownian: if two species diverged recently, then $T - t_{ij} \approx 0$ and hence $cov(X_i, X_j) \approx 1 - e^{-2\alpha t_{ij}} \approx 2\alpha t_{ij}$. Thus, recently diverged species provide little information relevant to estimating non-Brownian evolution according to an OU process. \\

  \item In the long-term the imprint of history is weakened: if $T$ is large (i.e. evolution proceeds for a long time), equation ??? predicts that the variance in $X_i$ tends to a constant, i.e. because the expected value of $X_i$ is $\mu$. Similarly, in equation (4), the covariance between traits tends to a constant because $T$ becomes large relative to $t_{ij}$. Consequently for large groups the model implies that the imprint of history is weak. As a result adding more species to a phylogeny will contribute only relatively weakly to fitting the model, as the deeper branches within a phylogeny will contribute little information to estimating the parameters.\\

\end{enumerate}

Figure 2 gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above. We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height. Figure 2 highlights that the behavior of $\alpha$ can be classified into three regions. When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time. Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent. Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig. 2B).  Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig. 2C). This will be indistinguishable from phylogenetically unstructured data.

In terms of statistical model fitting, a feature to highlight is that in Figure 2A and Figure 2C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig. 2A; 10-50 in Fig. 2C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.

The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.

\subsection{Simulations}
\label{section:sims.methods} 
We used simulations to explore the statistical properties of $\alpha$. We simulated phylogenies with 25, 50, 100, 150, 200, 500, or 1000 tips under pure birth, constant-rate Birth-Death (extinction fractions of 0.25, 0.5 and 0.75), or temporally varying speciation rate (speciation rate modeled as time from the root raised to the power 0.2, 0.5, 2 and 5) models. We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total. Trees were simulated using the R package TESS \cite{hohna2013fast}. We then simulated the evolution of a single trait under a Brownian motion model on each phylogeny using the R package MOTMOT \citep{Thomas:2011aa}.  We estimated α and compared the fit of a Brownian model to that of an OU model using: (i) a likelihood ratio test with 1 degree of freedom with the transformPhylo.ML function in MOTMOT; and, (ii) with Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). We treated Bayes factors $>$ 2 as evidence favouring the OU model. We used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20. For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 for $\sigma^2$. We also normalized the tree length to prevent the $\alpha$ parameter of the OU model from interacting with $\sigma^2$.

OU models have not previously been widely applied within a Bayesian framework, although some packages offer this functionality \citep[e.g., diversitree;][]{FitzJohn:2012aa} and the choice of prior has not been fully explored. We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior. We provide results derived from a strong exponential prior with mean = 1 and a bounded uniform prior (range 0-20) as supplementary material (Figs. S2-S5). We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. Multiple independent chains were run for each analysis to ensure convergence was reached.

We used the same procedure to simulate trait data under a Brownian motion model with known error. Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees. We then simulated data under a Brownian model on each tree. We compared the fit of a Brownian model to that of an OU model (with ML and MCMC stepping stone sampling as above) using the original trees without the addition of extra branch length to the tip edges. The expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.

Although our focus is on the OU model we also explored three other commonly used models \citep[$\kappa$, $\lambda$, and $\delta$;][]{Pagel:1997aa,Pagel:1999aa} fitted using maximum likelihood using the same simulated data. These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the supplementary material.

\subsection{Literature Review}
\label{section:literature.methods} 
To get an overview of the use of OU models in ecology and evolution, we used Google Scholar (accessed 21st November 2013) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2013 that contained the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” (the “biology” term was added to omit physics papers which also use the term “evolution”). We also recorded the total number of papers containing the terms “ecology”, or “evolution” and “biology” published between 2005 and 2013 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig. 1).

Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following ecology and evolution journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science. For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other). Where papers included multiple analyses using different numbers of species we used the median number of taxa. Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. We summarise these results in Figure S1 and the full dataset is available in Supplemental Table S1 along with the full list of references.

\section{Results}
\subsection{Simulations}
\label{section:sims.results} 
Figure 3 shows some examples of profile likelihoods for α generated on different datasets that were simulated according to a Brownian model. It is clear that the shape of the tree affects the shape of the likelihood profile. Trees with increasing speciation rates from root to tip (i.e., that have a lot of recently evolved species) generate very wide, flat likelihood profiles (Fig. 3A). On the other hand, trees with declining speciation rates from root to tip (Fig. 3B), or generated according to a Yule process (Fig. 3C), yield more strongly peaked profiles.

In all three cases, there are several features worth noting. First, the profiles are asymmetric. There is a hard bound at $\alpha = 0$, and as $\alpha$ becomes large, the rate of decrease in the likelihood slows. This is because, as shown in Figure 2, as values of $\alpha$ become large, the effects of changing $\alpha$ on model predictions are increasingly small. It is also notable, that the Maximum Likelihood estimates of $\alpha$ are, in most cases, between 0 and 1. In Figure 2 it was noted that in this region of parameter space, the OU model is very difficult to distinguish from Brownian. This, combined with the asymmetric nature of the profile likelihood is suggestive that estimates of $\alpha$ are likely to be biased, and that the asymptotic assumptions of likelihood ratio tests on $\alpha$ are likely to be violated.

The effects of the asymmetric profile likelihoods become clear when calculating Type I error for the OU model based on ML estimates across a range of tree shapes and sizes (Fig. 4; see Supplemental Figs. S2 and S3 for results with alternative priors for the Bayesian analyses). Regardless of tree shape, Type I error rates are unacceptably high when tree size is small (Figs. 4A, 4C and 4E). For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips. Bayesian estimation is not subject to bias in the log-likelihood statistic and consistently rejects the OU model. However, estimates of α are similar regardless of method.

Figure 5 shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs. S4 and S5). The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error). There are two points worth noting. First, the frequency with which the OU model is favoured increases with tree size for both ML and Bayesian estimation (Fig. 5A). With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species. This is important for the interpretation of the OU model. We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1 (Fig. 5B). Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.

The biases reported above are not unique to the OU model. The parameter $\delta$, which tests for accelerating or decelerating rates of evolution \citep{Pagel:1997aa,Pagel:1999aa} also suffers from elevated Type I error and is upwardly biased, favoring models that imply temporally increasing evolutionary rates (Supplemental Fig. S8). In contrast, although the speciational model ($\kappa$) has slightly elevated Type I errors, the parameter estimates are unbiased for the majority of tree sizes and shapes (Supplemental Fig. S6). As previously shown (Freckleton et al. 2002), the $\lambda$ model is conservative with low Type I error rates (Supplemental Fig. S7). As with the OU model, the $\delta$ models are all strongly favored over Brownian motion when there is error in the species data (Supplemental Figs. S9-S11).

\subsection{Literature Review}
\label{section:literature.results}
In total, 3720 papers containing the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” were published between 2005 and 2013, and the number has increased substantially since 2005 (Fig. 1). Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig. S1 and Table S1). The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig. S1).

\section{Discussion}
\label{section:discussion}
Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways. Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. The results we have presented here illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used.

In the case of the OU model, we can make a series of specific recommendations for future analyses:
\begin{enumerate}[(i)]
  \item At least 200 species should be included. The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
  \item Likelihood ratio tests are untrustworthy. We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. The simulations indicate that likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered. 
  \item Simulate fitted models and compare these to your empirical results. A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests. Any given tree will therefore generate unique parameter estimates. Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}. Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias. 
  \item Consider plausible alternative hypotheses. The results indicate that a simple parsimonious model can be mistaken for an OU process: namely a BM process in which a small amount of error is added to the data. Small ($<$ 1) values of $\alpha$ are quite compatible with this interpretation of the data. The effects of error become more severe with increasing tree size. 
\end{enumerate}

In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}. It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}. However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling. This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton unpub). If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance. However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.

Our review of the literature revealed that the OU model is frequently described and interpreted as a model of ‘stabilizing selection’. The OU model is attractive because it sets effective bounds in probability on the size of species’ traits, but to use the term ‘stabilizing selection’ is inaccurate and misleading. As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche. Under this formulation, α can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.

In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process. This is because according to the OU process traits move back towards the optimum as their values become particularly large or small. Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig. 2). Indeed, Figure 2 implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present. This interpretation is particular important for inference of phylogenetic niche conservatism. Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}. A more parsimonious explanation would be that there is measurement error in the data. This seems particularly likely for climatic niche traits that are often averaged across an entire species range. Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.

If ancillary data are available, however, this may not be the case. If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.  A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances in equation (6) depend on both the shared distances between species and the distance of a node to the nearest tip. This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. These implementations should not be used with fossil data.

The results we report are not unique to the OU model. As noted above, \citet{grafen1989phylogenetic} pointed out that his $\rho$ parameter would likely be biased, and other transformations \citep[e.g., $\lambda$, $\delta$, ACDC;][]{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa} will exhibit similar behavior \citep[e.g., see][]{freckleton2002phylogenetic}. \citet{freckleton2002phylogenetic} explored the behavior of Pagel’s $\lambda$ and found that this statistic was biased but in manner that was likely to be conservative in the rejection of the Brownian model (also see Supplemental Fig. S7). Some transformations may be unbiased: for example, Pagel’s $\kappa$ is a transformation of branch lengths, and is essentially the “standard diagnostic” that is used in testing the assumptions of phylogenetic contrasts, and known to behave appropriately \citep[Supplemental Fig. S6;][]{garland1992procedures}.

In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species. The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes. Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation. The results presented above highlight that in the case of the OU model, this is especially important.

\section{Acknowledgments}
Thanks to Mark Pagel, Matt Pennell, Graham Slater and Rich FitzJohn for fruitful discussions about OU models, and three anonymous reviewers for helpful comments on a previous version of this paper. NC was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696. GHT was supported by a Royal Society University Research Fellowship. AM was supported by BBSRC grant BB/K004344/1 and the computing time was funded by European Research Council Grant no. 268744, Mother Tongue.

\bibliographystyle{biolinnsoc}
\bibliography{ohyou}

\end{document}

### 2015-01-15 Manuscript/ohyou.tex
Addded figures to manuscript minus supplemental

@@
 The popularity of the OU model has grown exponentially in recent years (Fig.
-1), in part because these models are now easy to implement apply via packages in R \citep[e.g.
+\ref{figure.literature}), in part because these models are now easy to implement apply via packages in R \citep[e.g.
 ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}.
@@
 Additionally, although they are pattern based, fit to an OU model is now being used as evidence for the action of ecological, or ecologically driven, processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+
+\begin{figure}
+\centering
+    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure1.pdf}
+\caption{The number of evolutionary biology (closed circles) and ecology (open circles) papers published between 2005 and 2012 containing the phrase “Ornstein-Uhlenbeck”, as a proportion of the total number of evolutionary biology or ecology papers published that year.}
+\label{figure.literature}
+\end{figure}
 
@@
 The question is obviously what is a small value of $\alpha$?
-Figure 2 provides some guidance on this issue, and see point (iv) below.\\
+Figure \ref{figure.traitsim} provides some guidance on this issue, and see point (iv) below.\\
 
@@
 
-Figure 2 gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
+Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
 We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
-Figure 2 highlights that the behavior of $\alpha$ can be classified into three regions.
+Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
 When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
@@
 Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
-2B).
+\ref{figure.traitsim}B).
 Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
-2C).
+\ref{figure.traitsim}C).
 This will be indistinguishable from phylogenetically unstructured data.
 
-In terms of statistical model fitting, a feature to highlight is that in Figure 2A and Figure 2C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
-2A; 10-50 in Fig.
-2C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
+\begin{figure}
+\centering
+    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
+\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
+The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
+equal to correlation between species’ traits).
+This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
+}
+\label{figure.traitsim}
+\end{figure}
+
+In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
+\ref{figure.traitsim}A; 10-50 in Fig.
+\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
 
@@
 We also recorded the total number of papers containing the terms “ecology”, or “evolution” and “biology” published between 2005 and 2013 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
-1).
+\ref{figure.literature}).
 
@@
 \label{section:sims.results} 
-Figure 3 shows some examples of profile likelihoods for α generated on different datasets that were simulated according to a Brownian model.
+Figure \ref{figure.profiles} shows some examples of profile likelihoods for α generated on different datasets that were simulated according to a Brownian model.
 It is clear that the shape of the tree affects the shape of the likelihood profile.
 Trees with increasing speciation rates from root to tip (i.e., that have a lot of recently evolved species) generate very wide, flat likelihood profiles (Fig.
-3A).
+\ref{figure.profiles}A).
 On the other hand, trees with declining speciation rates from root to tip (Fig.
-3B), or generated according to a Yule process (Fig.
-3C), yield more strongly peaked profiles.
+\ref{figure.profiles}B), or generated according to a Yule process (Fig.
+\ref{figure.profiles}C), yield more strongly peaked profiles.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure3.pdf}
+  \caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ‘true’ value of alpha is 0.}
+\label{figure.profiles}
+\end{figure}
 
@@
 There is a hard bound at $\alpha = 0$, and as $\alpha$ becomes large, the rate of decrease in the likelihood slows.
-This is because, as shown in Figure 2, as values of $\alpha$ become large, the effects of changing $\alpha$ on model predictions are increasingly small.
+This is because, as shown in Figure \ref{figure.traitsim}, as values of $\alpha$ become large, the effects of changing $\alpha$ on model predictions are increasingly small.
 It is also notable, that the Maximum Likelihood estimates of $\alpha$ are, in most cases, between 0 and 1.
-In Figure 2 it was noted that in this region of parameter space, the OU model is very difficult to distinguish from Brownian.
+In Figure \ref{figure.traitsim} it was noted that in this region of parameter space, the OU model is very difficult to distinguish from Brownian.
 This, combined with the asymmetric nature of the profile likelihood is suggestive that estimates of $\alpha$ are likely to be biased, and that the asymptotic assumptions of likelihood ratio tests on $\alpha$ are likely to be violated.
@@
 The effects of the asymmetric profile likelihoods become clear when calculating Type I error for the OU model based on ML estimates across a range of tree shapes and sizes (Fig.
-4; see Supplemental Figs.
+\ref{figure.rejection}; see Supplemental Figs.
 S2 and S3 for results with alternative priors for the Bayesian analyses).
 Regardless of tree shape, Type I error rates are unacceptably high when tree size is small (Figs.
-4A, 4C and 4E).
+\ref{figure.rejection}A, \ref{figure.rejection}C and \ref{figure.rejection}E).
 For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips.
@@
 
-Figure 5 shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs.
+\begin{figure}
+  \centering
+  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure4.pdf}
+  \caption{Rejection rates and estimation of $\alpha$ for different tree sizes and shapes.
+Rejection rates (a,c,e) for the OU model at multiple tree sizes for models fitted using ML (circles) and Bayesian (diamonds) methods.
+Estimates of $\alpha$ (b,d,f) based on ML (circles show means and dotted lines show 95\% sampling interval from 1000 trees) and Bayesian (diamonds show means and dotted lines show 95\% sampling interval of the modal value from the posterior distribution from 1000 trees).}
+\label{figure.rejection}
+\end{figure}
+
+Figure \ref{figure.error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs.
 S4 and S5).
@@
 First, the frequency with which the OU model is favoured increases with tree size for both ML and Bayesian estimation (Fig.
-5A).
+\ref{figure.error}A).
 With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species.
@@
 Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1 (Fig.
-5B).
+\ref{figure.error}B).
 Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure5.pdf}
+  \caption{Rejection of BM with measurement error (a) and estimation of α (b). Symbols follow Figure \ref{figure.rejection}.}
+\label{figure.error}
+\end{figure}
 
@@
 In total, 3720 papers containing the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” were published between 2005 and 2013, and the number has increased substantially since 2005 (Fig.
-1).
+\ref{figure:literature}).
 Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig.
@@
 Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig.
-2).
-Indeed, Figure 2 implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.
+\ref{figure.traitsim}).
+Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.
 This interpretation is particular important for inference of phylogenetic niche conservatism.

### 2015-03-13 Analyses/OhYouFigures.R
Figure for literature review

@@ -1,0 +1,18 @@
+# Making figures 
+
+# Read in data
+ds <- read.delim("Data/papers.txt")
+
+# Calculate %
+ds$eco <- (ds$ecology / ds$ecology.total) * 100
+ds$evo <- (ds$evolution / ds$evolution.total) * 100
+
+# Plot figure 1
+
+png("Figures/PapersThruTime.png")
+par(bty = "l")
+plot(evo ~ Year, data = ds, las = 1, pch = 16, cex = 1.2, cex.lab = 1.2, cex.axis = 1.2,
+     ylab = "OU publications (% of total papers)", xlab = "Year published", ylim = c(0, 0.7))
+points(eco ~ Year, data = ds, cex = 1.2)
+legend("topleft", legend = c("evolution", "ecology"), pch = c(16, 1), cex = 1.2, bty = "n")
+dev.off()

### 2015-03-13 Manuscript/ohyou.tex
Started to re-write manuscript

@@
 \section{Abstract 100-200}
-Phylogenetic comparative methods are increasingly used to give new insight into variation, causes and consequences of trait variation among species.
+%Phylogenetic comparative methods are increasingly used to give new insight into variation, causes and consequences of trait variation among species.
 The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model.
@@
 \label{section:introduction} 
-The outcomes of evolutionary processes acting over millions of years can be seen in the distributions and covariances of species traits at the tips of phylogenetic trees.
+%The outcomes of evolutionary processes acting over millions of years can be seen in the distributions and covariances of species traits at the tips of phylogenetic trees.
 Such processes can be modeled using phylogenetic comparative methods (PCMs).
@@
 
-The majority of PCMs use an explicit evolutionary model to characterize trait evolution \citep{Freckleton:2011aa}.
+%The majority of PCMs use an explicit evolutionary model to characterize trait evolution \citep{Freckleton:2011aa}.
 A model can be fit using statistical methods (typically maximum likelihood) and model parameters interrogated to test ideas from evolutionary theory.
@@
 
-The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
+%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
 $\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
@@
 \section{Materials and Methods}
-\subsection{Model Outline} 
-\label{section:models} 
-According to the Brownian model, a trait X evolves at random at a rate $\sigma$:
+According to the Brownian model \cite{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
 
@@
 
-where $W(t)$ is a white noise function and is a random variate drawn from a normal distribution with mean $0$ and variance $\sigma^2$.
-This model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant.
-The model has two parameters, $\sigma$ and the state of the root at time zero, $X(0)$.
-The Brownian model predicts after a time $T$ the variance in trait value $X_i$ for species $i$ is:
+where $W(t)$ is drawn at random from a normal distribution with mean $0$ and variance $\sigma^2$.
+The model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant.
+Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.
+The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species.
+This means that close relatives will be more similar in their trait values than more distant relatives.
+It also means that variance in the trait will increase (linearly) in proportion to time.
+The model has two parameters, the Brownian rate parameter, $\sigma^2$ and the state of the root at time zero, $X(0)$.
 
-  \begin{equation}
-    var(X_i) = \sigma^2 T
-    \label{equation:BMvar} 
-  \end{equation}
-
-and the covariance in traits for species $i$ and $j$ is:
-
-  \begin{equation}
-    cov(X_i,X_j) = \sigma^2 t_{ij}
-    \label{equation:BMcov} 
-  \end{equation}
-
-where $t_{ij}$ is the shared evolutionary pathway for species $i$ and $j$, i.e.
-the time at which they last shared a common ancestor.
-Equations \ref{equation:BMvar} and \ref{equation:BMcov} encapsulate the simplicity of the Brownian model, namely it predicts that variances accrue as a linear function of time.
-
-The Ornstein-Uhlenbeck (OU) model describes a mean-reverting process and has the following form, adding an extra term to the Brownian model:
+The Ornstein-Uhlenbeck (OU) model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
+The model has the following form:
 
@@
 
-The parameter $\mu$ is a long-term mean, and it is assumed that species evolve around this value.
-$\sigma$ is the strength of evolutionary force that returns traits back towards the mean if they evolve away.
-The OU model was introduced to population genetics by \citet{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape.
-The process operating in comparative data is analogous, although clearly is not stabilizing selection despite being sometimes referred to as such (see Discussion).
-This model has two parameters in addition to those of the Brownian model, $\alpha$ and $\mu$.
-The OU model predicts that after a time $T$ for a species $i$, the variance in trait value $X_i$ is:
+Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$.
+The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value.
+$\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it.
+$\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$.
 
-  \begin{equation}
-    var(X_i) = 1 - e^{-2 \alpha T}
-    \label{equation:OUvar} 
-  \end{equation}
+The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape.
+The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see Discussion).
 
-And for a pair of species i and j, the covariance in traits is:
-
-  \begin{equation}
-    cov(X_i, X_j) = e^{-2 \alpha (T - t_{ij})} (1 - e^{-2\alpha t_{ij}})
-    \label{equation:OUcov} 
-  \end{equation}
-
-The variances and covariances predicted by equations \ref{equation:OUvar} and \ref{equation:OUcov} are more complex than those predicted by the Brownian model.
+The variances and covariances predicted by the OU model are more complex than those predicted by the Brownian model.
 In the light of the results below, some properties of this model are worth highlighting:
@@
 
-  \item If $\alpha$ is small then evolution is approximately Brownian: If $\alpha$ is small then $1 - e^{-2\alpha T} \approx 2\alpha T$, i.e.
-traits accrue variance as if evolving according to a Brownian process.
-The question is obviously what is a small value of $\alpha$?
-Figure \ref{figure.traitsim} provides some guidance on this issue, and see point (iv) below.\\
+  \item If $\alpha$ is small then evolution is approximately Brownian, The question is what is a small value of $\alpha$? Figure \ref{figure.traitsim} provides some guidance on this issue.\\
 
@@
 
-  \item In the long-term the imprint of history is weakened: if $T$ is large (i.e.
-evolution proceeds for a long time), equation ???
-predicts that the variance in $X_i$ tends to a constant, i.e.
+  \item In the long-term the imprint of history is weakened.
+If evolution proceeds for a long time the variance in $X_i$ tends to a constant, i.e.
 because the expected value of $X_i$ is $\mu$.
@@
 \ref{figure.traitsim}B).
+
 Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
@@
 
-In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
+In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0-0.5 in Fig.
 \ref{figure.traitsim}A; 10-50 in Fig.
-\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
+\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1-5) that the biggest discernible effects occur.
 
@@
 The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the supplementary material.
-
-\subsection{Literature Review}
-\label{section:literature.methods} 
-To get an overview of the use of OU models in ecology and evolution, we used Google Scholar (accessed 21st November 2013) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2013 that contained the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” (the “biology” term was added to omit physics papers which also use the term “evolution”).
-We also recorded the total number of papers containing the terms “ecology”, or “evolution” and “biology” published between 2005 and 2013 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
-\ref{figure.literature}).
-
-Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following ecology and evolution journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
-For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other).
-Where papers included multiple analyses using different numbers of species we used the median number of taxa.
-Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category.
-We summarise these results in Figure S1 and the full dataset is available in Supplemental Table S1 along with the full list of references.
 
@@
 S9-S11).
-
-\subsection{Literature Review}
-\label{section:literature.results}
-In total, 3720 papers containing the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” were published between 2005 and 2013, and the number has increased substantially since 2005 (Fig.
-\ref{figure:literature}).
-Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig.
-S1 and Table S1).
-The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
-S1).
 
@@
 
+\section{Appendix}
+\subsection{Model Outline} % correct the equations!
+\label{section:models} 
+According to the Brownian model, a trait X evolves at random at a rate $\sigma$:
+
+  \begin{equation}
+    dX(t) = \sigma dW(t)
+    %\label{equation:BMrate} 
+  \end{equation}
+
+where $W(t)$ is a white noise function and is a random variate drawn from a normal distribution with mean $0$ and variance $\sigma^2$.
+This model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant.
+The model has two parameters, $\sigma$ and the state of the root at time zero, $X(0)$.
+The Brownian model predicts after a time $T$ the variance in trait value $X_i$ for species $i$ is:
+
+  \begin{equation}
+    var(X_i) = \sigma^2 T
+    \label{equation:BMvar} 
+  \end{equation}
+
+and the covariance in traits for species $i$ and $j$ is:
+
+  \begin{equation}
+    cov(X_i,X_j) = \sigma^2 t_{ij}
+    \label{equation:BMcov} 
+  \end{equation}
+
+where $t_{ij}$ is the shared evolutionary pathway for species $i$ and $j$, i.e.
+the time at which they last shared a common ancestor.
+Equations \ref{equation:BMvar} and \ref{equation:BMcov} encapsulate the simplicity of the Brownian model, namely it predicts that variances accrue as a linear function of time.
+
+The Ornstein-Uhlenbeck (OU) model describes a mean-reverting process and has the following form, adding an extra term to the Brownian model:
+
+  \begin{equation}
+    dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
+    %\label{equation:OUrate} 
+  \end{equation}
+
+The parameter $\mu$ is a long-term mean, and it is assumed that species evolve around this value.
+$\sigma$ is the strength of evolutionary force that returns traits back towards the mean if they evolve away.
+The OU model was introduced to population genetics by \citet{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape.
+The process operating in comparative data is analogous, although clearly is not stabilizing selection despite being sometimes referred to as such (see Discussion).
+This model has two parameters in addition to those of the Brownian model, $\alpha$ and $\mu$.
+The OU model predicts that after a time $T$ for a species $i$, the variance in trait value $X_i$ is:
+
+  \begin{equation}
+    var(X_i) = 1 - e^{-2 \alpha T}
+    \label{equation:OUvar} 
+  \end{equation}
+
+And for a pair of species i and j, the covariance in traits is:
+
+  \begin{equation}
+    cov(X_i, X_j) = e^{-2 \alpha (T - t_{ij})} (1 - e^{-2\alpha t_{ij}})
+    \label{equation:OUcov} 
+  \end{equation}
+
+The variances and covariances predicted by equations \ref{equation:OUvar} and \ref{equation:OUcov} are more complex than those predicted by the Brownian model.
+In the light of the results below, some properties of this model are worth highlighting:
+
+\begin{enumerate}[(i)]
+
+  \item If $\alpha$ is small then evolution is approximately Brownian: If $\alpha$ is small then $1 - e^{-2\alpha T} \approx 2\alpha T$, i.e.
+traits accrue variance as if evolving according to a Brownian process.
+The question is obviously what is a small value of $\alpha$?
+Figure \ref{figure.traitsim} provides some guidance on this issue, and see point (iv) below.\\
+
+  \item If species $i$ and $j$ diverged recently, evolution is approximately Brownian: if two species diverged recently, then $T - t_{ij} \approx 0$ and hence $cov(X_i, X_j) \approx 1 - e^{-2\alpha t_{ij}} \approx 2\alpha t_{ij}$.
+Thus, recently diverged species provide little information relevant to estimating non-Brownian evolution according to an OU process.
+\\
+
+  \item In the long-term the imprint of history is weakened: if $T$ is large (i.e.
+evolution proceeds for a long time), equation ???
+predicts that the variance in $X_i$ tends to a constant, i.e.
+because the expected value of $X_i$ is $\mu$.
+Similarly, in equation (4), the covariance between traits tends to a constant because $T$ becomes large relative to $t_{ij}$.
+Consequently for large groups the model implies that the imprint of history is weak.
+As a result adding more species to a phylogeny will contribute only relatively weakly to fitting the model, as the deeper branches within a phylogeny will contribute little information to estimating the parameters.\\
+
+\end{enumerate}
+
+Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
+We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
+Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
+When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
+Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent.
+Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
+\ref{figure.traitsim}B).
+Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
+\ref{figure.traitsim}C).
+This will be indistinguishable from phylogenetically unstructured data.
+
+\begin{figure}
+\centering
+    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
+\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
+The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
+equal to correlation between species’ traits).
+This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
+}
+\label{figure.traitsim}
+\end{figure}
+
+In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
+\ref{figure.traitsim}A; 10-50 in Fig.
+\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
+
+The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
+
+\section{Supplementary Material}
+
+  \subsection{Literature Review}
+  \label{section:literature.methods} 
+  To get an overview of the use of OU models in ecology and evolution, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution'').
+We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'' published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
+\ref{figure.literature}).
+
+  Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following ecology and evolution journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
+For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other).
+Where papers included multiple analyses using different numbers of species we used the median number of taxa.
+Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category.
+We summarise these results in Figure S1 and the full dataset is available in Supplemental Table S1 along with the full list of references.
+
+  \label{section:literature.results}
+  In total, 3720 papers containing the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” were published between 2005 and 2013, and the number has increased substantially since 2005 (Fig.
+\ref{figure:literature}).
+Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig.
+S1 and Table S1).
+The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
+S1).
+
 \end{document}

### 2015-04-07 Manuscript/ohyou.tex
Beginning new edits to Biol Linn Soc version

@@
 \label{section:introduction}
-%The outcomes of evolutionary processes acting over millions of years can be seen in the distributions and covariances of species traits at the tips of phylogenetic trees.
-Such processes can be modeled using phylogenetic comparative methods (PCMs).
+
+  Phylogenetic comparative methods are powerful tools for identifying patterns in the evolution of species traits, and for potentially inferring the evolutionary processes that underlie them \citep[e.g.,][]{freckleton2009seven,Nunn:2011aa,o2012evolutionary,pennell2013integrative}.
   These approaches have been used, for example, to model niche conservatism \citep{Wiens:2010aa}, infer potential rates of species responses to climate change \citep{Quintero:2013aa}, test the role of ecological niche as a driver of morphological evolution \citep{pienaar2013macroevolution}, compare rates of evolution of species traits \citep{Claramunt:2012ab} and test for constraints in adaptive radiations \citep{blackburn2013adaptive}.
-PCMs are now widely recognized as powerful tools with which to model observed patterns in species traits and potentially to infer the evolutionary processes that drive them \citep[e.g.,][]{freckleton2009seven,Nunn:2011aa,o2012evolutionary,pennell2013integrative}.
-Despite these advances, the properties of some of the most widely used PCMs are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
+% Maybe cut to three examples?
+  %Unfortunately, although widely used, the properties of some phylogenetic comparative methods are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
 
-%The majority of PCMs use an explicit evolutionary model to characterize trait evolution \citep{Freckleton:2011aa}.
-A model can be fit using statistical methods (typically maximum likelihood) and model parameters interrogated to test ideas from evolutionary theory.
   Most model-based methods for characterizing trait evolution are based on the Brownian constant variance model \citep[for exceptions see][]{price1997correlated,harvey2000comparative,freckleton2006detecting}.
   The Brownian model, first applied in a phylogenetic context by \citet{cavalli1967} and to across-species data by \citet{felsenstein1973maximum}, is a simple model of trait evolution in which trait variance accrues as a linear function of time, and makes the prediction that traits of closely-related species are more similar than those of distantly-related ones.
-The Brownian model has been modified in various ways to account for a suite of ecological and evolutionary processes (e.g., Grafen 1989; Hansen 1997; Pagel 1997, 1999).
-Most of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters.
-These modified Brownian models have become very popular because they fit better, have attractive biological interpretations, and are implemented in freely available packages \citep[e.g., R;][]{R-Core-Team:2014aa}.
-
-%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
-$\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
-It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$.
-Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously.
-\citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989).
-They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes).
-This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.
+  The Brownian model has been modified in various ways to account for a suite of ecological and evolutionary processes \citep[e.g.,][]{grafen1989phylogenetic,hansen1997stabilizing,Pagel:1997aa,Pagel:1999aa}.
+  The majority of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. 
+  These modified Brownian models tend to fit better, have attractive biological interpretations, and are implemented in freely available packages \citep[e.g., R;][]{R-Core-Team:2014aa}.
 
@@
   This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
+  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape. 
+  The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see Discussion).
+
   The popularity of the OU model has grown exponentially in recent years (Fig.
@@
       \centering
-    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure1.pdf}
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
       \caption{The number of evolutionary biology (closed circles) and ecology (open circles) papers published between 2005 and 2012 containing the phrase “Ornstein-Uhlenbeck”, as a proportion of the total number of evolutionary biology or ecology papers published that year.}
@@
     \end{figure}
-
-Uses of the Ornstein-Uhlenbeck can be divided into three broad categories that are near equally represented in the literature: (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses.
-Here, we focus on the first of these because the principles apply to a range of other macroevolutionary models.
 
@@
 
-\section{Materials and Methods}
+We are not the first to mention these problems, but we hope by
+summarising the issues and giving recommendations we can help those
+less familiar with the literature.
+
+  \subsection{What is the Ornstein-Uhlenbeck model?}
     According to the Brownian model \cite{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
@@
     The model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant. 
-Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.
-The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species.
-This means that close relatives will be more similar in their trait values than more distant relatives.
-It also means that variance in the trait will increase (linearly) in proportion to time.
+    Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species.
+    This means that close relatives will be more similar in their trait values than more distant relatives. It also means that variance in the trait will increase (linearly) in proportion to time. 
     The model has two parameters, the Brownian rate parameter, $\sigma^2$ and the state of the root at time zero, $X(0)$.
@@
 
-The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape.
-The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see Discussion).
+%In the light of the results below, some properties of this model are worth highlighting.
+When $\alpha$ is small, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent.
+Eventually, when $\alpha$ is really large, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
+\ref{figure.traitsim}C).
+This will be indistinguishable from phylogenetically unstructured data.
 
-The variances and covariances predicted by the OU model are more complex than those predicted by the Brownian model.
-In the light of the results below, some properties of this model are worth highlighting:
-
-\begin{enumerate}[(i)]
-
-  \item If $\alpha$ is small then evolution is approximately Brownian, The question is what is a small value of $\alpha$? Figure \ref{figure.traitsim} provides some guidance on this issue.\\
-
-  \item If species $i$ and $j$ diverged recently, evolution is approximately Brownian: if two species diverged recently, then $T - t_{ij} \approx 0$ and hence $cov(X_i, X_j) \approx 1 - e^{-2\alpha t_{ij}} \approx 2\alpha t_{ij}$.
-Thus, recently diverged species provide little information relevant to estimating non-Brownian evolution according to an OU process.
-\\
-
-  \item In the long-term the imprint of history is weakened.
-If evolution proceeds for a long time the variance in $X_i$ tends to a constant, i.e.
-because the expected value of $X_i$ is $\mu$.
-Similarly, in equation (4), the covariance between traits tends to a constant because $T$ becomes large relative to $t_{ij}$.
-Consequently for large groups the model implies that the imprint of history is weak.
-As a result adding more species to a phylogeny will contribute only relatively weakly to fitting the model, as the deeper branches within a phylogeny will contribute little information to estimating the parameters.\\
-
-\end{enumerate}
-
-Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
+%Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
 We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
@@
 
-Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
-\ref{figure.traitsim}C).
-This will be indistinguishable from phylogenetically unstructured data.
+\subsection{What do people use the OU model for?}
+Uses of the Ornstein-Uhlenbeck can be divided into three broad categories that are near equally represented in the literature: (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses.
+Here, we focus on the first of these because the principles apply to a range of other macroevolutionary models.
 
-\begin{figure}
-\centering
-    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
-\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
-The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
-equal to correlation between species’ traits).
-This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
-}
-\label{figure.traitsim}
-\end{figure}
-
-In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0-0.5 in Fig.
-\ref{figure.traitsim}A; 10-50 in Fig.
-\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1-5) that the biggest discernible effects occur.
-
-The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
-
-\subsection{Simulations}
+\section{Materials and Methods}
+  \subsection{Maximum Likelihood simulations}
   \label{section:sims.methods} 
@@
   We simulated phylogenies with 25, 50, 100, 150, 200, 500, or 1000 tips under pure birth, constant-rate Birth-Death (extinction fractions of 0.25, 0.5 and 0.75), or temporally varying speciation rate (speciation rate modeled as time from the root raised to the power 0.2, 0.5, 2 and 5) models.
-We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total.
-Trees were simulated using the R package TESS \cite{hohna2013fast}.
+  We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total. Trees were simulated using the R package TESS \cite{hohna2013fast}. 
   We then simulated the evolution of a single trait under a Brownian motion model on each phylogeny using the R package MOTMOT \citep{Thomas:2011aa}. 
-We estimated α and compared the fit of a Brownian model to that of an OU model using: (i) a likelihood ratio test with 1 degree of freedom with the transformPhylo.ML function in MOTMOT; and, (ii) with Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits.
+  We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using: (i) a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
+
+\subsection{Effects of measurement error}
+  We used the same procedure to simulate trait data under a Brownian motion model with known error. 
+  Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees.
+We then simulated data under a Brownian model on each tree.
+  We compared the fit of a Brownian model to that of an OU model as described above, using the original trees without the addition of extra branch length to the tip edges. 
+  The expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.
+
+\subsection{Bayesian simulations}
+
+and, (ii) with Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits.
 The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1).
@@
 
-We used the same procedure to simulate trait data under a Brownian motion model with known error.
-Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees.
-We then simulated data under a Brownian model on each tree.
-We compared the fit of a Brownian model to that of an OU model (with ML and MCMC stepping stone sampling as above) using the original trees without the addition of extra branch length to the tip edges.
-The expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.
-
+\subsection{Other models}
 Although our focus is on the OU model we also explored three other commonly used models \citep[$\kappa$, $\lambda$, and $\delta$;][]{Pagel:1997aa,Pagel:1999aa} fitted using maximum likelihood using the same simulated data.
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure3.pdf}
-  \caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ‘true’ value of alpha is 0.}
-\label{figure.profiles}
-\end{figure}
+%\begin{figure}
+ % \centering
+  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure3.pdf}
+  %\caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ‘true’ value of alpha is 0.}
+%\label{figure.profiles}
+%\end{figure}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure4.pdf}
-  \caption{Rejection rates and estimation of $\alpha$ for different tree sizes and shapes.
+%\begin{figure}
+  %\centering
+  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure4.pdf}
+  %\caption{Rejection rates and estimation of $\alpha$ for different tree sizes and shapes.
 Rejection rates (a,c,e) for the OU model at multiple tree sizes for models fitted using ML (circles) and Bayesian (diamonds) methods.
 Estimates of $\alpha$ (b,d,f) based on ML (circles show means and dotted lines show 95\% sampling interval from 1000 trees) and Bayesian (diamonds show means and dotted lines show 95\% sampling interval of the modal value from the posterior distribution from 1000 trees).}
-\label{figure.rejection}
-\end{figure}
+%\label{figure.rejection}
+%\end{figure}
 
-Figure \ref{figure.error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs.
+%Figure \ref{figure.error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs.
 S4 and S5).
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure5.pdf}
-  \caption{Rejection of BM with measurement error (a) and estimation of α (b). Symbols follow Figure \ref{figure.rejection}.}
-\label{figure.error}
-\end{figure}
+%\begin{figure}
+ % \centering
+  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure5.pdf}
+  %\caption{Rejection of BM with measurement error (a) and estimation of α (b). Symbols follow Figure \ref{figure.rejection}.}
+%\label{figure.error}
+%\end{figure}
 
@@
 \label{section:discussion}
+
+Recommendations
+
+Interpreting results
+ - OU is favoured over BM. If alpha is small it's basically BM. What
+is alpha? Talk about half lives.
+ - Not stabilising selection
+ - Not an issue if you're just mopping up error structure for PGLS (?)
+ - If really want to be certain OU over BM you need 200 species
+ - Account for measurement error
+
+%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
+$\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
+It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$.
+Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously.
+\citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989).
+They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes).
+This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.
+
 Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways.

### 2015-04-07 Analyses/OhYouFigures.R
Added paleo dtat to figure 1

@@ -6,6 +6,7 @@ ds <- read.delim("Data/papers.txt")
 # Calculate %
 ds$eco <- (ds$ecology / ds$ecology.total) * 100
 ds$evo <- (ds$evolution / ds$evolution.total) * 100
+ds$pal <- (ds$paleo / ds$paleo.total) * 100
 
 # Plot figure 1
 
@@ -14,5 +15,6 @@ par(bty = "l")
 plot(evo ~ Year, data = ds, las = 1, pch = 16, cex = 1.2, cex.lab = 1.2, cex.axis = 1.2,
      ylab = "OU publications (% of total papers)", xlab = "Year published", ylim = c(0, 0.7))
 points(eco ~ Year, data = ds, cex = 1.2)
-legend("topleft", legend = c("evolution", "ecology"), pch = c(16, 1), cex = 1.2, bty = "n")
+points(pal ~ Year, data = ds, cex = 1.2, pch = 6)
+legend("topleft", legend = c("ecology", "evolution", "palaeontology"), pch = c(1, 16, 6), cex = 1.2, bty = "n")
 dev.off()

### 2015-04-07 Analyses/OhYouFigures.R
Added paleo dtat to figure 1

@@ -10,7 +10,7 @@ ds$pal <- (ds$paleo / ds$paleo.total) * 100
 
 # Plot figure 1
 
-png("Figures/PapersThruTime.png")
+png("Manuscript/Figures/PapersThruTime.png")
 par(bty = "l")
 plot(evo ~ Year, data = ds, las = 1, pch = 16, cex = 1.2, cex.lab = 1.2, cex.axis = 1.2,
      ylab = "OU publications (% of total papers)", xlab = "Year published", ylim = c(0, 0.7))

### 2015-04-07 Manuscript/ohyou.tex
Completed most of intriduction

@@
 
-\runninghead{Limits of comparative analyses}
+\runninghead{A cautionary note on OU models}
 \keywords{Bayesian method - comparative methods - delta - kappa - lambda - OU - phylogeny - stabilizing selection - selection} %meant to be after abstract
@@
   The majority of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. 
-  These modified Brownian models tend to fit better, have attractive biological interpretations, and are implemented in freely available packages \citep[e.g., R;][]{R-Core-Team:2014aa}.
+  These modified Brownian models tend to fit better and often have attractive biological interpretations.
 
@@
   This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
-  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape. 
+  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait mean was recast as a fitness optimum on an adaptive landscape. 
   The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see Discussion).
 
-  The popularity of the OU model has grown exponentially in recent years (Fig.
-\ref{figure.literature}), in part because these models are now easy to implement apply via packages in R \citep[e.g.
-ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}.
-OU models have become particularly prevalent in ecological studies: 652 papers containing the phrases “Ornstein Uhlenbeck” and “ecology” were published between 2011 and 2013 (Google Scholar search 30th January 2014).
-Additionally, although they are pattern based, fit to an OU model is now being used as evidence for the action of ecological, or ecologically driven, processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  Here we investigate the OU model in more detail. 
+  We present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+We also show that very small amounts of measurement error in data can have profound effects on the performance of models.
+  Many of our findings are applicable to other models of evolution (see Discussion and Supplemental Figs. S6-S11) 
+  % Add figure refs here
+  , but we focus on the OU model because of the recent increase in publications using the model (see below) and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
+  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic}.
 
-    \begin{figure}
-      \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
-      \caption{The number of evolutionary biology (closed circles) and ecology (open circles) papers published between 2005 and 2012 containing the phrase “Ornstein-Uhlenbeck”, as a proportion of the total number of evolutionary biology or ecology papers published that year.}
-      \label{figure.literature}
-    \end{figure}
-
-Here we investigate the OU model because it does not appear to have been subject to a rigorous critique. 
-The $\alpha$ parameter of the OU model is very similar to Grafen’s $\rho$ (along with other parameters in the literature) in that it is in effect a non-linear transformation of the phylogeny. 
-Consequently, one might imagine that it shares some of the same statistical properties. 
-Results from \citet{ho2013asymptotic}, indicate that there is the potential for revealing strong biases in $\alpha$ , but the implications for practice (e.g.
-likelihood of rejection of alternative models) is unexplored other than to note that tree shape may play an influence.
-We present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in comparative analyses.
-We also show that very small amounts of measurement error in data can have profound effects on the performance of models.
-Many of our findings are applicable to other models of evolution (see Discussion and Supplemental Figs.
-S6-S11), but we focus here on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the α parameter.
-
-We are not the first to mention these problems, but we hope by
-summarising the issues and giving recommendations we can help those
-less familiar with the literature.
-
-  \subsection{What is the Ornstein-Uhlenbeck model?}
-    According to the Brownian model \cite{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
+  \subsection{What is the OU model?}
+    According to the Brownian model \citep{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
 
@@
     The model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant. 
-    Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species.
+    Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.
+    The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species. 
     This means that close relatives will be more similar in their trait values than more distant relatives. It also means that variance in the trait will increase (linearly) in proportion to time. 
@@
 
-    The Ornstein-Uhlenbeck (OU) model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
+    The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
     The model has the following form:
@@
     $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. 
-
-%In the light of the results below, some properties of this model are worth highlighting.
-When $\alpha$ is small, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent.
-Eventually, when $\alpha$ is really large, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
-\ref{figure.traitsim}C).
-This will be indistinguishable from phylogenetically unstructured data.
-
-%Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
-We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
-Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
-When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
-Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent.
-Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
-\ref{figure.traitsim}B).
+    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger (around 0.5 in our simulations) the non-Brownian behaviour of the model starts to become apparent.
+    Eventually, when $\alpha$ is really large (i.e., greateer than XXX), all imprint of history is lost and the evolution is essentially a rapid burst at the present (FIGURE 2). 
+    This will be indistinguishable from phylogenetically unstructured data. % needs better phrasing
+    % Fix the figure or send to supplemental
 
   \subsection{What do people use the OU model for?}
-Uses of the Ornstein-Uhlenbeck can be divided into three broad categories that are near equally represented in the literature: (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses.
-Here, we focus on the first of these because the principles apply to a range of other macroevolutionary models.
+    The popularity of the OU model has grown exponentially in recent years (Fig.
+\ref{figure.literature}), in part because these models are now easy to implement apply via packages in R \citep[e.g.
+ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}.
+    OU models have become particularly prevalent in ecological studies: 829 papers containing the phrases ``Ornstein Uhlenbeck'' and ``ecology'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
+% Need to add ref to supp mat
+    Additionally, although they are pattern based, fit to an OU model is now being used as evidence for the action of ecological, or ecologically driven, processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+OU models are also increasingly used in palaeontological studies \citep[e.g.,][]{slater2013,soul2015taxonomy}.
+
+      \begin{figure}[h]
+        \centering
+        \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
+        \caption{The number of ecology (open circles), evolutionary biology (closed circles) and palaeontology (open triangles) papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.}
+        \label{figure.literature}
+      \end{figure}
+
+    Uses of the OU model can be divided into three broad categories that are near equally represented in the literature (see Supplementary Material) % Need to add ref to supp mat
+    : (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses.
+    In our simulations (see below), we focus on the first of these because the principles apply to a range of other macroevolutionary models.
 
@@
     We simulated phylogenies with 25, 50, 100, 150, 200, 500, or 1000 tips under pure birth, constant-rate Birth-Death (extinction fractions of 0.25, 0.5 and 0.75), or temporally varying speciation rate (speciation rate modeled as time from the root raised to the power 0.2, 0.5, 2 and 5) models.
-  We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total. Trees were simulated using the R package TESS \cite{hohna2013fast}. 
+    We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total. 
+    Trees were simulated using the R package TESS \cite{hohna2013fast}. 
     We then simulated the evolution of a single trait under a Brownian motion model on each phylogeny using the R package MOTMOT \citep{Thomas:2011aa}. 
-  We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using: (i) a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
+    We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
 
   \subsection{Effects of measurement error}
-  We used the same procedure to simulate trait data under a Brownian motion model with known error. 
+    Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015} so we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
+    We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
     Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees.
     We then simulated data under a Brownian model on each tree. 
-  We compared the fit of a Brownian model to that of an OU model as described above, using the original trees without the addition of extra branch length to the tip edges. 
-  The expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.
+    We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges.
+    Our expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.
 
@@
 
-\subsection{Other models}
-Although our focus is on the OU model we also explored three other commonly used models \citep[$\kappa$, $\lambda$, and $\delta$;][]{Pagel:1997aa,Pagel:1999aa} fitted using maximum likelihood using the same simulated data.
+  \subsection{Other evolutionary models}
+    Although our focus is on the OU model we also explored three other commonly used models of trait evolution -- $\kappa$, $\lambda$, and $\delta$ \citep{Pagel:1997aa,Pagel:1999aa} -- fitted using Maximum Likelihood with the same simulated data.
     These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
@@
   \label{section:sims.results}
+
     Figure \ref{figure.profiles} shows some examples of profile likelihoods for α generated on different datasets that were simulated according to a Brownian model.
@@
  - Account for measurement error
+
+%Results from \citet{ho2013asymptotic}, indicate that there is the potential for revealing strong biases in $\alpha$ , but the implications for practice (e.g.
+likelihood of rejection of alternative models) is unexplored other than to note that tree shape may play an influence.
 
@@
   \label{section:literature.methods} 
-  To get an overview of the use of OU models in ecology and evolution, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution'').
-We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'' published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
+  To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''.
+We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'', or ``paleo/palaeo'', published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
 \ref{figure.literature}).
 
-  Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following ecology and evolution journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
+  Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
 For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other).

### 2015-04-09 Manuscript/ohyou.tex
Edits to intro and appendix

@@
 \usepackage{enumerate} 
+%\usepackage[nomarkers, nolists]{endfloat}  % To push floats to the end   
 \pagenumbering{arabic}
@@
   Phylogenetic comparative methods are powerful tools for identifying patterns in the evolution of species traits, and for potentially inferring the evolutionary processes that underlie them \citep[e.g.,][]{freckleton2009seven,Nunn:2011aa,o2012evolutionary,pennell2013integrative}.
-  These approaches have been used, for example, to model niche conservatism \citep{Wiens:2010aa}, infer potential rates of species responses to climate change \citep{Quintero:2013aa}, test the role of ecological niche as a driver of morphological evolution \citep{pienaar2013macroevolution}, compare rates of evolution of species traits \citep{Claramunt:2012ab} and test for constraints in adaptive radiations \citep{blackburn2013adaptive}.
-% Maybe cut to three examples?
+  These approaches have been used, for example, to infer potential rates of species responses to climate change \citep{Quintero:2013aa}, test the role of ecological niche as a driver of morphological evolution \citep{pienaar2013macroevolution}, and test for constraints in adaptive radiations \citep{blackburn2013adaptive}.
   %Unfortunately, although widely used, the properties of some phylogenetic comparative methods are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
@@
   % Add figure refs here
-  , but we focus on the OU model because of the recent increase in publications using the model (see below) and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic}.
+  , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
+  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your}.
+%ives garland 2010, hansen bartoszek et al
 
+\section{Materials and Methods}
   \subsection{What is the OU model?}
@@
         dX(t) = \sigma dW(t)
-      \label{equation:BMrate} 
       \end{equation}
@@
         dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
-        \label{equation:OUrate} 
       \end{equation}
@@
     $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
-    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. 
-    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger (around 0.5 in our simulations) the non-Brownian behaviour of the model starts to become apparent.
-    Eventually, when $\alpha$ is really large (i.e., greateer than XXX), all imprint of history is lost and the evolution is essentially a rapid burst at the present (FIGURE 2). 
-    This will be indistinguishable from phylogenetically unstructured data. % needs better phrasing
+    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see Appendix.
+
+    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
+    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2).
+    We discuss how to interpret values of $\alpha$ below.
+
     % Fix the figure or send to supplemental
@@
   \subsection{What do people use the OU model for?}
-    The popularity of the OU model has grown exponentially in recent years (Fig.
+    The popularity of the OU model has grown extensively in recent years (Fig.
 \ref{figure.literature}), in part because these models are now easy to implement apply via packages in R \citep[e.g.
 ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}.
+
     OU models have become particularly prevalent in ecological studies: 829 papers containing the phrases ``Ornstein Uhlenbeck'' and ``ecology'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
@@
 
-\section{Materials and Methods}
+     % OK somewhere here need to clarify that a) most ou studies are looking at fitting the model, and comparing models, not using multiple optima stuff.
+Then clarify that here we are focussing on that popular use of the
+
+     % Problems with the OU model and where they are reported in the literature?
+
+     % This maybe belongs in the main text in shorter form  
+  Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
+We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
+Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
+When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
+Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent.
+Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
+\ref{figure.traitsim}B).
+Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
+\ref{figure.traitsim}C).
+This will be indistinguishable from phylogenetically unstructured data.
+
+\begin{figure}
+\centering
+    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
+\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
+The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
+equal to correlation between species’ traits).
+This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
+}
+\label{figure.traitsim}
+\end{figure}
+
+In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
+\ref{figure.traitsim}A; 10-50 in Fig.
+\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
+
+The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
+
   \subsection{Maximum Likelihood simulations}
@@
 
-    Figure \ref{figure.profiles} shows some examples of profile likelihoods for α generated on different datasets that were simulated according to a Brownian model.
+    Figure \ref{figure.profiles} shows some examples of profile likelihoods for $\alpha$ generated on different datasets that were simulated according to a Brownian model.
 It is clear that the shape of the tree affects the shape of the likelihood profile.
@@
 For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips.
-Bayesian estimation is not subject to bias in the log-likelihood statistic and consistently rejects the OU model.
+Bayesian estimation is not subject to bias % check in the log-likelihood statistic and consistently rejects the OU model.
 However, estimates of α are similar regardless of method.
@@
 
-The biases reported above are not unique to the OU model.
+The bias % check
+ reported above are not unique to the OU model.
 The parameter $\delta$, which tests for accelerating or decelerating rates of evolution \citep{Pagel:1997aa,Pagel:1999aa} also suffers from elevated Type I error and is upwardly biased, favoring models that imply temporally increasing evolutionary rates (Supplemental Fig.
@@
  - OU is favoured over BM. If alpha is small it's basically BM. What
-is alpha? Talk about half lives.
+is alpha? Talk about half lives. add in ives and garland advice on what is big and what is small here?
  - Not stabilising selection
@@
 \section{Appendix}
-\subsection{Model Outline} % correct the equations!
   \label{section:models}
+  \setcounter{equation}{0}
+
   According to the Brownian model, a trait X evolves at random at a rate $\sigma$:
@@
       dX(t) = \sigma dW(t)
-    %\label{equation:BMrate} 
+      \label{equation:BMrate} 
     \end{equation}
@@
 
-where $t_{ij}$ is the shared evolutionary pathway for species $i$ and $j$, i.e.
-the time at which they last shared a common ancestor.
+  where $t_{ij}$ is the shared evolutionary pathway for species $i$ and $j$, i.e. the time at which they last shared a common ancestor. 
   Equations \ref{equation:BMvar} and \ref{equation:BMcov} encapsulate the simplicity of the Brownian model, namely it predicts that variances accrue as a linear function of time.
@@
     dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
-    %\label{equation:OUrate} 
+    \label{equation:OUrate} 
   \end{equation}
@@
   $\sigma$ is the strength of evolutionary force that returns traits back towards the mean if they evolve away. 
-The OU model was introduced to population genetics by \citet{Lande:1976aa} to model stabilizing selection in which the mean was recast as a fitness optimum on an adaptive landscape.
-The process operating in comparative data is analogous, although clearly is not stabilizing selection despite being sometimes referred to as such (see Discussion).
   This model has two parameters in addition to those of the Brownian model, $\alpha$ and $\mu$.
@@
 
+% Still not sure of these two equations!
     \begin{equation}
-    var(X_i) = 1 - e^{-2 \alpha T}
+      var(X_i) = \frac{(1 - e^{-2 \alpha T}) \times \sigma^2}{2 \alpha}
       \label{equation:OUvar} 
@@
     \begin{equation}
-    cov(X_i, X_j) = e^{-2 \alpha (T - t_{ij})} (1 - e^{-2\alpha t_{ij}})
+      cov(X_i, X_j) = \frac{e^{-2 \alpha (T - t_{ij})} \times (1 - e^{-2\alpha t_{ij}})\times \sigma^2}{2 \alpha}
       \label{equation:OUcov} 
@@
   The variances and covariances predicted by equations \ref{equation:OUvar} and \ref{equation:OUcov} are more complex than those predicted by the Brownian model. 
-In the light of the results below, some properties of this model are worth highlighting:
+  In the light of the results above, some properties of this model are worth highlighting:
 
   \begin{enumerate}[(i)]
-
     \item If $\alpha$ is small then evolution is approximately Brownian: If $\alpha$ is small then $1 - e^{-2\alpha T} \approx 2\alpha T$, i.e.
-traits accrue variance as if evolving according to a Brownian process.
-The question is obviously what is a small value of $\alpha$?
-Figure \ref{figure.traitsim} provides some guidance on this issue, and see point (iv) below.\\
+traits accrue variance as if evolving according to a Brownian process.\\
 
     \item If species $i$ and $j$ diverged recently, evolution is approximately Brownian: if two species diverged recently, then $T - t_{ij} \approx 0$ and hence $cov(X_i, X_j) \approx 1 - e^{-2\alpha t_{ij}} \approx 2\alpha t_{ij}$.
-Thus, recently diverged species provide little information relevant to estimating non-Brownian evolution according to an OU process.
-\\
+    Thus, recently diverged species provide little information relevant to estimating non-Brownian evolution according to an OU process.\\
 
-  \item In the long-term the imprint of history is weakened: if $T$ is large (i.e.
-evolution proceeds for a long time), equation ???
-predicts that the variance in $X_i$ tends to a constant, i.e.
+    \item In the long-term, the imprint of history is weakened: if $T$ is large (i.e.
+evolution proceeds for a long time), equation \ref{equation:OUvar} predicts that the variance in $X_i$ tends to a constant, i.e.
 because the expected value of $X_i$ is $\mu$.
-Similarly, in equation (4), the covariance between traits tends to a constant because $T$ becomes large relative to $t_{ij}$.
-Consequently for large groups the model implies that the imprint of history is weak.
-As a result adding more species to a phylogeny will contribute only relatively weakly to fitting the model, as the deeper branches within a phylogeny will contribute little information to estimating the parameters.\\
-
+    Similarly, in equation \ref{equation:OUrate}, the covariance between traits tends to a constant because $T$ becomes large relative to $t_{ij}$.
+    Consequently for large groups the model implies that the imprint of history is weak.\\
   \end{enumerate}
 
-Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
-We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
-Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
-When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
-Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent.
-Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
-\ref{figure.traitsim}B).
-Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
-\ref{figure.traitsim}C).
-This will be indistinguishable from phylogenetically unstructured data.
-
-\begin{figure}
-\centering
-    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
-\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
-The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
-equal to correlation between species’ traits).
-This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
-}
-\label{figure.traitsim}
-\end{figure}
-
-In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
-\ref{figure.traitsim}A; 10-50 in Fig.
-\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
-
-The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
-
-\section{Supplementary Material}
-
-  \subsection{Literature Review}
-  \label{section:literature.methods} 
-  To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''.
-We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'', or ``paleo/palaeo'', published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
-\ref{figure.literature}).
-
-  Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers) published in the following journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
-For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other).
-Where papers included multiple analyses using different numbers of species we used the median number of taxa.
-Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category.
-We summarise these results in Figure S1 and the full dataset is available in Supplemental Table S1 along with the full list of references.
-
-  \label{section:literature.results}
-  In total, 3720 papers containing the terms “Ornstein Uhlenbeck” and either “ecology”, or “evolution” and “biology” were published between 2005 and 2013, and the number has increased substantially since 2005 (Fig.
-\ref{figure:literature}).
-Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig.
-S1 and Table S1).
-The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
-S1).
+\input{ohyou_suppmat}
 

### 2015-04-09 Manuscript/ohyou_suppmat.tex
Edits to intro and appendix

\section{Supplementary Material}

  \subsection{Literature Review}
    To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''. 
    We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'', or ``paleo/palaeo'', published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig. \ref{figure.literature}).

    Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers with empirical proof of concept) published in the following journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science. 
    For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other). 
    Where papers included multiple analyses using different numbers of species we used the median number of taxa. 
    Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. 
    We summarise these results in Fig. %figure link
    and the full dataset is available in Supplemental Table S1 along with the full list of references.

    In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. \ref{figure:literature}). 
    Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig. S1 and Table S1). 
    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig. S1).

### 2015-04-10 Manuscript/ohyou.tex
Edits to intro and fixed equations

@@
 
-\runninghead{A cautionary note on OU models}
+\runninghead{A cautionary note on macroevolutionary models}
 \keywords{Bayesian method - comparative methods - delta - kappa - lambda - OU - phylogeny - stabilizing selection - selection} %meant to be after abstract
@@
   The majority of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. 
-  These modified Brownian models tend to fit better and often have attractive biological interpretations.
+  These modified Brownian models tend to fit better and often have links to process-based interpretations.
 
@@
 
-  Here we investigate the OU model in more detail. 
-  We present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+  The popularity of the OU model has grown extensively in recent years (Fig.
+\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
+% Need to add ref to supp mat
+  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
+  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
+  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
+
+    \begin{figure}[h]
+      \centering
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
+      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+See Supplementary Material for details %linkto supp mat
+      }
+      \label{figure.literature}
+    \end{figure}
+
+  Uses of the OU model can be divided into three broad categories. % improve citations in this section
+    \begin{enumerate}
+      \item \textbf{Modeling the evolution of a continuous character.}\\ 
+            Generally in these analyses, multiple models of evolution are fit to the same continuous character \citep[e.g.,][]{Harmon:2010aa}.
+Model selection is then used to determine which model best fits the data.
+            If the OU model is favoured this is often used as evidence of evolutionary constraints, stabilizing selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}
+      \item \textbf{Modeling the effects of different hypothesized adaptive regimes on a continuous character.}\\
+            OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
+            The latter represents evolution under multiple selective regimes, and is probably more biological realistic. 
+            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012ouwie}.
+      \item \textbf{Controlling for phylogeny in correlative analyses.}\\ 
+            In these analyses, . This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the models \citep[e.g.,][]{}.%cite maia?
+    \end{enumerate}
+
+  \noindent
+  The OU model is also used to reconstruct ancestral states % citation
+  and to detect clade-wide convergent evolution \citep{ingram2013surface}. 
+  The majority of papers use the OU model to model the evolution of a continuous character (Fig 2 and Supplementary Material) % Need to add ref to supp mat
+  Therefore, we focus on this use of the OU model in our simulations below.
+The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures.
+
+    \begin{figure}[h]
+      \centering
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
+      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+See Supplementary Material for details %linkto supp mat
+      }
+      \label{figure.uses}
+    \end{figure}
+
+  Here we present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
 We also show that very small amounts of measurement error in data can have profound effects on the performance of models.
@@
   , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your}.
-%ives garland 2010, hansen bartoszek et al
+  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}.
 
 \section{Materials and Methods}
-  \subsection{What is the OU model?}
+  \subsection{The Ornstein Uhlenbeck (OU) model}
     According to the Brownian model \citep{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
@@
     $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
-    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see Appendix.
+    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see the Appendix.
 
     When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
-    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2).
+    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2). % Fix the figure or send to supplemental
     We discuss how to interpret values of $\alpha$ below.
 
-    % Fix the figure or send to supplemental
-
-  \subsection{What do people use the OU model for?}
-    The popularity of the OU model has grown extensively in recent years (Fig.
-\ref{figure.literature}), in part because these models are now easy to implement apply via packages in R \citep[e.g.
-ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}.
-
-    OU models have become particularly prevalent in ecological studies: 829 papers containing the phrases ``Ornstein Uhlenbeck'' and ``ecology'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
-% Need to add ref to supp mat
-    Additionally, although they are pattern based, fit to an OU model is now being used as evidence for the action of ecological, or ecologically driven, processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
-OU models are also increasingly used in palaeontological studies \citep[e.g.,][]{slater2013,soul2015taxonomy}.
-
-      \begin{figure}[h]
-        \centering
-        \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
-        \caption{The number of ecology (open circles), evolutionary biology (closed circles) and palaeontology (open triangles) papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.}
-        \label{figure.literature}
-      \end{figure}
-
-    Uses of the OU model can be divided into three broad categories that are near equally represented in the literature (see Supplementary Material) % Need to add ref to supp mat
-    : (i) modeling the evolution of a continuous character to test for evolutionary constraints or stabilizing selection; (ii) modeling the effects of different hypothesized adaptive regimes on a continuous character and, (iii) controlling for phylogeny in correlative analyses.
-    In our simulations (see below), we focus on the first of these because the principles apply to a range of other macroevolutionary models.
-
-     % OK somewhere here need to clarify that a) most ou studies are looking at fitting the model, and comparing models, not using multiple optima stuff.
-Then clarify that here we are focussing on that popular use of the
-
+  \subsection{Known statistical issues with the OU model}
      % Problems with the OU model and where they are reported in the literature?
@@
      % This maybe belongs in the main text in shorter form  
-  Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
+  %Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
 We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
@@
 
-In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
+%In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
 \ref{figure.traitsim}A; 10-50 in Fig.
@@
 
-The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
+%The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
 
@@
     We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
+    This procedure ... % explain WHY we are doing this, how does it relate to uses of the OU model.
 
@@
 
-  \subsection{Bayesian simulations}
-
-and, (ii) with Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits.
+  \subsection{Bayesian simulations} % keep? or move it all to supp mat?
+    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
+    The choice of priors for OU models has not been fully explored. 
+    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
+    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
+    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
+S2-S5).
+% add links to this
+    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
+    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
+    Multiple independent chains were run for each analysis to ensure convergence was reached. 
+    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
     The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
@@
     We treated Bayes factors $>$ 2 as evidence favouring the OU model.
-We used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
-For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 for $\sigma^2$.
-We also normalized the tree length to prevent the $\alpha$ parameter of the OU model from interacting with $\sigma^2$.
-
-OU models have not previously been widely applied within a Bayesian framework, although some packages offer this functionality \citep[e.g., diversitree;][]{FitzJohn:2012aa} and the choice of prior has not been fully explored.
-We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior.
-We provide results derived from a strong exponential prior with mean = 1 and a bounded uniform prior (range 0-20) as supplementary material (Figs.
-S2-S5).
-We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
-Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample.
-Multiple independent chains were run for each analysis to ensure convergence was reached.
 
@@
     These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
-    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the supplementary material.
+    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the Supplementary Material.
 
@@
 
-% Still not sure of these two equations!
     \begin{equation}
-      var(X_i) = \frac{(1 - e^{-2 \alpha T}) \times \sigma^2}{2 \alpha}
+      var(X_i) = \frac{\sigma^2}{2 \alpha} 1 - e^{-2 \alpha T}
       \label{equation:OUvar} 
@@
     \begin{equation}
-      cov(X_i, X_j) = \frac{e^{-2 \alpha (T - t_{ij})} \times (1 - e^{-2\alpha t_{ij}})\times \sigma^2}{2 \alpha}
+      cov(X_i, X_j) = \frac{\sigma^2}{2 \alpha} e^{-2 \alpha (T - t_{ij})} (1 - e^{-2\alpha t_{ij}})
       \label{equation:OUcov} 

### 2015-04-13 Manuscript/ohyou.tex
Minor edits to intro:

@@
 
-  Uses of the OU model can be divided into three broad categories. % improve citations in this section
+  Uses of the OU model can be divided into three broad categories. % improve citations in this section?
     \begin{enumerate}
@@
 Model selection is then used to determine which model best fits the data.
-            If the OU model is favoured this is often used as evidence of evolutionary constraints, stabilizing selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}
-      \item \textbf{Modeling the effects of different hypothesized adaptive regimes on a continuous character.}\\
+            If the OU model is favoured, this is often used as evidence of evolutionary constraints, stabilising selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}.
+            OU models with different numbers of optima may also be included in the set of models considered (see below).
+      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character.}\\
             OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
-            The latter represents evolution under multiple selective regimes, and is probably more biological realistic. 
-            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012ouwie}.
+            The latter represents evolution under multiple selective regimes, and is probably more biologically realistic. 
+            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012modeling}.
       \item \textbf{Controlling for phylogeny in correlative analyses.}\\ 
-            In these analyses, . This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the models \citep[e.g.,][]{}.%cite maia?
+            Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
+            This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data.
+            In these analyses, the OU model is used to transform the tree and the variance-covariance matrix.
+            This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
     \end{enumerate}
@@
   \noindent
-  The OU model is also used to reconstruct ancestral states % citation
+  The OU model is also used to reconstruct ancestral states \citep{martins1999estimation}
   and to detect clade-wide convergent evolution \citep{ingram2013surface}. 
-  The majority of papers use the OU model to model the evolution of a continuous character (Fig 2 and Supplementary Material) % Need to add ref to supp mat
-  Therefore, we focus on this use of the OU model in our simulations below.
+  The majority of papers use the OU model to model the evolution of a continuous character (Fig 2 and Supplementary Material), % Need to add ref to supp mat
+  therefore, we focus on this use of the OU model in our simulations below. 
   The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures.

### 2015-04-13 Manuscript/ohyou.tex
Edits to 3 uses, thinking to change to two uses...

@@
   Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
-  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
   Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
@@
 
-  Uses of the OU model can be divided into three broad categories. % improve citations in this section?
+  Uses of the OU model can be divided into three broad categories. % improve citations in this section? % fix up numbers. Argggggggghhhhhhhhhhh.
     \begin{enumerate}
-      \item \textbf{Modeling the evolution of a continuous character.}\\ 
+      \item \textbf{Modeling the evolution of a continuous character (N = 58).}\\ 
             Generally in these analyses, multiple models of evolution are fit to the same continuous character \citep[e.g.,][]{Harmon:2010aa}.
@@
             If the OU model is favoured, this is often used as evidence of evolutionary constraints, stabilising selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}.
-            OU models with different numbers of optima may also be included in the set of models considered (see below).
-      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character.}\\
+      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character (N = 27).}\\
             OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
@@
             In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012modeling}.
-      \item \textbf{Controlling for phylogeny in correlative analyses.}\\ 
+            Note that the first and second uses of the OU model are often combined, so that OU models with various numbers of optima are included in the pool of models being tested (N = 27).
+      \item \textbf{Controlling for phylogeny in correlative analyses (N = 35).}\\ 
             Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
@@
 
-    \begin{figure}[h]
-      \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
-      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supplementary Material for details %linkto supp mat
-      }
-      \label{figure.uses}
-    \end{figure}
+    \begin{table}[h]
+      \begin{centering}
+        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013.
+See Supplementary Material for details linkto supp mat.}
+        \bigskip  
+          \begin{tabular}{p{8cm}lc}
+            \hline
+            Use of OU model & Optima & Number of papers\\
+            \hline
+            Ancestral state reconstruction & single & 8\\
+            & multiple & 2\\
+            Convergent evolution & single & 0\\
+            & multiple & 2\\
+            Model of evolution & single & 31\\
+            & multiple & 27\\
+            Phylogenetic generalised least squares & single & 35\\
+            & multiple & 0\\
+            Other & single & 5\\
+            & multiple & 5\\
+            Total & single &  36\\
+            & multiple & 79\\
+            \hline
+            \label{table.uses}
+          \end{tabular}
+    \end{table}
 
@@
     Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2). % Fix the figure or send to supplemental
-    We discuss how to interpret values of $\alpha$ below.
+
+    %$\alpha$ can most easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+
+      \begin{equation}
+        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
+        \label{equation:halflife}
+      \end{equation}
+
+    %If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, adaptation is quick, residual phylogenetic correlations are weak, and there is little influence of the past.
+This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term
 

### 2015-04-13 Analyses/OhYouFigures.R
Fixed errors and collated lit review

@@ -9,7 +9,6 @@ ds$evo <- (ds$evolution / ds$evolution.total) * 100
 ds$pal <- (ds$paleo / ds$paleo.total) * 100
 
 # Plot figure 1
-
 png("Manuscript/Figures/PapersThruTime.png")
 par(bty = "l")
 plot(evo ~ Year, data = ds, las = 1, pch = 16, cex = 1.2, cex.lab = 1.2, cex.axis = 1.2,
@@ -18,3 +17,17 @@ points(eco ~ Year, data = ds, cex = 1.2)
 points(pal ~ Year, data = ds, cex = 1.2, pch = 6)
 legend("topleft", legend = c("ecology", "evolution", "palaeontology"), pch = c(1, 16, 6), cex = 1.2, bty = "n")
 dev.off()
+
+# Read in data for figure 2 and reshape
+ds1 <- read.delim("Data/literature.txt")
+ds1 <- subset(ds1, Noptima != "?")
+ds1$Noptima <- ds1$Noptima[, drop = TRUE]
+
+omit <- which(ds1$General.use == "other" | ds1$General.use == "convergent evolution" |
+	  ds1$General.use == "ancestral state reconstruction" | 
+	  ds1$General.use == "phylogenetic inertia" | ds1$General.use == "phylogenetic signal" )
+
+ds2 <- ds1[-omit, ]
+ds2$General.use <- ds2$General.use[, drop = TRUE]
+
+summary.table <- tapply(ds2$Paper, list(ds2$General.use, ds2$Noptima), length)

### 2015-04-13 Manuscript/ohyou.tex
More eidts and moved out Bayesian stuff

@@
   Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
-  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
   Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
@@
 
-  Uses of the OU model can be divided into three broad categories. % improve citations in this section? % fix up numbers. Argggggggghhhhhhhhhhh.
-    \begin{enumerate}
-      \item \textbf{Modeling the evolution of a continuous character (N = 58).}\\ 
-            Generally in these analyses, multiple models of evolution are fit to the same continuous character \citep[e.g.,][]{Harmon:2010aa}.
-Model selection is then used to determine which model best fits the data.
-            If the OU model is favoured, this is often used as evidence of evolutionary constraints, stabilising selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}.
-      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character (N = 27).}\\
-            OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
-            The latter represents evolution under multiple selective regimes, and is probably more biologically realistic. 
-            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012modeling}.
-            Note that the first and second uses of the OU model are often combined, so that OU models with various numbers of optima are included in the pool of models being tested (N = 27).
-      \item \textbf{Controlling for phylogeny in correlative analyses (N = 35).}\\ 
+  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material). 
+  Multiple models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
+  OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
+  The latter represents evolution under multiple selective regimes, and may be more biologically realistic. 
+  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
+
+  OU models are also commonly used to control for phylogeny in correlative analyses (Tables \ref{table.uses}; Supplementary Material).  
   Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
   This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data.
-            In these analyses, the OU model is used to transform the tree and the variance-covariance matrix.
+  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
   This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
-    \end{enumerate}
+Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.
 
-  \noindent
-  The OU model is also used to reconstruct ancestral states \citep{martins1999estimation}
-  and to detect clade-wide convergent evolution \citep{ingram2013surface}. 
-  The majority of papers use the OU model to model the evolution of a continuous character (Fig 2 and Supplementary Material), % Need to add ref to supp mat
-  therefore, we focus on this use of the OU model in our simulations below. 
+  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material) therefore, we focus on this use of the OU model in our simulations below.
   The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures.
+  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate, but our general conclusions also apply to multiple optima OU models.
 
     \begin{table}[h]
-      \begin{centering}
-        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013.
-See Supplementary Material for details linkto supp mat.}
+        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supplementary Material for details.}
         \bigskip  
@@
     When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
-    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2). % Fix the figure or send to supplemental
+    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
 
-    %$\alpha$ can most easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+    $\alpha$ can most easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
 
@@
 
-    %If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, adaptation is quick, residual phylogenetic correlations are weak, and there is little influence of the past.
-This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term
+    This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
+    If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
 
@@
 
-     % This maybe belongs in the main text in shorter form  
-  %Figure \ref{figure.traitsim} gives some numerical examples, chosen to outline the behavior of $\alpha$ in relation to the above.
-We show the predicted covariance in traits \ref{equation:OUcov} scaled by the overall variance \ref{equation:OUvar} plotted against the time since divergence, $t_{ij}$ scaled by tree height.
-Figure \ref{figure.traitsim} highlights that the behavior of $\alpha$ can be classified into three regions.
-When $\alpha$ is low the model behaves very much like Brownian motion, with the covariance scaling approximately linearly with time.
-Only as the value becomes moderately larger (in this case approaching a value of $0.5$), does the non-Brownian behavior become apparent.
-Secondly, there is a region of moderate values in which the model leads to markedly non-Brownian patterns of trait variation (Fig.
-\ref{figure.traitsim}B).
-Finally, when $\alpha$ is large the model behaves differently again: in this case, all imprint of history is lost and the evolution is essentially a rapid burst at the present (Fig.
-\ref{figure.traitsim}C).
-This will be indistinguishable from phylogenetically unstructured data.
-
-\begin{figure}
-\centering
-    \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure2.pdf}
-\caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein-Uhlenbeck model.
-The covariance between species’ traits values is scaled by the intra-specific trait variance (i.e.
-equal to correlation between species’ traits).
-This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height).
-}
-\label{figure.traitsim}
-\end{figure}
-
-%In terms of statistical model fitting, a feature to highlight is that in Figure \ref{figure.traitsim}A and Figure \ref{figure.traitsim}C, changing the value of α across large relative ranges at the extremes (0 – 0.5 in Fig.
-\ref{figure.traitsim}A; 10-50 in Fig.
-\ref{figure.traitsim}C) produces relatively little quantitative change in the model behavior, and it is in the middle region of parameter space (1 – 5) that the biggest discernible effects occur.
-
-%The point of highlighting these aspects of expected behavior is that we believe these properties need to be understood if we are to build up a full picture of when the model is likely to be well estimated, as well as to understand the bounds of what can be inferred from real data.
+    \begin{enumerate}
+      \item $\alpha$ is bounded at zero.
+      \item blah
+      \item measurement errors
+    \end{enumerate}
 
@@
     We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
-    This procedure ... % explain WHY we are doing this, how does it relate to uses of the OU model.
+    This procedure mirrors a situation where the OU model  %%% argh
 
@@
 
-  \subsection{Bayesian simulations} % keep? or move it all to supp mat?
-    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
-    The choice of priors for OU models has not been fully explored. 
-    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
-    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
-    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
-S2-S5).
-% add links to this
-    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
-    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
-    Multiple independent chains were run for each analysis to ensure convergence was reached. 
-    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
-    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
-    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). 
-    We treated Bayes factors $>$ 2 as evidence favouring the OU model.
-
   \subsection{Other evolutionary models}
@@
 \section{Results}
-  \subsection{Simulations}
+  \subsection{Maximum Likelihood simulations}
   \label{section:sims.results}

### 2015-04-13 Manuscript/ohyou_suppmat.tex
Edits to intro and fixed equations
More eidts and moved out Bayesian stuff

@@
 
-    Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than methods papers with empirical proof of concept) published in the following journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
+    Next we filtered our Google Scholar search results to focus on empirical papers using OU models (rather than pure methods papers) published in the following journals: The American Naturalist, Ecology Letters, Evolution, Journal of Evolutionary Biology, Nature, Proceedings of the National Academy of Sciences, USA, Proceedings of the Royal Society B: Biological Sciences, and Science.
+    We only include papers up to the end of 2013 to ensure completeness.
+
     For each of these papers we recorded the number of species in the analysis, the study group (amphibians, birds, fish, mammals, reptiles, invertebrates or plants), the statistical package or specific R package used to fit the models, and the reason the authors state for using an OU model (ancestral state reconstructions, detecting convergent evolution, controlling for phylogeny, selecting a model of trait evolution, or other).
@@
     Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. 
-    We summarise these results in Fig. %figure link
-    and the full dataset is available in Supplemental Table S1 along with the full list of references.
+    We summarise these results in Figure 1 and Table 1, and the full dataset is available in Supplemental Table S1 % create
+    along with the full list of references.
 
@@
     Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig. S1 and Table S1). 
-    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the “best” model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
+    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the ``best'' model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
 S1).
+
+  \subsection{Bayesian simulations} % keep? or move it all to supp mat?
+    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
+    The choice of priors for OU models has not been fully explored. 
+    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
+    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
+    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
+S2-S5).
+% add links to this
+    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
+    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
+    Multiple independent chains were run for each analysis to ensure convergence was reached. 
+    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
+    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
+    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
+    We treated Bayes factors $>$ 2 as evidence favouring the OU model.

### 2015-04-13 Manuscript/ohyou.tex
Minor edits to methods

@@
 
+    \noindent
     where $W(t)$ is drawn at random from a normal distribution with mean $0$ and variance $\sigma^2$. 
@@
 
+    \noindent
     Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
@@
 
-    $\alpha$ can most easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+    $\alpha$ can more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
 
@@
 
+    \noindent
     This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
@@
   \subsection{Known statistical issues with the OU model}
-     % Problems with the OU model and where they are reported in the literature?
+  ???Problems with the OU model and where they are reported in the literature?
 

### 2015-04-13 Manuscript/ohyou.tex
Minor edits to methods

@@
     \noindent
-    This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
+    This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
     If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.

### 2015-04-14 Analyses/OUfigures.R
Added analyses from GHT

@@ -1,0 +1,572 @@
+###################################################
+#### Plot proportion error and alpha estimates ####
+###################################################
+setwd("/Users/ghthomas/Google Drive/CurrentResearch/OrnsteinUhlenbeckDiatribe")
+
+
+load("OU_simulations.rda")
+out <- data.frame(type=out[,1], index=as.numeric(out[,2]), bm=as.numeric(out[,3]), ou=as.numeric(out[,4]), LR=as.numeric(out[,5]), alpha=as.numeric(out[,6]), I.prime=as.numeric(out[,7]))
+
+
+t1err <- function(x) { return(x > 3.84) }
+HPD <- function(x) {qu <- quantile(x, probs=seq(0,1,0.025))
+    return(c(qu[2],qu[40]))
+}
+
+BF <- function (x) { return(x > 2) }
+
+
+xx1 <- grep("yule", out[,1])
+xx2 <- grep("err", out[,1])
+xx <- out[setdiff(xx1, xx2), ]
+yule <- aggregate(xx[,5:6], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+yule_HPD <- aggregate(xx[,6], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
+yuleerr <- rowSums(aggregate(xx[,5], by=list(xx[,1]), t1err)[c(5,6,1,3,4,7,2),2])
+
+xx <- out[grep("bdLow", out[,1]), ]
+bdlow <- aggregate(xx[,5:6], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+bdlow_HPD <- aggregate(xx[,6], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
+bdlowerr <- rowSums(aggregate(xx[,5], by=list(xx[,1]), t1err)[c(5,6,1,3,4,7,2),2])
+
+xx <- out[grep("bdMid", out[,1]), ]
+bdmid <- aggregate(xx[,5:6], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
[…]

### 2015-04-14 Analyses/OUsims.R

@@ -1,0 +1,2328 @@
+# Code for OU paper
+# Gavin Thomas 2014
+
+library(TESS)
+library(motmot)
+library(caper)
+
+# setwd to location of project
+
+out <- as.data.frame(matrix(NA, ncol=7, nrow=56000))
+
+for (i in 1:1000) {
+
+	
+	tr <- sim.globalBiDe.taxa(n=1,nTaxa=26,max=10,1,0,MRCA=TRUE)[[1]]
+    tipidx <- which(tr$edge.length==min(tr$edge.length[match(1:Ntip(tr), tr$edge[,2])]))
+    tip.d <- tr$tip.label[tr$edge[tipidx[1],2]]
+    tr <- drop.tip(tr, tip.d)
+    tr$tip.label <- paste("t", 1:Ntip(tr), sep="")
+	write.tree(tr, file="output/yule25.tre", append=TRUE)
+	dat <- transformPhylo.sim(tr, n=1, model="bm")
+	if (i==1) {write.table(t(dat[order(rownames(dat)),]), file="output/yule25.txt", append=FALSE, col.names=TRUE, row.names=FALSE)}
+	if (i!=1) {write.table(t(dat[order(rownames(dat)),]), file="output/yule25.txt", append=TRUE, col.names=FALSE, row.names=FALSE)}	
+	bm_mod <- transformPhylo.ML(dat, tr, model="bm", modelCIs=FALSE)
+	ou_mod <- transformPhylo.ML(dat, tr, model="OU", modelCIs=FALSE)	
+	sprich <- data.frame(species=tr$tip.label, rich=rep(1, Ntip(tr)))
+	cdat <- comparative.data(phy=tr, dat=sprich, names.col=species)
+	ft <- fusco.test(cdat, rich=rich, reps=1)
+	out[i,1] <- "yule25"
+	out[i,2] <- i
+	out[i,3] <- bm_mod$logLikelihood
+	out[i,4] <- ou_mod$MaximumLikelihood
+	out[i,5] <- 2*(ou_mod$MaximumLikelihood - bm_mod$logLikelihood)
+	out[i,6] <- ou_mod$Alpha
+	out[i,7] <- mean(ft[[1]][,"I.prime"])
+
+	tr <- sim.globalBiDe.taxa(n=1,nTaxa=51,max=10,1,0,MRCA=TRUE)[[1]]
[…]

### 2015-04-14 Analyses/PagelParamPlots.R

@@ -1,0 +1,255 @@
+setwd("/Users/ghthomas/Google Drive/CurrentResearch/OrnsteinUhlenbeckDiatribe")
+
+
+out <- read.csv("data/DeltaLambdaKappaMLRes2_Working.csv")
+
+#### Delta
+xx1 <- grep("yule", out[,1])
+xx2 <- grep("err", out[,1])
+xx <- out[setdiff(xx1, xx2), ]
+yule <- aggregate(xx[,c("Delta_LR", "Delta")], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+yule_HPD <- aggregate(xx[,c("Delta")], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
+yuleerr <- rowSums(aggregate(xx[,"Delta_LR"], by=list(xx[,1]), t1err)[c(5,6,1,3,4,7,2),2])
+
+xx <- out[grep("bdLow", out[,1]), ]
+bdlow <- aggregate(xx[,c("Delta_LR", "Delta")], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+bdlow_HPD <- aggregate(xx[,c("Delta")], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
+bdlowerr <- rowSums(aggregate(xx[,"Delta_LR"], by=list(xx[,1]), t1err)[c(5,6,1,3,4,7,2),2])
+
+xx <- out[grep("bdMid", out[,1]), ]
+bdmid <- aggregate(xx[,c("Delta_LR", "Delta")], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+bdmid_HPD <- aggregate(xx[,c("Delta")], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
+bdmiderr <- rowSums(aggregate(xx[,"Delta_LR"], by=list(xx[,1]), t1err)[c(5,6,1,3,4,7,2),2])
+
+xx <- out[grep("bdHigh", out[,1]), ]
+bdhigh <- aggregate(xx[,c("Delta_LR", "Delta")], by=list(xx[,1]), mean)[c(5,6,1,3,4,7,2),]
+bdhigh_HPD <- aggregate(xx[,c("Delta")], by=list(xx[,1]), HPD)[c(5,6,1,3,4,7,2),]
[…]

### 2015-04-14 R/OhYou_simulations.R
All simulateddate from GHT

@@ -1,4 +1,4 @@
-# Functions for simulating trees and data for the paper
+# Functions for simulating trees and data
 
 #----------------------------------
 # Simulating trees
@@ -161,43 +161,4 @@ get.tree.with.error <- function(phy, error, treetype, write.tree = FALSE) {
     write.trees(phy.error, paste(ntaxa, "_", error, sep = ""), treetype)
   }
   return(phy.error)
-}
-
-# Fine up to this point...
-all.trees.error <- function(ntaxa, error, write.tree = FALSE) {
-  yule <- yule.trees(ntaxa, write.tree = write.tree)
-  bdlow <- bdlow.trees(ntaxa, write.tree = write.tree)
-  bdmid <- bdmid.trees(ntaxa, write.tree = write.tree)
-  bdhigh <- bdhigh.trees(ntaxa, write.tree = write.tree)
-  return(list(yule, bdlow, bdmid, bdhigh))
-}
-
-bm.sim <- function(ntaxa, sigma, write.tree = FALSE, write.data = FALSE) {
-  phy <- all.trees(ntaxa, write.tree = write.tree)
-  yule <- bm.data(phy[[1]], sigma)
-  bdlow <- bm.data(phy[[2]], sigma)
-  bdmid <- bm.data(phy[[3]], sigma)
-  bdhigh <- bm.data(phy[[4]], sigma)
-  if (write.data == TRUE) {
-    write.data(yule, ntaxa, "yule", paste("_BM", sigma, sep = ""))
-    write.data(bdlow, ntaxa, "bdlow", paste("_BM", sigma, sep = ""))
-    write.data(bdmid, ntaxa, "bdmid", paste("_BM", sigma, sep = ""))
-    write.data(bdhigh, ntaxa, "bdhigh", paste("_BM", sigma, sep = ""))
-  }
-  return(list(phy, list(yule, bdlow, bdmid, bdhigh)))
-}
-
-ou.sim <- function(ntaxa, sigma, alpha, write.tree = FALSE, write.data = FALSE) {
-  phy <- all.trees(ntaxa, write.tree = write.tree)
-  yule <- ou.data(phy[[1]], sigma, alpha)
-  bdlow <- ou.data(phy[[2]], sigma, alpha)
-  bdmid <- ou.data(phy[[3]], sigma, alpha)
-  bdhigh <- ou.data(phy[[4]], sigma, alpha)
[…]

### 2015-04-14 R/OhYou_fittingmodels.R
Removed defunct fiule

@@ -1,4 +1,0 @@
-#------------------------------------------------------
-# Fitting BM and OU models to simulated trees and data
-# ML and Bayesian models if possible
-#------------------------------------------------------

### 2015-04-15 Manuscript/Tables/table_literature.tex
Minor edits

\begin{longtable}{p{6cm}llllll}
\hline

\textbf{Paper}	&	Year	&	Journal	&	Taxon	&	Ntips	&	Use in paper	&	Stats/R package	\\
\hline
(Hansen and Orzack 2005)	&	2005	&	Evolution	&	insects	&	15	&	other	&	OUCH precursor?	\\
(Edwards and Donoghue 2006) 	&	2006	&	Am Nat	&	plants	&	12	&	ancestral state reconstruction	&	COMPARE	\\
(Gvoždík and Damme 2006)	&	2006	&	Evolution	&	amphibians	&	10	&	phylogenetic correction	&	COMPARE	\\
(Halsey et al. 2006)	&	2006	&	Am Nat	&	birds/mammals	&	90	&	phylogenetic correction	&	Custom code	\\
(Ives and Godfray 2006)	&	2006	&	Am Nat	&	insects	&	8	&	phylogenetic signal	&	MATLAB	\\
(Valiente-Banuet et al. 2006)	&	2006	&	PNAS	&	plants	&	47	&	model of evolution	&	? (OUCH)	\\
(Clabaut et al. 2007)	&	2007	&	Evolution	&	fish	&	45	&	phylogenetic correction	&	APE	\\
(Gomez and Théry 2007)	&	2007	&	Am Nat	&	birds	&	40	&	model of evolution	&	OUCH	\\
(Hipp 2007)	&	2007	&	Evolution	&	plants	&	53	&	model of evolution	&	BayesTraits	\\
(Rezende et al. 2007)	&	2007	&	Nature	&	plants/insects	&	?	&	phylogenetic signal	&	? (OUCH)	\\
(Spoor et al. 2007)	&	2007	&	PNAS	&	mammals	&	210	&	phylogenetic correction	&	PDAP	\\
(Stuart‐Fox et al. 2007)	&	2007	&	Am Nat	&	reptiles	&	21	&	phylogenetic correction	&	COMPARE	\\
(Buchwalter et al. 2008)	&	2008	&	PNAS	&	insects	&	21	&	phylogenetic correction	&	MATLAB	\\
(Dumont and Payseur 2008)	&	2008	&	Evolution	&	mammals	&	13	&	model of evolution	&	OUCH	\\
(Hansen et al. 2008)	&	2008	&	Evolution	&	mammals	&	105	&	other	&	SLOUCH	\\
(Smith et al. 2008)	&	2008	&	Evolution	&	plants	&	15	&	phylogenetic correction	&	APE	\\
(Warne and Charnov 2008)	&	2008	&	Am Nat	&	reptiles	&	71	&	phylogenetic correction	&	MATLAB	\\
(Adams et al. 2009)	&	2009	&	PRSB	&	amphibians	&	10	&	model of evolution	&	OUCH	\\
(Addison et al. 2009)	&	2009	&	PRSB	&	birds	&	23	&	phylogenetic correction	&	MATLAB	\\
(Agrawal et al. 2009)	&	2009	&	PNAS	&	plants	&	53	&	model of evolution	&	GEIGER	\\
(Bergmann et al. 2009)	&	2009	&	Evolution	&	reptiles	&	38	&	model of evolution	&	OUCH	\\
(Collar et al. 2009)	&	2009	&	Evolution	&	fish	&	29	&	model of evolution	&	OUCH	\\
(Gonzalez-Voyer et al. 2009)	&	2009	&	PRSB	&	fish	&	39	&	phylogenetic correction	&	APE/COMPARE	\\
(Goodman et al. 2009)	&	2009	&	Evolution	&	reptiles	&	20	&	phylogenetic correction	&	COMPARE	\\
(Huey et al. 2009)	&	2009	&	PRSB	&	reptiles	&	70	&	phylogenetic correction	&	MATLAB	\\
(Kozak et al. 2009)	&	2009	&	Evolution	&	amphibians	&	184	&	model of evolution	&	GEIGER	\\
(Labra et al. 2009)	&	2009	&	Am Nat	&	reptiles	&	83	&	other	&	SLOUCH	\\
(Rezende et al. 2009)	&	2009	&	ELE	&	fish	&	116	&	model of evolution	&	PDTREE	\\
(Swanson and Garland 2009)	&	2009	&	Evolution	&	birds	&	44	&	phylogenetic correction	&	MATLAB	\\
(Van Buskirk 2009)	&	2009	&	JEB	&	amphibians	&	82	&	other	&	SLOUCH	\\
(Burbrink and Pyron 2010)	&	2010	&	Evolution	&	reptiles	&	29	&	model of evolution	&	GEIGER	\\
(Cooper and Purvis 2010)	&	2010	&	Am Nat	&	mammals	&	45	&	model of evolution	&	GEIGER	\\
(Edwards and Smith 2010)	&	2010	&	PNAS	&	plants	&	300	&	model of evolution	&	OUCH	\\
(Harmon et al. 2010)	&	2010	&	Evolution	&	multiple	&	17	&	model of evolution	&	GEIGER	\\
(Helmus et al. 2010)	&	2010	&	ELE	&	zooplankton	&	15	&	phylogenetic correction	&	?	\\
(Kalinka et al. 2010)	&	2010	&	Nature	&	insects	&	6	&	model of evolution	&	OUCH	\\
(Kozak and Wiens 2010b)	&	2010	&	Am Nat	&	amphibians	&	84	&	model of evolution/	&	GEIGER/	\\
	&		&		&		&		&	ancestral state reconstruction	&	OUCH	\\
(Kozak and Wiens 2010a)	&	2010	&	ELE	&	amphibians	&	11	&	model of evolution	&	OUCH	\\
(Ord et al. 2010)	&	2010	&	Evolution	&	reptiles	&	16	&	other	&	SLOUCH	\\
(Price et al. 2010)	&	2010	&	Evolution	&	fish	&	122	&	model of evolution	&	GEIGER	\\
(Slater et al. 2010)	&	2010	&	PRSB	&	mammals	&	84	&	model of evolution	&	OUCH	\\
(Angielczyk et al. 2011)	&	2011	&	Evolution	&	reptiles	&	8	&	model of evolution	&	OUCH	\\
(Benesh et al. 2011)	&	2011	&	Evolution	&	helminths	&	310	&	model of evolution	&	OUCH	\\
(Collar et al. 2011)	&	2011	&	Evolution	&	reptiles	&	37	&	model of evolution	&	Brownie	\\
(Derryberry et al. 2011)	&	2011	&	Evolution	&	birds	&	285	&	model of evolution	&	GEIGER	\\
(Galvan and Moller 2011)	&	2011	&	JEB	&	birds	&	323	&	phylogenetic correction	&	COMPARE	\\
(Gonzalez-Voyer and Kolm 2011)	&	2011	&	JEB	&	fish	&	49	&	model of evolution	&	GEIGER	\\
(Ord et al. 2011)	&	2011	&	Evolution	&	multiple	&	23	&	phylogenetic signal	&	SLOUCH	\\
(Oufiero et al. 2011) 	&	2011	&	Evolution	&	reptiles	&	106	&	phylogenetic correction	&	SLOUCH	\\
(Perez et al. 2011)	&	2011	&	JEB	&	mammals	&	29	&	phylogenetic correction	&	APE	\\
(Raia and Meiri 2011)	&	2011	&	Evolution	&	mammals	&	842	&	model of evolution	&	MOTMOT	\\
(Rosas-Guerrero et al. 2011)	&	2011	&	Evolution	&	plants	&	20	&	phylogenetic correction	&	APE	\\
(Setiadi et al. 2011)	&	2011	&	Am Nat	&	amphibians	&	22	&	model of evolution	&	OUCH	\\
(Smith et al. 2011)	&	2011	&	Evolution	&	reptiles	&	15	&	model of evolution	&	GEIGER	\\
(Tulli et al. 2011)	&	2011	&	JEB	&	reptiles	&	29	&	phylogenetic correction	&	?	\\
(Turbill et al. 2011)	&	2011	&	PRSB	&	mammals	&	19	&	phylogenetic correction	&	GEIGER	\\
(Valido et al. 2011)	&	2011	&	JEB	&	plants	&	111	&	phylogenetic correction	&	APE	\\
(Wiens et al. 2011)	&	2011	&	ELE	&	amphibians	&	337	&	model of evolution/	&	GEIGER/	\\
	&		&		&		&		&	phylogenetic correction	&	COMPARE	\\
(Weir and Wheatcroft 2011)	&	2011	&	PRSB	&	birds	&	232	&	model of evolution	&	GEIGER	\\
(Beaulieu et al. 2012)	&	2012	&	Evolution	&	plants	&	590	&	model of evolution	&	OUwie	\\
(Betancur-R et al. 2012)	&	2012	&	ELE	&	fish	&	123	&	model of evolution	&	GEIGER	\\
(Blankers et al. 2012)	&	2012	&	JEB	&	amphibians	&	189	&	phylogenetic correction	&	GEIGER	\\
(Boettiger et al. 2012)	&	2012	&	Evolution	&	reptiles	&	23	&	ancestral state reconstruction	&	OUCH	\\
(Burbrink et al. 2012)	&	2012	&	PRSB	&	multiple	&	41	&	model of evolution	&	Custom code	\\
(Calosi et al. 2012)	&	2012	&	JEB	&	insects	&	25	&	phylogenetic correction	&	MATLAB	\\
(Claramunt et al. 2012a)	&	2012	&	Am Nat	&	birds	&	290	&	model of evolution	&	GEIGER	\\
(Claramunt et al. 2012b)	&	2012	&	PRSB	&	birds	&	282	&	model of evolution	&	GEIGER	\\
(Davis et al. 2012)	&	2012	&	JEB	&	insects	&	53	&	phylogenetic correction	&	SLOUCH	\\
(Diniz-Filho et al. 2012)	&	2012	&	Evolution	&	mammals	&	209	&	other	&	PAM	\\
(Fusco et al. 2012)	&	2012	&	Evolution	&	trilobites	&	60	&	model of evolution	&	?	\\
(Gomez-Mestre et al. 2012)	&	2012	&	Evolution	&	amphibians	&	720	&	phylogenetic correction	&	APE	\\
(Ingram et al. 2012)	&	2012	&	JEB	&	food webs	&	20	&	model of evolution	&	GEIGER	\\
(Kellermann et al. 2012a)	&	2012	&	Evolution	&	insects	&	94	&	phylogenetic correction	&	SLOUCH	\\
(Kellermann et al. 2012b)	&	2012	&	PNAS	&	insects	&	94	&	phylogenetic correction	&	SLOUCH	\\
(Nogueira et al. 2012)	&	2012	&	JEB	&	plants	&	105	&	phylogenetic signal	&	GEIGER	\\
(Ord 2012)	&	2012	&	JEB	&	reptiles	&	32	&	ancestral state reconstruction	&	SLOUCH	\\
(Pearse and Hipp 2012)	&	2012	&	Evolution	&	plants	&	56	&	phylogenetic correction	&	SLOUCH	\\
(Pellissier et al. 2012)	&	2012	&	JEB	&	insects	&	83	&	model of evolution	&	? (APE)	\\
(Price et al. 2012)	&	2012	&	Evolution	&	fish	&	50	&	model of evolution	&	GEIGER	\\
(Sallan and Friedman 2012)	&	2012	&	PRSB	&	fish	&	100	&	model of evolution	&	GEIGER	\\
(Santana et al. 2012)	&	2012	&	Evolution	&	mammals	&	85	&	model of evolution	&	OUCH	\\
(Schmerler et al. 2012)	&	2012	&	PRSB	&	plants	&	88	&	phylogenetic correction	&	nlme	\\
(Smith 2012)	&	2012	&	Evolution	&	birds	&	42	&	phylogenetic correction	&	APE/nlme	\\
(Sookias et al. 2012)	&	2012	&	PRSB	&	multiple	&	43	&	model of evolution	&	GEIGER	\\
(Stireman et al. 2012)	&	2012	&	JEB	&	insects	&	24	&	model of evolution	&	GEIGER	\\
(Voje and Hansen 2012)	&	2012	&	Evolution	&	insects	&	30	&	phylogenetic correction	&	SLOUCH	\\
(Weir et al. 2012)	&	2012	&	Evolution	&	birds	&	232	&	model of evolution	&	GEIGER	\\
(Arbour and López-Fernández 2013)	&	2013	&	PRSB	&	fish	&	27	&	model of evolution	&	OUCH	\\
(Benesh et al. 2013)	&	2013	&	Am Nat	&	helminths	&	143	&	phylogenetic correction	&	APE	\\
(Blackburn et al. 2013)	&	2013	&	Evolution	&	amphibians	&	18	&	model of evolution	&	GEIGER	\\
(Christin et al. 2013)	&	2013	&	PNAS	&	plants	&	545	&	model of evolution	&	GEIGER/OUCH	\\
(Frédérich et al. 2013)	&	2013	&	Am Nat	&	fish	&	208	&	model of evolution	&	OUwie	\\
(Friedman et al. 2013)	&	2013	&	Evolution	&	birds	&	15	&	model of evolution	&	OUCH	\\
(Guerrero et al. 2013)	&	2013	&	PNAS	&	plants/reptiles	&	49	&	ancestral state reconstruction	&	GEIGER/	\\
	&		&		&		&		&		&	COMPARE	\\
(Hertz et al. 2013)	&	2013	&	Evolution	&	reptiles	&	100	&	model of evolution	&	GEIGER	\\
(Hossie et al. 2013)	&	2013	&	JEB	&	amphibians/	&	104	&	phylogenetic correction	&	GEIGER	\\
	&		&		&	reptiles	&		&		&		\\
(Knope and Scales 2013)	&	2013	&	JEB	&	fish	&	26	&	model of evolution	&	OUCH	\\
(Kostikova et al. 2013)	&	2013	&	Am Nat	&	plants	&	68	&	model of evolution	&	OUwie	\\
(Lambert and Wiens 2013)	&	2013	&	Evolution	&	reptiles	&	117	&	ancestral state reconstruction	&	GEIGER	\\
(Lapiedra et al. 2013)	&	2013	&	PRSB	&	birds	&	154	&	model of evolution	&	OUwie	\\
(Litsios et al. 2013)	&	2013	&	Evolution	&	plants	&	382	&	model of evolution	&	OUwie	\\
(López-Fernández et al. 2013)	&	2013	&	Evolution	&	fish	&	135	&	model of evolution	&	GEIGER	\\
(Machac et al. 2013)	&	2013	&	Evolution	&	mammals	&	231	&	model of evolution	&	GEIGER	\\
(Mahler et al. 2013)	&	2013	&	Science	&	reptiles	&	100	&	convergent evolution	&	SURFACE	\\
(Maia et al. 2013)	&	2013	&	PNAS	&	birds	&	47	&	model of evolution	&	OUwie	\\
(Mirceta et al. 2013)	&	2013	&	Science	&	mammals	&	130	&	phylogenetic correction/	&	Mesquite/	\\
	&		&		&		&		&	ancestral state reconstruction	&	MATLAB	\\
(Moen et al. 2013)	&	2013	&	PRSB	&	amphibians	&	44	&	convergent evolution	&	GEIGER	\\
(Pérez i de Lanuza et al. 2013)	&	2013	&	JEB	&	reptiles	&	42	&	ancestral state reconstruction	&	GEIGER	\\
(Pienaar et al. 2013)	&	2013	&	ELE	&	birds	&	382	&	model of evolution	&	SLOUCH	\\
(Quintero and Wiens 2013)	&	2013	&	ELE	&	multiple	&	500	&	ancestral state reconstruction	&	GEIGER/COMPARE	\\
(Ryan and Shaw 2013)	&	2013	&	PRSB	&	mammals	&	34	&	phylogenetic correction	&	MATLAB	\\
(Seddon et al. 2013) 	&	2013	&	PRSB	&	birds	&	153	&	model of evolution	&	? (GEIGER)	\\
(Tanabe and Sota 2013)	&	2013	&	Evolution	&	millipedes	&	84	&	phylogenetic correction	&	APE	\\
(Voje et al. 2013)	&	2013	&	JEB	&	fish	&	87	&	other	&	SLOUCH	\\
(Wiens et al. 2013)	&	2013	&	Evolution	&	reptiles	&	117	&	ancestral state reconstruction	&	GEIGER	\\
\hline
\caption[Literature review]
			{Details of the papers used in our literature review. For each paper we recorded the study taxon, the number of tips in the phylogeny used to fit the Ornstein Uhlenbeck (OU) model, how the authors used the OU model in the paper, and the statistical package (usually R) used to carry out the analyses.For a full reference list see below.}
\label{tab:sk.measurements}

\end{longtable}

### 2015-04-15 Manuscript/ohyou_suppmat.tex
Minor edits

@@
 
-  \subsection{Literature Review}
+  \subsection{Literature review}
     To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''.
@@
     Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. 
-    We summarise these results in Figure 1 and Table 1, and the full dataset is available in Supplemental Table S1 % create
-    along with the full list of references.
+    We summarise these results in Figure 1 and Table 1, and the full dataset is available below along with the full list of references.
 
-    In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. \ref{figure:literature}). 
+    In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. 1). 
     Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig. S1 and Table S1). 
@@
 S1).
+
+\begin{landscape}
+\begin{center}
+\begin{longtable}{p{6cm}llllll}
+\caption[Literature review]
+        {Details of the papers used in our literature review.
+For each paper we recorded the study taxon, the number of tips in the phylogeny used to fit the Ornstein Uhlenbeck (OU) model, how the authors used the OU model in the paper, and the statistical package (usually R) used to carry out the analyses.
+For a full reference list see below.
+Journal abbreviations: Am Nat = The American Naturalist, ELE = Ecology Letters, JEB = Journal of Evolutionary Biology, PRSB = Proceedings of the Royal Society B: Biological Sciences, PNAS = Proceedings of the National Academy of the USA.
+†Number of tips in the phylogeny the OU model was fitted to.
+Where there were multiple analyses in a paper we use the median number of tips.
+All packages mentioned are R packages apart from BayesTraits, COMPARE, MATLAB, Mesquite, PAM, PDAP and PDTREE.}\\
+\hline
+
+\textbf{Paper}  &   Year    &   Journal &   Taxon   &   Ntips   &   Use in paper    &   Stats/R package \\
+\hline
+Hansen and Orzack 2005    &   2005    &   Evolution   &   insects &   15  &   other   &   OUCH precursor? \\
+Edwards and Donoghue 2006     &   2006    &   Am Nat  &   plants  &   12  &   ancestral state reconstruction  &   COMPARE \\
+Gvoždík and Damme 2006    &   2006    &   Evolution   &   amphibians  &   10  &   phylogenetic correction &   COMPARE \\
+Halsey et al. 2006    &   2006    &   Am Nat  &   birds/mammals   &   90  &   phylogenetic correction &   Custom code \\
+Ives and Godfray 2006 &   2006    &   Am Nat  &   insects &   8   &   phylogenetic signal &   MATLAB  \\
+Valiente-Banuet et al. 2006   &   2006    &   PNAS    &   plants  &   47  &   model of evolution  &   ? OUCH    \\
+Clabaut et al. 2007   &   2007    &   Evolution   &   fish    &   45  &   phylogenetic correction &   APE \\
+Gomez and Théry 2007  &   2007    &   Am Nat  &   birds   &   40  &   model of evolution  &   OUCH    \\
+Hipp 2007 &   2007    &   Evolution   &   plants  &   53  &   model of evolution  &   BayesTraits \\
+Rezende et al. 2007   &   2007    &   Nature  &   plants/insects  &   ?   &   phylogenetic signal &   ? OUCH    \\
+Spoor et al. 2007 &   2007    &   PNAS    &   mammals &   210 &   phylogenetic correction &   PDAP    \\
+Stuart‐Fox et al. 2007    &   2007    &   Am Nat  &   reptiles    &   21  &   phylogenetic correction &   COMPARE \\
+Buchwalter et al. 2008    &   2008    &   PNAS    &   insects &   21  &   phylogenetic correction &   MATLAB  \\
+Dumont and Payseur 2008   &   2008    &   Evolution   &   mammals &   13  &   model of evolution  &   OUCH    \\
+Hansen et al. 2008    &   2008    &   Evolution   &   mammals &   105 &   other   &   SLOUCH  \\
+Smith et al. 2008 &   2008    &   Evolution   &   plants  &   15  &   phylogenetic correction &   APE \\
+Warne and Charnov 2008    &   2008    &   Am Nat  &   reptiles    &   71  &   phylogenetic correction &   MATLAB  \\
+Adams et al. 2009 &   2009    &   PRSB    &   amphibians  &   10  &   model of evolution  &   OUCH    \\
+Addison et al. 2009   &   2009    &   PRSB    &   birds   &   23  &   phylogenetic correction &   MATLAB  \\
+Agrawal et al. 2009   &   2009    &   PNAS    &   plants  &   53  &   model of evolution  &   GEIGER  \\
+Bergmann et al. 2009  &   2009    &   Evolution   &   reptiles    &   38  &   model of evolution  &   OUCH    \\
+Collar et al. 2009    &   2009    &   Evolution   &   fish    &   29  &   model of evolution  &   OUCH    \\
+Gonzalez-Voyer et al. 2009    &   2009    &   PRSB    &   fish    &   39  &   phylogenetic correction &   APE/COMPARE \\
+Goodman et al. 2009   &   2009    &   Evolution   &   reptiles    &   20  &   phylogenetic correction &   COMPARE \\
+Huey et al. 2009  &   2009    &   PRSB    &   reptiles    &   70  &   phylogenetic correction &   MATLAB  \\
+Kozak et al. 2009 &   2009    &   Evolution   &   amphibians  &   184 &   model of evolution  &   GEIGER  \\
+Labra et al. 2009 &   2009    &   Am Nat  &   reptiles    &   83  &   other   &   SLOUCH  \\
+Rezende et al. 2009   &   2009    &   ELE &   fish    &   116 &   model of evolution  &   PDTREE  \\
+Swanson and Garland 2009  &   2009    &   Evolution   &   birds   &   44  &   phylogenetic correction &   MATLAB  \\
+Van Buskirk 2009  &   2009    &   JEB &   amphibians  &   82  &   other   &   SLOUCH  \\
+Burbrink and Pyron 2010   &   2010    &   Evolution   &   reptiles    &   29  &   model of evolution  &   GEIGER  \\
+Cooper and Purvis 2010    &   2010    &   Am Nat  &   mammals &   45  &   model of evolution  &   GEIGER  \\
+Edwards and Smith 2010    &   2010    &   PNAS    &   plants  &   300 &   model of evolution  &   OUCH    \\
+Harmon et al. 2010    &   2010    &   Evolution   &   multiple    &   17  &   model of evolution  &   GEIGER  \\
+Helmus et al. 2010    &   2010    &   ELE &   zooplankton &   15  &   phylogenetic correction &   ?   \\
+Kalinka et al. 2010   &   2010    &   Nature  &   insects &   6   &   model of evolution  &   OUCH    \\
+Kozak and Wiens 2010b &   2010    &   Am Nat  &   amphibians  &   84  &   model of evolution/ &   GEIGER/ \\
+    &       &       &       &       &   ancestral state reconstruction  &   OUCH    \\
+Kozak and Wiens 2010a &   2010    &   ELE &   amphibians  &   11  &   model of evolution  &   OUCH    \\
+Ord et al. 2010   &   2010    &   Evolution   &   reptiles    &   16  &   other   &   SLOUCH  \\
+Price et al. 2010 &   2010    &   Evolution   &   fish    &   122 &   model of evolution  &   GEIGER  \\
+Slater et al. 2010    &   2010    &   PRSB    &   mammals &   84  &   model of evolution  &   OUCH    \\
+Angielczyk et al. 2011    &   2011    &   Evolution   &   reptiles    &   8   &   model of evolution  &   OUCH    \\
+Benesh et al. 2011    &   2011    &   Evolution   &   helminths   &   310 &   model of evolution  &   OUCH    \\
+Collar et al. 2011    &   2011    &   Evolution   &   reptiles    &   37  &   model of evolution  &   Brownie \\
+Derryberry et al. 2011    &   2011    &   Evolution   &   birds   &   285 &   model of evolution  &   GEIGER  \\
+Galvan and Moller 2011    &   2011    &   JEB &   birds   &   323 &   phylogenetic correction &   COMPARE \\
+Gonzalez-Voyer and Kolm 2011  &   2011    &   JEB &   fish    &   49  &   model of evolution  &   GEIGER  \\
+Ord et al. 2011   &   2011    &   Evolution   &   multiple    &   23  &   phylogenetic signal &   SLOUCH  \\
+Oufiero et al. 2011   &   2011    &   Evolution   &   reptiles    &   106 &   phylogenetic correction &   SLOUCH  \\
+Perez et al. 2011 &   2011    &   JEB &   mammals &   29  &   phylogenetic correction &   APE \\
+Raia and Meiri 2011   &   2011    &   Evolution   &   mammals &   842 &   model of evolution  &   MOTMOT  \\
+Rosas-Guerrero et al. 2011    &   2011    &   Evolution   &   plants  &   20  &   phylogenetic correction &   APE \\
+Setiadi et al. 2011   &   2011    &   Am Nat  &   amphibians  &   22  &   model of evolution  &   OUCH    \\
+Smith et al. 2011 &   2011    &   Evolution   &   reptiles    &   15  &   model of evolution  &   GEIGER  \\
+Tulli et al. 2011 &   2011    &   JEB &   reptiles    &   29  &   phylogenetic correction &   ?   \\
+Turbill et al. 2011   &   2011    &   PRSB    &   mammals &   19  &   phylogenetic correction &   GEIGER  \\
+Valido et al. 2011    &   2011    &   JEB &   plants  &   111 &   phylogenetic correction &   APE \\
+Wiens et al. 2011 &   2011    &   ELE &   amphibians  &   337 &   model of evolution/ &   GEIGER/ \\
+    &       &       &       &       &   phylogenetic correction &   COMPARE \\
+Weir and Wheatcroft 2011  &   2011    &   PRSB    &   birds   &   232 &   model of evolution  &   GEIGER  \\
+Beaulieu et al. 2012  &   2012    &   Evolution   &   plants  &   590 &   model of evolution  &   OUwie   \\
+Betancur-R et al. 2012    &   2012    &   ELE &   fish    &   123 &   model of evolution  &   GEIGER  \\
+Blankers et al. 2012  &   2012    &   JEB &   amphibians  &   189 &   phylogenetic correction &   GEIGER  \\
+Boettiger et al. 2012 &   2012    &   Evolution   &   reptiles    &   23  &   ancestral state reconstruction  &   OUCH    \\
+Burbrink et al. 2012  &   2012    &   PRSB    &   multiple    &   41  &   model of evolution  &   Custom code \\
+Calosi et al. 2012    &   2012    &   JEB &   insects &   25  &   phylogenetic correction &   MATLAB  \\
+Claramunt et al. 2012a    &   2012    &   Am Nat  &   birds   &   290 &   model of evolution  &   GEIGER  \\
+Claramunt et al. 2012b    &   2012    &   PRSB    &   birds   &   282 &   model of evolution  &   GEIGER  \\
+Davis et al. 2012 &   2012    &   JEB &   insects &   53  &   phylogenetic correction &   SLOUCH  \\
+Diniz-Filho et al. 2012   &   2012    &   Evolution   &   mammals &   209 &   other   &   PAM \\
+Fusco et al. 2012 &   2012    &   Evolution   &   trilobites  &   60  &   model of evolution  &   ?   \\
+Gomez-Mestre et al. 2012  &   2012    &   Evolution   &   amphibians  &   720 &   phylogenetic correction &   APE \\
+Ingram et al. 2012    &   2012    &   JEB &   food webs   &   20  &   model of evolution  &   GEIGER  \\
+Kellermann et al. 2012a   &   2012    &   Evolution   &   insects &   94  &   phylogenetic correction &   SLOUCH  \\
+Kellermann et al. 2012b   &   2012    &   PNAS    &   insects &   94  &   phylogenetic correction &   SLOUCH  \\
+Nogueira et al. 2012  &   2012    &   JEB &   plants  &   105 &   phylogenetic signal &   GEIGER  \\
+Ord 2012  &   2012    &   JEB &   reptiles    &   32  &   ancestral state reconstruction  &   SLOUCH  \\
+Pearse and Hipp 2012  &   2012    &   Evolution   &   plants  &   56  &   phylogenetic correction &   SLOUCH  \\
+Pellissier et al. 2012    &   2012    &   JEB &   insects &   83  &   model of evolution  &   ? APE \\
+Price et al. 2012 &   2012    &   Evolution   &   fish    &   50  &   model of evolution  &   GEIGER  \\
+Sallan and Friedman 2012  &   2012    &   PRSB    &   fish    &   100 &   model of evolution  &   GEIGER  \\
+Santana et al. 2012   &   2012    &   Evolution   &   mammals &   85  &   model of evolution  &   OUCH    \\
+Schmerler et al. 2012 &   2012    &   PRSB    &   plants  &   88  &   phylogenetic correction &   nlme    \\
+Smith 2012    &   2012    &   Evolution   &   birds   &   42  &   phylogenetic correction &   APE/nlme    \\
+Sookias et al. 2012   &   2012    &   PRSB    &   multiple    &   43  &   model of evolution  &   GEIGER  \\
+Stireman et al. 2012  &   2012    &   JEB &   insects &   24  &   model of evolution  &   GEIGER  \\
+Voje and Hansen 2012  &   2012    &   Evolution   &   insects &   30  &   phylogenetic correction &   SLOUCH  \\
+Weir et al. 2012  &   2012    &   Evolution   &   birds   &   232 &   model of evolution  &   GEIGER  \\
+Arbour and López-Fernández 2013   &   2013    &   PRSB    &   fish    &   27  &   model of evolution  &   OUCH    \\
+Benesh et al. 2013    &   2013    &   Am Nat  &   helminths   &   143 &   phylogenetic correction &   APE \\
+Blackburn et al. 2013 &   2013    &   Evolution   &   amphibians  &   18  &   model of evolution  &   GEIGER  \\
+Christin et al. 2013  &   2013    &   PNAS    &   plants  &   545 &   model of evolution  &   GEIGER/OUCH \\
+Frédérich et al. 2013 &   2013    &   Am Nat  &   fish    &   208 &   model of evolution  &   OUwie   \\
+Friedman et al. 2013  &   2013    &   Evolution   &   birds   &   15  &   model of evolution  &   OUCH    \\
+Guerrero et al. 2013  &   2013    &   PNAS    &   plants/reptiles &   49  &   ancestral state reconstruction  &   GEIGER/ \\
+    &       &       &       &       &       &   COMPARE \\
+Hertz et al. 2013 &   2013    &   Evolution   &   reptiles    &   100 &   model of evolution  &   GEIGER  \\
+Hossie et al. 2013    &   2013    &   JEB &   amphibians/ &   104 &   phylogenetic correction &   GEIGER  \\
+    &       &       &   reptiles    &       &       &       \\
+Knope and Scales 2013 &   2013    &   JEB &   fish    &   26  &   model of evolution  &   OUCH    \\
+Kostikova et al. 2013 &   2013    &   Am Nat  &   plants  &   68  &   model of evolution  &   OUwie   \\
+Lambert and Wiens 2013    &   2013    &   Evolution   &   reptiles    &   117 &   ancestral state reconstruction  &   GEIGER  \\
+Lapiedra et al. 2013  &   2013    &   PRSB    &   birds   &   154 &   model of evolution  &   OUwie   \\
+Litsios et al. 2013   &   2013    &   Evolution   &   plants  &   382 &   model of evolution  &   OUwie   \\
+López-Fernández et al. 2013   &   2013    &   Evolution   &   fish    &   135 &   model of evolution  &   GEIGER  \\
+Machac et al. 2013    &   2013    &   Evolution   &   mammals &   231 &   model of evolution  &   GEIGER  \\
+Mahler et al. 2013    &   2013    &   Science &   reptiles    &   100 &   convergent evolution    &   SURFACE \\
+Maia et al. 2013  &   2013    &   PNAS    &   birds   &   47  &   model of evolution  &   OUwie   \\
+Mirceta et al. 2013   &   2013    &   Science &   mammals &   130 &   phylogenetic correction/    &   Mesquite/   \\
+    &       &       &       &       &   ancestral state reconstruction  &   MATLAB  \\
+Moen et al. 2013  &   2013    &   PRSB    &   amphibians  &   44  &   convergent evolution    &   GEIGER  \\
+Pérez i de Lanuza et al. 2013 &   2013    &   JEB &   reptiles    &   42  &   ancestral state reconstruction  &   GEIGER  \\
+Pienaar et al. 2013   &   2013    &   ELE &   birds   &   382 &   model of evolution  &   SLOUCH  \\
+Quintero and Wiens 2013   &   2013    &   ELE &   multiple    &   500 &   ancestral state reconstruction  &   GEIGER/COMPARE  \\
+Ryan and Shaw 2013    &   2013    &   PRSB    &   mammals &   34  &   phylogenetic correction &   MATLAB  \\
+Seddon et al. 2013    &   2013    &   PRSB    &   birds   &   153 &   model of evolution  &   ? GEIGER  \\
+Tanabe and Sota 2013  &   2013    &   Evolution   &   millipedes  &   84  &   phylogenetic correction &   APE \\
+Voje et al. 2013  &   2013    &   JEB &   fish    &   87  &   other   &   SLOUCH  \\
+Wiens et al. 2013 &   2013    &   Evolution   &   reptiles    &   117 &   ancestral state reconstruction  &   GEIGER  \\
+\hline
+\label{table:litreview}
+\end{longtable}
+\end{center}
+\end{landscape}
+
+\subsection{Literature review references}
+
+\indent
+Adams, D.
+C., C.
+M.
+Berns, K.
+H.
+Kozak, and J.
+J.
+Wiens.
+2009.
+Are rates of species diversification correlated with rates of morphological evolution?
+Proceedings of the Royal Society B: Biological Sciences 276:2729-2738.\\
+Addison, B., K.
+C.
+Klasing, W.
+D.
+Robinson, S.
+H.
+Austin, and R.
+E.
+Ricklefs.
+2009.
+Ecological and life-history factors influencing the evolution of maternal antibody allocation: a phylogenetic comparison.
+Proceedings of the Royal Society B: Biological Sciences 276:3979-3987.\\
+Agrawal, A.
+A., M.
+Fishbein, R.
+Halitschke, A.
+P.
+Hastings, D.
+L.
+Rabosky, and S.
+Rasmann.
+2009.
+Evidence for adaptive radiation from a phylogenetic study of plant defenses.
+Proceedings of the National Academy of Sciences 106:18067-18072.\\
+Angielczyk, K. D., C. R. Feldman, and G. R. Miller. 2011. Adaptive evolution of plastron shape in emydine turtles. Evolution 65:377-394.\\
+Arbour, J.
+H., and H.
+López-Fernández.
+2013.
+Ecological variation in South American geophagine cichlids arose during an early burst of adaptive morphological and functional evolution.
+Proceedings of the Royal Society B: Biological Sciences 280.\\
+Beaulieu, J. M., D.-C. Jhwueng, C. Boettiger, and B. C. O’Meara. 2012. Modeling stabilizing selection: expanding the Ornstein-Uhlenbeck model of adaptive evolution. Evolution 66:2369-2383.\\
+Benesh, D.
+P., J.
+C.
+Chubb, and G.
+A.
+Parker.
+2011.
+Exploitation of the same trophic link favors convergence of larval life-history strategies in complex life cycle helminths.
+Evolution 65:2286-2299.\\
+Benesh, D. P., J. C. Chubb, and G. A. Parker. 2013. Complex life cycles: why refrain from growth before reproduction in the adult niche? The American Naturalist 181:39-51.\\
+Bergmann, P. J., J. J. Meyers, and D. J. Irschick. 2009. Directional evolution of stockiness coevolves with ecology and locomotion in lizards. Evolution 63:215-227.\\
+Betancur-R, R., G.
+Ortí, A.
+M.
+Stein, A.
+P.
+Marceniuk, and R.
+Alexander Pyron.
+2012.
+Apparent signal of competition limiting diversification after ecological transitions from marine to freshwater habitats.
+Ecology Letters 15:822-830.\\
+Blackburn, D. C., C. D. Siler, A. C. Diesmos, J. A. McGuire, D. C. Cannatella, and R. M. Brown. 2013. An adaptive radiation of frogs in a Southeat Asian island archipelago. Evolution 67:2631–2646.\\
+Blankers, T., D. C. Adams, and J. J. Wiens. 2012. Ecological radiation with limited morphological diversification in salamanders. Journal of evolutionary biology 25:634-646.\\
+Boettiger, C., G. Coop, and P. Ralph. 2012. Is your phylogeny informative? Measuring the power of comparative methods. Evolution 66:2240-2251.\\
+Buchwalter, D.
+B., D.
+J.
+Cain, C.
+A.
+Martin, L.
+Xie, S.
+N.
+Luoma, and T.
+Garland.
+2008.
+Aquatic insect ecophysiological traits reveal phylogenetically based differences in dissolved cadmium susceptibility.
+Proceedings of the National Academy of Sciences 105:8321-8326.\\
+Burbrink, F.
+T., X.
+Chen, E.
+A.
+Myers, M.
+C.
+Brandley, and R.
+A.
+Pyron.
+2012.
+Evidence for determinism in species diversification and contingency in phenotypic evolution during adaptive radiation.
+Proceedings of the Royal Society B: Biological Sciences 279:4817-4826.\\
+Burbrink, F.
+T., and R.
+A.
+Pyron.
+2010.
+How does ecological opportunity influence rates of speciation, extinction, and morphological diversification in New World ratsnakes (tribe Lampropeltini)?
+Evolution 64:934-943.\\
+Calosi, P., D.
+T.
+Bilton, J.
+I.
+Spicer, W.
+C.
+E.
+P.
+Verberk, A.
+Atfield, and T.
+Garland.
+2012.
+The comparative biology of diving in two genera of European Dytiscidae (Coleoptera).
+Journal of Evolutionary Biology 25:329-341.\\
+Christin, P.-A., C.
+P.
+Osborne, D.
+S.
+Chatelet, J.
+T.
+Columbus, G.
+Besnard, T.
+R.
+Hodkinson, L.
+M.
+Garrison, M.
+S.
+Vorontsova, and E.
+J.
+Edwards.
+2013.
+Anatomical enablers and the evolution of C4 photosynthesis in grasses.
+Proceedings of the National Academy of Sciences 110:1381-1386.\\
+Clabaut, C., P.
+M.
+E.
+Bunje, W.
+Salzburger, and A.
+Meyer.
+2007.
+Geometric morphometric analyses provide evidence for the adaptive character of the Tanganyikan cichlid fish radiations.
+Evolution 61:560-578.\\
+Claramunt, S., E.
+P.
+Derryberry, R.
+T.
+Brumfield, and J.
+V.
+Remsen Jr.
+2012a.
+Ecological opportunity and diversification in a continental radiation of birds: climbing adaptations and cladogenesis in the Furnariidae.
+The American Naturalist 179:649-666.\\
+Claramunt, S., E.
+P.
+Derryberry, J.
+V.
+Remsen, and R.
+T.
+Brumfield.
+2012b.
+High dispersal ability inhibits speciation in a continental radiation of passerine birds.
+Proceedings of the Royal Society B: Biological Sciences 279:1567-1574.\\
+Collar, D. C., B. C. O'Meara, P. C. Wainwright, and T. J. Near. 2009. Piscivory limits diversification of feeding morphology in centrarchid fishes. Evolution 63:1557-1573.\\
+Collar, D. C., J. A. Schulte Ii, and J. B. Losos. 2011. Evolution of extreme body size disparity in monitor lizards (Varanus). Evolution 65:2664-2680.\\
+Cooper, N., and A. Purvis. 2010. Body size evolution in mammals: complexity in tempo and mode. The American Naturalist 175:727-738.\\
+Davis, R.
+B., J.
+Javois, J.
+Pienaar, E.
+Õunap, and T.
+Tammaru.
+2012.
+Disentangling determinants of egg size in the Geometridae (Lepidoptera) using an advanced phylogenetic comparative method.
+Journal of Evolutionary Biology 25:210-219.\\
+Derryberry, E.
+P., S.
+Claramunt, G.
+Derryberry, R.
+T.
+Chesser, J.
+Cracraft, A.
+Aleixo, J.
+Pérez-Emán, J.
+J.
+V.
+Remsen, and R.
+T.
+Brumfield.
+2011.
+Lineage diversification and morphological evolution in a large-scale continental radiation: the Neotropical ovenbirds and woodcreepers (Aves: Furnariidae).
+Evolution 65:2973-2986.\\
+Diniz-Filho, J.
+A.
+F., T.
+F.
+Rangel, T.
+Santos, and L.
+Mauricio Bini.
+2012.
+Exploring patterns of interspecific variation in quantitative traits using sequential phylogenetic eigenvector regressions.
+Evolution 66:1079-1090.\\
+Dumont, B. L., and B. A. Payseur. 2008. Evolution of the genomic rate of recombination in mammals Evolution 62:276-294.\\
+Edwards, E. J., and M. J. Donoghue. 2006. Pereskia and the origin of the cactus life-form. American Naturalist 167:777-793.\\
+Edwards, E. J., and S. A. Smith. 2010. Phylogenetic analyses reveal the shady history of C4 grasses. Proceedings of the National Academy of Sciences 107:2532-2537.\\
+Fr\'{e}d\'{e}rich, B., L.
+Sorenson, F.
+Santini, G.
+J.
+Slater, and M.
+E.
+Alfaro.
+2013.
+Iterative ecological radiation and convergence during the evolutionary history of damselfishes (Pomacentridae).
+The American Naturalist 181:94-113.\\
+Friedman, N.
+R., K.
+J.
+McGraw, and K.
+E.
+Omland.
+2013.
+Evolution of carotenoid pigmentation in caciques and meadowlarks (Icteridae): repeated gains of red plumage coloration by carotenoid C4-oxygenation.
+Evolution in press.\\
+Fusco, G., J. T. Garland, G. Hunt, and N. C. Hughes. 2012. Developmental trait evolution in trilobites. Evolution 66:314-329.\\
+Galvan, I., and A. P. Moller. 2011. Brain size and the expression of pheomelanin-based colour in birds. Journal of Evolutionary Biology 24:999-1006.\\
+Gomez, D., and M. Théry. 2007. Simultaneous crypsis and conspicuousness in color patterns: comparative analysis of a neotropical rainforest bird community. The American Naturalist 169:S42-S61.\\
+Gomez-Mestre, I., R. A. Pyron, and J. J. Wiens. 2012. Phylogenetic analyses reveal unexpected patterns in the evolution of reproductive modes in frogs. Evolution 66:3687-3700.\\
+Gonzalez-Voyer, A., and N.
+Kolm.
+2011.
+Rates of phenotypic evolution of ecological characters and sexual traits during the Tanganyikan cichlid adaptive radiation.
+Journal of Evolutionary Biology 24:2378-2388.\\
+Gonzalez-Voyer, A., S. Winberg, and N. Kolm. 2009. Social fishes and single mothers: brain evolution in African cichlids. Proceedings of the Royal Society B: Biological Sciences 276:161-167.\\
+Goodman, B. A., S. C. Hudson, J. L. Isaac, and L. Schwarzkopf. 2009. The evolution of body shape in response to habitat: is reproductive output reduced in flat lizards? Evolution 63:1279-1291.\\
+Guerrero, P.
+C., M.
+Rosas, M.
+T.
+K.
+Arroyo, and J.
+J.
+Wiens.
+2013.
+Evolutionary lag times and recent origin of the biota of an ancient desert (Atacama–Sechura).
+Proceedings of the National Academy of Sciences 110:11469-11474.\\
+Gvoždík, L., and R. V. Damme. 2006. Triturus newts defy the running-swimming dilemma. Evolution 60:2110-2121.\\
+Halsey, L. G., P. J. Butler, and T. M. Blackburn. 2006. A phylogenetic analysis of the allometry of diving. The American Naturalist 167:276-287.\\
+Hansen, T. F., and S. H. Orzack. 2005. Assessing current adaptation and phylogenetic inertia as explanations of trait evolution: the need for controlled comparisons. Evolution 59:2063-2072.\\
+Hansen, T. F., J. Pienaar, and S. H. Orzack. 2008. A comparative method for studying adaptation to a randomly evolving environment. Evolution 62:1965-1977.\\
+Harmon, L.
+J., J.
+B.
+Losos, T.
+Jonathan Davies, R.
+G.
+Gillespie, J.
+L.
+Gittleman, W.
+Bryan Jennings, K.
+H.
+Kozak, M.
+A.
+McPeek, F.
+Moreno Roark, and T.
+J.
+Near.
+2010.
+Early bursts of body size and shape evolution are rare in comparative data.
+Evolution 64:2385-2396.\\
+Helmus, M. R., W. Keller, M. J. Paterson, N. D. Yan, C. H. Cannon, and J. A. Rusak. 2010. Communities contain closely related species during ecosystem disturbance. Ecology Letters 13:162-174.\\
+Hertz, P. E., Y. Arima, A. Harrison, R. B. Huey, J. B. Losos, and R. E. Glor. 2013. Asynchronous evolution of physiology and morphology in Anolis lizards. Evolution 67:2101-2113.\\
+Hipp, A. L. 2007. Nonuniform processes of chromosome evolution in sedges (Carex: Cyperaceae). Evolution 61:2175-2194.\\
+Hossie, T. J., C. Hassall, W. Knee, and T. N. Sherratt. 2013. Species with a chemical defence, but not chemical offence, live longer. Journal of Evolutionary Biology 26:1598-1602.\\
+Huey, R.
+B., C.
+A.
+Deutsch, J.
+J.
+Tewksbury, L.
+J.
+Vitt, P.
+E.
+Hertz, H.
+J.
+Álvarez Pérez, and T.
+Garland.
+2009.
+Why tropical forest lizards are vulnerable to climate warming.
+Proceedings of the Royal Society of London Series B-Biological Sciences 276:1939-1948.\\
+Ingram, T., L.
+J.
+Harmon, and J.
+B.
+Shurin.
+2012.
+When should we expect early bursts of trait evolution in comparative data?
+Predictions from an evolutionary food web model.
+Journal of Evolutionary Biology 25:1902-1910.\\
+Ives, A. R., and H. C. J. Godfray. 2006. Phylogenetic analysis of trophic associations. The American Naturalist 168:E1-E14.\\
+Kalinka, A.
+T., K.
+M.
+Varga, D.
+T.
+Gerrard, S.
+Preibisch, D.
+L.
+Corcoran, J.
+Jarrells, U.
+Ohler, C.
+M.
+Bergman, and P.
+Tomancak.
+2010.
+Gene expression divergence recapitulates the developmental hourglass model.
+Nature 468:811-814.\\
+Kellermann, V., V.
+Loeschcke, A.
+A.
+Hoffmann, T.
+N.
+Kristensen, C.
+Fløjgaard, J.
+R.
+David, J.-C.
+Svenning, and J.
+Overgaard.
+2012a.
+Phylogenetic constraints in key functional traits behind species' climate niches: patterns of desiccation and cold resistance across 95 Drosophila species.
+Evolution 66:3377-3389.\\
+Kellermann, V., J.
+Overgaard, A.
+A.
+Hoffmann, C.
+Fløjgaard, J.-C.
+Svenning, and V.
+Loeschcke.
+2012b.
+Upper thermal limits of Drosophila are linked to species distributions and strongly constrained phylogenetically.
+Proceedings of the National Academy of Sciences 109:16228-16233.\\
+Knope, M. L., and J. A. Scales. 2013. Adaptive morphological shifts to novel habitats in marine sculpin fishes. Journal of Evolutionary Biology 26:472-482.\\
+Kostikova, A., G.
+Litsios, N.
+Salamin, and P.
+B.
+Pearman.
+2013.
+Linking life-history traits, ecology, and niche breadth evolution in North American eriogonoids (Polygonaceae).
+The American Naturalist 182:760-774.\\
+Kozak, K.
+H., R.
+W.
+Mendyk, and J.
+J.
+Wiens.
+2009.
+Can parallel diversification occur in sympatry?
+Repeated patterns of body-size evolution in coexisting clades of North American salamanders.
+Evolution 63:1769-1784.\\
+Kozak, K. H., and J. J. Wiens. 2010a. Accelerated rates of climatic-niche evolution underlie rapid species diversification. Ecology Letters 13:1378-1389.\\
+Kozak, K. H., and J. J. Wiens. 2010b. Niche conservatism drives elevational diversity patterns in Appalachian salamanders. The American Naturalist 176:40-54.\\
+Labra, A., J. Pienaar, and T. F. Hansen. 2009. Evolution of thermal physiology in Liolaemus lizards: adaptation, phylogenetic inertia, and niche tracking. The American Naturalist 174:204-220.\\
+Lambert, S. M., and J. J. Wiens. 2013. Evolution of viviparity: a phylogenetic test of the cold-climate hypothesis in phrynosomatid lizards. Evolution 67:2614-2630.\\
+Lapiedra, O., D. Sol, S. Carranza, and J. M. Beaulieu. 2013. Behavioural changes and the adaptive diversification of pigeons and doves. Proceedings of the Royal Society B: Biological Sciences 280.\\
+Litsios, G., R.
+O.
+Wüest, A.
+Kostikova, F.
+Forest, C.
+Lexer, H.
+P.
+Linder, P.
+B.
+Pearman, N.
+E.
+Zimmermann, and N.
+Salamin.
+2013.
+Effects of a fire response trait on diversification in replicated radiations.
+Evolution in press.\\
+López-Fernández, H., J. H. Arbour, K. O. Winemiller, and R. L. Honeycutt. 2013. Testing for ancient adaptive radiations in Neotropical cichlid fishes. Evolution 67:1321-1337.\\
+Machac, A., D. Storch, and J. J. Wiens. 2013. Ecological causes of decelerating diversification in carnivoran mammals. Evolution 67:2423–2433.\\
+Mahler, D. L., T. Ingram, L. J. Revell, and J. B. Losos. 2013. Exceptional convergence on the macroevolutionary landscape in island lizard radiations. Science 341:292-295.\\
+Maia, R., D. R. Rubenstein, and M. D. Shawkey. 2013. Key ornamental innovations facilitate diversification in an avian radiation. Proceedings of the National Academy of Sciences in press.\\
+Mirceta, S., A. V. Signore, J. M. Burns, A. R. Cossins, K. L. Campbell, and M. Berenbrink. 2013. Evolution of mammalian diving capacity traced by myoglobin net surface charge. Science 340:1234192.\\
+Moen, D.
+S., D.
+J.
+Irschick, and J.
+J.
+Wiens.
+2013.
+Evolutionary conservatism and convergence both lead to striking similarity in ecology, morphology and performance across continents in frogs.
+Proceedings of the Royal Society B: Biological Sciences 280:20132156.\\
+Nogueira, A., P.
+J.
+Rey, and L.
+G.
+Lohmann.
+2012.
+Evolution of extrafloral nectaries: adaptive process and selective regime changes from forest to savanna.
+Journal of Evolutionary Biology 25:2325-2340.\\
+Ord, T. J. 2012. Historical contingency and behavioural divergence in territorial Anolis lizards. Journal of Evolutionary Biology 25:2047-2055.\\
+Ord, T. J., L. King, and A. R. Young. 2011. Contrasting theory with the empirical data of species recognition. Evolution 65:2572-2591.\\
+Ord, T. J., J. A. Stamps, and J. B. Losos. 2010. Adaptation and plasticity of animal communication in fluctuating environments. Evolution 64:3134-3148.\\
+Oufiero, C.
+E., G.
+E.
+A.
+Gartner, S.
+C.
+Adolph, and T.
+Garland.
+2011.
+Latitudinal and climatic variation in body size and dorsal scale counts in Sceloporus lizards: a phylogenetic perspective.
+Evolution 65:3590-3607.\\
+Pearse, I. S., and A. L. Hipp. 2012. Global patterns of leaf defenses in oak species. Evolution 66:2272-2286.\\
+Pellissier, L., S.
+Rasmann, G.
+Litsios, K.
+Fiedler, A.
+Dubuis, J.
+Pottier, and A.
+Guisan.
+2012.
+High host-plant nitrogen content: a prerequisite for the evolution of ant–caterpillar mutualism?
+Journal of Evolutionary Biology 25:1658-1666.\\
+P\'{e}rez i de Lanuza, G., E.
+Font, and J.
+L.
+Monterde.
+2013.
+Using visual modelling to study the evolution of lizard coloration: sexual selection drives the evolution of sexual dichromatism in lacertids.
+Journal of Evolutionary Biology 26:1826-1835.\\
+Perez, S.
+I., J.
+Klaczko, G.
+Rocatti, and S.
+F.
+Dos Reis.
+2011.
+Patterns of cranial shape diversification during the phylogenetic branching process of New World monkeys (Primates: Platyrrhini).
+Journal of Evolutionary Biology 24:1826-1835.\\
+Pienaar, J., A. Ilany, E. Geffen, and Y. Yom-Tov. 2013. Macroevolution of life‐history traits in passerine birds: adaptation and phylogenetic inertia. Ecology Letters 16:571–576.\\
+Price, S. A., J. J. Tavera, T. J. Near, and P. Wainwright. 2012. Elevated rates of morphological and functional diversification in reef-dwelling haemulid fishes. Evolution 67:417–428.\\
+Price, S. A., P. C. Wainwright, D. R. Bellwood, E. Kazancioglu, D. C. Collar, and T. J. Near. 2010. Functional innovations and morphological diversification in parrotfish. Evolution 64:3057–3068.\\
+Quintero, I., and J. J. Wiens. 2013. Rates of projected climate change dramatically exceed past rates of climatic niche evolution among vertebrate species. Ecology letters 16:1095-1103.\\
+Raia, P., and S. Meiri. 2011. The tempo and mode of evolution: body sizes of island mammals. Evolution 65:1927-1934.\\
+Rezende, E. L., E. M. Albert, M. A. Fortuna, and J. Bascompte. 2009. Compartments in a marine food web associated with phylogeny, body mass, and habitat structure. Ecology Letters 12:779-788.\\
+Rezende, E. L., J. E. Lavabre, P. R. Guimarães, P. Jordano, and J. Bascompte. 2007. Non-random coextinctions in phylogenetically structured mutualistic networks. Nature 448:925-928.\\
+Rosas-Guerrero, V., M.
+Quesada, W.
+S.
+Armbruster, R.
+Pérez-Barrales, and S.
+D.
+Smith.
+2011.
+Influence of pollination specialization and breeding system on floral integration and phenotypic variation in Ipomoea.
+Evolution 65:350-364.\\
+Ryan, T. M., and C. N. Shaw. 2013. Trabecular bone microstructure scales allometrically in the primate humerus and femur. Proceedings of the Royal Society B: Biological Sciences 280.\\
+Sallan, L. C., and M. Friedman. 2012. Heads or tails: staged diversification in vertebrate evolutionary radiations. Proceedings of the Royal Society B: Biological Sciences 279:2025-2032.\\
+Santana, S. E., I. R. Grosse, and E. R. Dumont. 2012. Dietary hardness, loading behavior, and the evolution of skull form in bats. Evolution 66:2587-2598.\\
+Schmerler, S.
+B., W.
+L.
+Clement, J.
+M.
+Beaulieu, D.
+S.
+Chatelet, L.
+Sack, M.
+J.
+Donoghue, and E.
+J.
+Edwards.
+2012.
+Evolution of leaf form correlates with tropical–temperate transitions in Viburnum (Adoxaceae).
+Proceedings of the Royal Society B: Biological Sciences 279:3905-3913.\\
+Seddon, N., C.
+A.
+Botero, J.
+A.
+Tobias, P.
+O.
+Dunn, H.
+E.
+A.
+MacGregor, D.
+R.
+Rubenstein, J.
+A.
+C.
+Uy, J.
+T.
+Weir, L.
+A.
+Whittingham, and R.
+J.
+Safran.
+2013.
+Sexual selection accelerates signal evolution during speciation in birds.
+Proceedings of the Royal Society B: Biological Sciences 280.\\
+Setiadi, M.
+I., J.
+A.
+McGuire, R.
+M.
+Brown, M.
+Zubairi, D.
+T.
+Iskandar, N.
+Andayani, J.
+Supriatna, and B.
+J.
+Evans.
+2011.
+Adaptive radiation and ecological opportunity in Sulawesi and Philippine fanged frog (Limnonectes) communities.
+The American Naturalist 178:221-240.\\
+Slater, G.
+J., S.
+A.
+Price, F.
+Santini, and M.
+E.
+Alfaro.
+2010.
+Diversity versus disparity and the radiation of modern cetaceans.
+Proceedings of the Royal Society B: Biological Sciences 277:3097-3104.\\
+Smith, K. L., L. J. Harmon, L. P. Shoo, and J. Melville. 2011. Evidence of constrained phenotypic evolution in a cryptic species complex of agamid lizards. Evolution 65:976-992.\\
+Smith, N. D. 2012. Body mass and foraging ecology predict evolutionary patterns of skeletal pnematicity in the diverse "waterbird" clade. Evolution 66:1059-1078.\\
+Smith, S. D., C. Ané, and D. A. Baum. 2008. The role of pollinator shifts in the floral diversification of Iochroma (Solanaceae). Evolution 62:793-806.\\
+Sookias, R.
+B., R.
+J.
+Butler, and R.
+B.
+J.
+Benson.
+2012.
+Rise of dinosaurs reveals major body-size transitions are driven by passive processes of trait evolution.
+Proceedings of the Royal Society B: Biological Sciences 279:2180-2187.\\
+Spoor, F., T.
+Garland, G.
+Krovitz, T.
+M.
+Ryan, M.
+T.
+Silcox, and A.
+Walker.
+2007.
+The primate semicircular canal system and locomotion.
+Proceedings of the National Academy of Sciences 104:10808-10812.\\
+Stireman, J. O., H. Devlin, and P. Abbot. 2012. Rampant host- and defensive phenotype-associated diversification in a goldenrod gall midge. Journal of Evolutionary Biology 25:1991-2004.\\
+Stuart‐Fox, D., A. Moussalli, and M. J. Whiting. 2007. Natural selection on social signals: signal efficacy and the evolution of chameleon display coloration. The American Naturalist 170:916-930.\\
+Swanson, D. L., and J. T. Garland. 2009. The evolution of high summit metabolism and cold tolerance in birds and its impact on present-day distributions. Evolution 63:184-194.\\
+Tanabe, T., and T. Sota. 2013. Both male and female novel traits promote the correlated evolution of genitalia between the sexes in an arthropod. Evolution in press.\\
+Tulli, M. J., V. Abdala, and F. B. Cruz. 2011. Relationships among morphology, clinging performance and habitat use in Liolaemini lizards. Journal of Evolutionary Biology 24:843-855.\\
+Turbill, C., C.
+Bieber, and T.
+Ruf.
+2011.
+Hibernation is associated with increased survival and the evolution of slow life histories among mammals.
+Proceedings of the Royal Society B: Biological Sciences 278:3355-3363.\\
+Valido, A., H. M. Schaefer, and P. Jordano. 2011. Colour, design and reward: phenotypic integration of fleshy fruit displays. Journal of Evolutionary Biology 24:751-760.\\
+Valiente-Banuet, A., A.
+V.
+Rumebe, M.
+Verdú, and R.
+M.
+Callaway.
+2006.
+Modern Quaternary plant lineages promote diversity through facilitation of ancient Tertiary lineages.
+Proceedings of the National Academy of Sciences 103:16812-16817.\\
+Van Buskirk, J. 2009. Getting in shape: adaptation and phylogenetic inertia in morphology of Australian anuran larvae. Journal of evolutionary biology 22:1326-1337.\\
+Voje, K. L., and T. F. Hansen. 2012. Evolution of static allometries: adaptive change in allometric slopes of eye span in stalk-eyed flies. Evolution 67:453-467.\\
+Voje, K.
+L., A.
+B.
+Mazzarella, T.
+F.
+Hansen, K.
+Østbye, T.
+Klepaker, A.
+Bass, A.
+Herland, K.
+M.
+Bærum, F.
+Gregersen, and L.
+A.
+Vøllestad.
+2013.
+Adaptation and constraint in a stickleback radiation.
+Journal of evolutionary biology 26:2396-2414.\\
+Warne, R. W., and E. L. Charnov. 2008. Reproductive allometry and the size‐number trade‐off for lizards. The American Naturalist 172:E80-E98.\\
+Weir, J. T., and D. Wheatcroft. 2011. A latitudinal gradient in rates of evolution of avian syllable diversity and song length. Proceedings of the Royal Society B: Biological Sciences 278:1713-1720.\\
+Weir, J. T., D. J. Wheatcroft, and T. D. Price. 2012. The role of ecological constraint in driving the evolution of avian song frequency across a latitudinal gradient. Evolution 66:2773-2783.\\
+Wiens, J. J., K. H. Kozak, and N. Silva. 2013. Diversity and niche evolution along aridity gradients in North American lizards (Phrynosomatidae). Evolution 67:1715-1728.\\
+Wiens, J. J., R. A. Pyron, and D. S. Moen. 2011. Phylogenetic origins of local-scale diversity patterns and the causes of Amazonian megadiversity. Ecology Letters 14:643-652.
 

### 2015-04-15 R/OUtables.R
More tables

@@ -25,7 +25,7 @@ OUresTable[seq(7, 28, 7), "Tree size"] <- 1000
 # Now extract rejection rate and alpha estimates (median and 95% HPDs)
 
 # Read in data with no error (rename data object to avoid confusion with same name of object in error sims)
-load("OU_simulations.rda")
+load("Data/OU_simulations.rda")
 noerror_res <- out
 
 for (i in 1:28) {
@@ -34,22 +34,19 @@ for (i in 1:28) {
 	OUresTable[i,"Median alpha (95% HPD)"] <- paste(round(alpha[1], 3), " (", round(alpha[2], 3), "-", round(alpha[3], 3), ")", sep="")		
 	}
 
-write.csv(OUresTable, "tables/OU_bdTrees.csv")
+write.csv(OUresTable, "Manuscript/tables/OU_bd_noerror.csv")
 
 
 
 
 # Read in data with error (rename data object to avoid confusion with same name of object in error sims)
-load("OU_error_simulations.rda")
+load("Data/OU_error_simulations.rda")
 error_res <- out
 
 OUerrorTable <- as.data.frame(matrix(nrow=7, ncol=8))
 colnames(OUerrorTable) <- c("Rejection (0% error)", "Rejection (1% error)", "Rejection (5% error)", "Rejection (10% error)", "Alpha (0% error)", "Alpha (1% error)", "Alpha (5% error)", "Alpha (10% error)")
 
 rownames(OUerrorTable)<- c(25,50,100,150,200,500,1000)
-
-
-
 
 for (i in 1:7) {
 	OUerrorTable[i,"Rejection (0% error)"] <- sum(noerror_res[(1 + (i-1)*1000) : (i*1000), 5]>3.84)/1000	
@@ -68,65 +65,37 @@ for (i in 1:7) {
 	OUerrorTable[i,8] <- paste(round(alpha[1], 3), " (", round(alpha[2], 3), "-", round(alpha[3], 3), ")", sep="")
 		}
 
-
-write.csv(OUerrorTable, "tables/OU_yule_error_Trees.csv")
-
+write.csv(OUerrorTable, "Manucript/tables/OU_yule_error.csv")
 
 
+# Now for rate variable through time trees
+# Set up empty table with 4 columns and 28 rows, assign names to columns
[…]

### 2015-04-15 R/OUtables.R
Typo

@@ -95,7 +95,7 @@ for (i in 1:28) {
 	OUresVarRateTable[i,"Median alpha (95% HPD)"] <- paste(round(alpha[1], 3), " (", round(alpha[2], 3), "-", round(alpha[3], 3), ")", sep="")		
 	}
 
-write.csv(OUresVarRateTable, "Manuscript/tables/OU_var_rate_nerror.csv")
+write.csv(OUresVarRateTable, "Manuscript/tables/OU_var_rate_noerror.csv")
 
 
 

### 2015-04-16 Manuscript/ohyou.tex
Minor edits
Edits to structure

@@
 \usepackage{enumerate} 
+\usepackage{longtable}
+\usepackage{pdflscape}
+\usepackage{fullpage}
+\usepackage{hyperref}
 %\usepackage[nomarkers, nolists]{endfloat}  % To push floats to the end   
@@
   This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
-  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait mean was recast as a fitness optimum on an adaptive landscape. 
-  The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see Discussion).
-
-  The popularity of the OU model has grown extensively in recent years (Fig.
-\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
-% Need to add ref to supp mat
-  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
-  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
-  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  The model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait mean was recast as a fitness optimum on an adaptive landscape. 
+  The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see section X).
+  % Add section no here
   Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
 
-    \begin{figure}[h]
-      \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
-      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supplementary Material for details %linkto supp mat
-      }
-      \label{figure.literature}
-    \end{figure}
-
-  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material). 
-  Multiple models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
-  OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
-  The latter represents evolution under multiple selective regimes, and may be more biologically realistic. 
-  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
-
-  OU models are also commonly used to control for phylogeny in correlative analyses (Tables \ref{table.uses}; Supplementary Material).  
-  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
-  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data.
-  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
-  This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
-Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.
-
-  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material) therefore, we focus on this use of the OU model in our simulations below.
-  The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures.
-  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate, but our general conclusions also apply to multiple optima OU models.
-
-    \begin{table}[h]
-        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supplementary Material for details.}
-        \bigskip  
-          \begin{tabular}{p{8cm}lc}
-            \hline
-            Use of OU model & Optima & Number of papers\\
-            \hline
-            Ancestral state reconstruction & single & 8\\
-            & multiple & 2\\
-            Convergent evolution & single & 0\\
-            & multiple & 2\\
-            Model of evolution & single & 31\\
-            & multiple & 27\\
-            Phylogenetic generalised least squares & single & 35\\
-            & multiple & 0\\
-            Other & single & 5\\
-            & multiple & 5\\
-            Total & single &  36\\
-            & multiple & 79\\
-            \hline
-            \label{table.uses}
-          \end{tabular}
-    \end{table}
-
-  Here we present simulations demonstrating the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+% This paragraph may need modifying once we finish
+  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology. 
+  We use simulations to demonstrate the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
   We also show that very small amounts of measurement error in data can have profound effects on the performance of models. 
-  Many of our findings are applicable to other models of evolution (see Discussion and Supplemental Figs. S6-S11) 
-  % Add figure refs here
+  Many of our findings are applicable to other models of evolution (see section X) 
+  % Add section no here
   , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
@@
 
-\section{Materials and Methods}
-  \subsection{The Ornstein Uhlenbeck (OU) model}
+\section{The Ornstein Uhlenbeck (OU) model}
+
   According to the Brownian model \citep{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
@@
 
+% Could move this down to the section on recommendations
     $\alpha$ can more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
@@
 
-  \subsection{Known statistical issues with the OU model}
-  ???Problems with the OU model and where they are reported in the literature?
+  % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
+  \section{Uses of OU models}
+  The popularity of the OU model has grown extensively in recent years (Fig.
+\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
+% Need to add ref to supp mat
+  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
+  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
+  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+
+    \begin{figure}[h]
+      \centering
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
+      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+See Supplementary Material for details %linkto supp mat
+      }
+      \label{figure.literature}
+    \end{figure}
+
+  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material). 
+  Multiple models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
+  OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
+  The latter represents evolution under multiple selective regimes, and may be more biologically realistic. 
+  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
+
+  OU models are also commonly used to control for phylogeny in correlative analyses (Tables \ref{table.uses}; Supplementary Material).  
+  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
+  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data.
+  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
+  This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
+Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.
+
+  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material) therefore, we focus on this use of the OU model in our simulations below.
+  The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
+  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate, but our general conclusions also apply to multiple optima OU models.
+
+    \begin{table}[h]
+      \begin{center}
+        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supplementary Material for details.}
+        \bigskip  
+          \begin{tabular}{p{8cm}lc}
+            \hline
+            Use of OU model & Optima & Number of papers\\
+            \hline
+            Ancestral state reconstruction & single & 8\\
+            & multiple & 2\\
+            Convergent evolution & single & 0\\
+            & multiple & 2\\
+            Model of evolution & single & 31\\
+            & multiple & 27\\
+            Phylogenetic generalised least squares & single & 35\\
+            & multiple & 0\\
+            Other & single & 5\\
+            & multiple & 5\\
+            Total & single &  79\\
+            & multiple & 36\\
+            \hline
+            \label{table.uses}
+          \end{tabular}
+      \end{center}
+    \end{table}
+
+\section{Known statistical issues with the OU model}
+% Problems with the OU model and where they are reported in the literature?
+% Could rename this section also or move it. Whatever makes most sense.
 
     \begin{enumerate}
-      \item $\alpha$ is bounded at zero.
+      \item $\alpha$ has a hard bound at zero, but no upper bound, meaning that the Maximum Likelihood profile is asymmetric.
+As values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small.
+This suggests that estimates of $\alpha$ are likely to be biased
+      % make sure we really mean biased here
+      , and that the asymptotic assumptions of Likelihood ratio tests on $\alpha$ are likely to be violated.
       \item blah
-      \item measurement errors
+      \item measurement errors?
     \end{enumerate}
 
-  \subsection{Maximum Likelihood simulations}
-  \label{section:sims.methods} 
-    We used simulations to explore the statistical properties of $\alpha$. 
+% I've separated this out a bit to try and make the progression more logical
+% It's all a bit drafty at present so feel free to edit heavily - you've a better idea than me about the results anyway
+\section{Performance of OU models using Maximum Likelihood simulations}
+  To explore some issues with the OU model in more detail, we ran a number of simulations designed to mirror the use of OU models in the literature.
+
+  \subsection{Simulating phylogenies and data}
     We simulated phylogenies with 25, 50, 100, 150, 200, 500, or 1000 tips under pure birth, constant-rate Birth-Death (extinction fractions of 0.25, 0.5 and 0.75), or temporally varying speciation rate (speciation rate modeled as time from the root raised to the power 0.2, 0.5, 2 and 5) models.
@@
     We then simulated the evolution of a single trait under a Brownian motion model on each phylogeny using the R package MOTMOT \citep{Thomas:2011aa}. 
-    We estimated $\alpha$ and compared the fit of a Brownian model to that of an OU model using a likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
-    This procedure mirrors a situation where the OU model  %%% argh
+    All our simulated trees and data are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
 
-  \subsection{Effects of measurement error}
-    Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015} so we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
+  \subsection{Performance of $\alpha$ and Likelihood ratio tests}
+    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
+    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model \citep{}. % Add citations
+    We then estimated the rejection rate for each set of simulations, i.e., the proportion of times OU models are favoured relative to a Brownian motion model, even though the data were simulated under Brownian model.
+This gives us a measure of Type I error rate for the OU model.
+
+% I ended up just LaTeXing this for our own amusement
+      \begin{table}[!htbp]
+        \begin{center}
+        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees.
+Tree type refers to the extinction fraction for the birth-death trees.
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+        \bigskip
+        \input{Tables/OU_bd_noerror.tex}
+        \label{tables:bd_noerror}
+        \end{center}
+      \end{table}
+
+    We find that Type I error rates are unacceptably high when tree size is small (Table \ref{tables:bd_noerror}). 
+    %For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips.
+    % I think we're cutting this?
+    This has clear implications for studies with low sample sizes where OU is favoured.
+
+% Could put in something about interpreting parameters here?
+
+\section{Effects of measurement error on OU model performance}
+  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{} % add citation to Matt's model adequacy stuff
+  Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
+
+  \subsection{Simulating phylogenies and data with error}
     We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
+    % should we say "as above" here or be specific i.e. copy paste from above?
     Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees.
     We then simulated data under a Brownian model on each tree. 
+    All our simulated trees and data are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
+
+  \subsection{Effects of measurement error on OU model performance}
     We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges.
@@
 
-  \subsection{Other evolutionary models}
-    Although our focus is on the OU model we also explored three other commonly used models of trait evolution -- $\kappa$, $\lambda$, and $\delta$ \citep{Pagel:1997aa,Pagel:1999aa} -- fitted using Maximum Likelihood with the same simulated data.
-    These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
-    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the Supplementary Material.
+% OK this is still broken!!!
+  \begin{landscape}
+      \begin{table}[!htbp]
+        \begin{center}
+        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.).
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+        \bigskip
+        \input{Tables/OU_yule_error.tex}
+        \label{tables:yule_error}
+        \end{center}
+      \end{table}
+  \end{landscape}
 
-\section{Results}
-  \subsection{Maximum Likelihood simulations}
-  \label{section:sims.results}
-
-    Figure \ref{figure.profiles} shows some examples of profile likelihoods for $\alpha$ generated on different datasets that were simulated according to a Brownian model.
-It is clear that the shape of the tree affects the shape of the likelihood profile.
-Trees with increasing speciation rates from root to tip (i.e., that have a lot of recently evolved species) generate very wide, flat likelihood profiles (Fig.
-\ref{figure.profiles}A).
-On the other hand, trees with declining speciation rates from root to tip (Fig.
-\ref{figure.profiles}B), or generated according to a Yule process (Fig.
-\ref{figure.profiles}C), yield more strongly peaked profiles.
-
-%\begin{figure}
- % \centering
-  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure3.pdf}
-  %\caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ‘true’ value of alpha is 0.}
-%\label{figure.profiles}
-%\end{figure}
-
-In all three cases, there are several features worth noting.
-First, the profiles are asymmetric.
-There is a hard bound at $\alpha = 0$, and as $\alpha$ becomes large, the rate of decrease in the likelihood slows.
-This is because, as shown in Figure \ref{figure.traitsim}, as values of $\alpha$ become large, the effects of changing $\alpha$ on model predictions are increasingly small.
-It is also notable, that the Maximum Likelihood estimates of $\alpha$ are, in most cases, between 0 and 1.
-In Figure \ref{figure.traitsim} it was noted that in this region of parameter space, the OU model is very difficult to distinguish from Brownian.
-This, combined with the asymmetric nature of the profile likelihood is suggestive that estimates of $\alpha$ are likely to be biased, and that the asymptotic assumptions of likelihood ratio tests on $\alpha$ are likely to be violated.
-
-The effects of the asymmetric profile likelihoods become clear when calculating Type I error for the OU model based on ML estimates across a range of tree shapes and sizes (Fig.
-\ref{figure.rejection}; see Supplemental Figs.
-S2 and S3 for results with alternative priors for the Bayesian analyses).
-Regardless of tree shape, Type I error rates are unacceptably high when tree size is small (Figs.
-\ref{figure.rejection}A, \ref{figure.rejection}C and \ref{figure.rejection}E).
-For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips.
-Bayesian estimation is not subject to bias % check in the log-likelihood statistic and consistently rejects the OU model.
-However, estimates of α are similar regardless of method.
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure4.pdf}
-  %\caption{Rejection rates and estimation of $\alpha$ for different tree sizes and shapes.
-Rejection rates (a,c,e) for the OU model at multiple tree sizes for models fitted using ML (circles) and Bayesian (diamonds) methods.
-Estimates of $\alpha$ (b,d,f) based on ML (circles show means and dotted lines show 95\% sampling interval from 1000 trees) and Bayesian (diamonds show means and dotted lines show 95\% sampling interval of the modal value from the posterior distribution from 1000 trees).}
-%\label{figure.rejection}
-%\end{figure}
-
-%Figure \ref{figure.error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error (also see Figs.
-S4 and S5).
+    Table \ref{tables:yule_error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error.
 The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error).
@@
 First, the frequency with which the OU model is favoured increases with tree size for both ML and Bayesian estimation (Fig.
-\ref{figure.error}A).
+\ref{tables:yule_error}).
 With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species.
@@
 We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for.
-Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1 (Fig.
-\ref{figure.error}B).
+Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1.
 Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.
 
-%\begin{figure}
- % \centering
-  %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OU_figure5.pdf}
-  %\caption{Rejection of BM with measurement error (a) and estimation of α (b). Symbols follow Figure \ref{figure.rejection}.}
-%\label{figure.error}
-%\end{figure}
-
-The bias % check
- reported above are not unique to the OU model.
-The parameter $\delta$, which tests for accelerating or decelerating rates of evolution \citep{Pagel:1997aa,Pagel:1999aa} also suffers from elevated Type I error and is upwardly biased, favoring models that imply temporally increasing evolutionary rates (Supplemental Fig.
-S8).
-In contrast, although the speciational model ($\kappa$) has slightly elevated Type I errors, the parameter estimates are unbiased for the majority of tree sizes and shapes (Supplemental Fig.
-S6).
-As previously shown (Freckleton et al.
-2002), the $\lambda$ model is conservative with low Type I error rates (Supplemental Fig.
-S7).
-As with the OU model, the $\delta$ models are all strongly favored over Brownian motion when there is error in the species data (Supplemental Figs.
-S9-S11).
-
-\section{Discussion}
-\label{section:discussion}
+\section{Prospects/ recommendations}
+  \subsection{Brief discussion on methods and limitations for including measurement error}
 
@@
 
-%Results from \citet{ho2013asymptotic}, indicate that there is the potential for revealing strong biases in $\alpha$ , but the implications for practice (e.g.
-likelihood of rejection of alternative models) is unexplored other than to note that tree shape may play an influence.
+\section{Bayesian methods (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)}
 
+\section{Just an OU problem? - Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
+
+ % The bias % check we mean bias
+ % reported above are not unique to the OU model.
+The parameter $\delta$, which tests for accelerating or decelerating rates of evolution \citep{Pagel:1997aa,Pagel:1999aa} also suffers from elevated Type I error and is upwardly biased, favoring models that imply temporally increasing evolutionary rates (Supplemental Fig.
+S8).
+In contrast, although the speciational model ($\kappa$) has slightly elevated Type I errors, the parameter estimates are unbiased for the majority of tree sizes and shapes (Supplemental Fig.
+S6).
+As previously shown (Freckleton et al.
+2002), the $\lambda$ model is conservative with low Type I error rates (Supplemental Fig.
+S7).
+As with the OU model, the $\delta$ models are all strongly favored over Brownian motion when there is error in the species data (Supplemental Figs.
+S9-S11).
+
+\section{Concluding comments/Conclusions}
+\section{Epilogue - vaguely reproducible science}
+
+%--------------------------------------
+% I've left all this as is for now as we can slot relevant paragraphs into the appropriate places
+
+\section{Discussion}
+
+% This came from the intro but fits better later
 %The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
@@
 The results presented above highlight that in the case of the OU model, this is especially important.
+
+  \subsection{Other evolutionary models}
+    Although our focus is on the OU model we also explored three other commonly used models of trait evolution -- $\kappa$, $\lambda$, and $\delta$ \citep{Pagel:1997aa,Pagel:1999aa} -- fitted using Maximum Likelihood with the same simulated data.
+    These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
+    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the Supplementary Material.
 

### 2015-04-17 Manuscript/ohyou.tex
fixed tables and added var error

@@
 
-    We find that Type I error rates are unacceptably high when tree size is small (Table \ref{tables:bd_noerror}). 
-    %For some tree shapes (particularly where speciation rates accelerate towards the present), Type I error remains $>0.05$ even for trees with 1000 tips.
-    % I think we're cutting this?
+      \begin{table}[!htbp]
+        \begin{center}
+        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on trees simulated under time variable speciation rates.
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+        \bigskip
+        \input{Tables/OU_var_rate_noerror.tex}
+        \label{tables:varrate_noerror}
+        \end{center}
+      \end{table}
+
+    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was actually used to generate the data.
+    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
     This has clear implications for studies with low sample sizes where OU is favoured.
@@
 
-% OK this is still broken!!!
   \begin{landscape}
@@
 There are two points worth noting.
-First, the frequency with which the OU model is favoured increases with tree size for both ML and Bayesian estimation (Fig.
-\ref{tables:yule_error}).
+First, the frequency with which the OU model is favoured increases with tree size.
 With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species.
@@
 \section{Bayesian methods (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)}
+
+\subsection{Bayesian simulations} % keep? or move it all to supp mat?
+    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
+    The choice of priors for OU models has not been fully explored. 
+    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
+    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
+    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
+S2-S5).
+% add links to this
+    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
+    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
+    Multiple independent chains were run for each analysis to ensure convergence was reached. 
+    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
+    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
+    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
+    We treated Bayes factors $>$ 2 as evidence favouring the OU model.
 

### 2015-04-17 Manuscript/ohyou.tex
Added ntaxa figure

@@
 
-% I ended up just LaTeXing this for our own amusement
+% I ended up just LaTeXing the tables for my own amusement
       \begin{table}[!htbp]
@@
 
-    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was actually used to generate the data.
+    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data.
     For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
-    This has clear implications for studies with low sample sizes where OU is favoured.
+    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
+    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure ; see Supplemental Material). % link to supp mat
+
+    \begin{figure}[h]
+      \centering
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/NTaxa.png}
+      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+Two studies with \morethan 3000 taxa have been omitted for clarity.
+See Supplementary Material for details %linkto supp mat
+      }
+      \label{figure.ntaxa}
+    \end{figure}
 
@@
 
-\section{Bayesian methods (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)}
+\section{Bayesian methods}
+ (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
 
@@
 
-\section{Just an OU problem? - Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
+\section{Just an OU problem?}
+Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
 

### 2015-04-17 Analyses/OhYouFigures.R

@@ -18,11 +18,12 @@ points(pal ~ Year, data = ds, cex = 1.2, pch = 6)
 legend("topleft", legend = c("ecology", "evolution", "palaeontology"), pch = c(1, 16, 6), cex = 1.2, bty = "n")
 dev.off()
 
-# Read in data for figure 2 and reshape
+# Read in data for table 2 and figure 3 and reshape
 ds1 <- read.delim("Data/literature.txt")
 ds1 <- subset(ds1, Noptima != "?")
 ds1$Noptima <- ds1$Noptima[, drop = TRUE]
 
+# Table 2
 omit <- which(ds1$General.use == "other" | ds1$General.use == "convergent evolution" |
 	  ds1$General.use == "ancestral state reconstruction" | 
 	  ds1$General.use == "phylogenetic inertia" | ds1$General.use == "phylogenetic signal" )
@@ -31,3 +32,10 @@ ds2 <- ds1[-omit, ]
 ds2$General.use <- ds2$General.use[, drop = TRUE]
 
 summary.table <- tapply(ds2$Paper, list(ds2$General.use, ds2$Noptima), length)
+
+# Figure 3
+png("Manuscript/Figures/NTaxa.png")
+  par(bty = "l")
+  hist(ds1$Ntips[which(ds1$Ntips <= 1000)], las = 1, cex.lab = 1.2, cex.axis = 1.2,
+       ylab = "Number of papers", xlab = "Number of taxa", ylim = c(0, 80), main = "")
+dev.off()

### 2015-04-21 Manuscript/ohyou.tex

@@
 
-% Could move this down to the section on recommendations
-    $\alpha$ can more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
-
-      \begin{equation}
-        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
-        \label{equation:halflife}
-      \end{equation}
-
-    \noindent
-    This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
-    If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
-
   % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
@@
   For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading.
+As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche.
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
 
@@
       \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supplementary Material for details %linkto supp mat
+See Supplementary Material for details.
+%linkto supp mat
       }
@@
 
-  OU models are also commonly used to control for phylogeny in correlative analyses (Tables \ref{table.uses}; Supplementary Material).  
+  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supplementary Material).  
   Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
@@
     \begin{enumerate}
-      \item $\alpha$ has a hard bound at zero, but no upper bound, meaning that the Maximum Likelihood profile is asymmetric.
+      \item $\alpha$ has a hard bound at zero but no upper bound, meaning that the Maximum Likelihood profile is asymmetric.
 As values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small.
@@
     We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data.
-    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
+    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
     This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
-    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure ; see Supplemental Material). % link to supp mat
+    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supplemental Material). % link to supp mat
 
@@
       \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
-Two studies with \morethan 3000 taxa have been omitted for clarity.
-See Supplementary Material for details %linkto supp mat
+Two studies with \textgreater 3000 taxa have been omitted for clarity.
+See Supplementary Material for details.
+%linkto supp mat
       }
-      \label{figure.ntaxa}
+      \label{figure:ntaxa}
     \end{figure}
 
-% Could put in something about interpreting parameters here?
+    We provide recommendations related to these findings below. % name of section?
 
 \section{Effects of measurement error on OU model performance}
-  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{} % add citation to Matt's model adequacy stuff
+  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{}.
+% add citation to Matt's model adequacy stuff
   Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
@@
     Table \ref{tables:yule_error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error. 
-The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error).
-There are two points worth noting.
+    The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error). There are two points worth noting. 
     First, the frequency with which the OU model is favoured increases with tree size.
@@
     We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. 
-Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1.
+    Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently \textgreater 1. 
     Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.
 
-\section{Prospects/ recommendations}
-  \subsection{Brief discussion on methods and limitations for including measurement error}
+\section{Prospects: limitations, recommendations and potential solutions?}
+  \subsection{Limitations of the OU model}
+  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways.
+  Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. 
+  But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. 
+  The results of our simulations illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supplementary Table S1).
 
-  Recommendations
+  In the case of the OU model, there are several limitations worth highlighting:
+\begin{enumerate}[(i)]
+  \item \textbf{Type I error rates are high when sample size is low.} 
+  The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
+  \item \textbf{Likelihood ratio tests are untrustworthy.}
+  We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
+  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered.
+  \item \textbf{Measurement error increases Type I error rates.} 
+  Our results indicate that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
+  The effects of measurement error become more severe with increasing tree size.
+\end{enumerate}
 
-Interpreting results
- - OU is favoured over BM. If alpha is small it's basically BM. What
-is alpha? Talk about half lives. add in ives and garland advice on what is big and what is small here?
- - Not stabilising selection
- - Not an issue if you're just mopping up error structure for PGLS (?)
- - If really want to be certain OU over BM you need 200 species
- - Account for measurement error
+  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution. 
+  This is particularly important for inference of phylogenetic niche conservatism. 
+  Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
+  A more parsimonious explanation would be that there is measurement error in the data.
+  This seems particularly likely for climatic niche traits that are often averaged across an entire species range. 
+  Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.
 
-\section{Bayesian methods}
- (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
+% Not sure this is needed anywhere
+%In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process.
+This is because according to the OU process traits move back towards the optimum as their values become particularly large or small.
+Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig.
+\ref{figure.traitsim}).
+Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.
 
-\subsection{Bayesian simulations} % keep? or move it all to supp mat?
+\subsection{Potential solutions and recommendations}
+  Although there are limitations with the OU model, there are also several ways researchers can get around these problems as follows:
+
+  \begin{enumerate}[(i)]
+    \item \textbf{Use at least 200 species.} 
+    If you really want to be confident that the OU model is favoured over the Brownian model, you need to fit the models to more than 200 species. 
+    Note that this does not solve the issue of measurement error. in the data.
+    \item \textbf{Simulate fitted models and compare these to your empirical results.}
+    A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values.
+These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}.
+This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests.
+Any given tree will therefore generate unique parameter estimates.
+Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}.
+Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.
+    \item \textbf{Consider plausible alternative hypotheses.}
+    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$parameter when an OU model is favoured. 
+    We explain this in detail below.
+    \item \textbf{Account for measurement error.} % details
+    % This is from the original Discussion
+    %In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
+It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
+However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
+This is because the variances will themselves be estimates and subject to error.
+Treating such variances as error-free can potentially create biases (Freckleton unpub).
+If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance.
+However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
+    \item \textbf{If possible, include ancillary data.}
+    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
+    This is implemented in GEIGER \citep{Harmon:2008aa}.
+    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances (see Appendix \ref{equation:OUcov} depend on both the shared distances between species and the distance of a node to the nearest tip.
+    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
+    Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
+    These implementations should not be used with fossil data.
+    \item \textbf{Use a Bayesian approach.} See below. %Or cite our preprint?
+\end{enumerate}
+
+  Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
+  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
+  Thus this does not reflect any kind of OU process underlying the data.
+
+  To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
+  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.).
+
+  $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+
+      \begin{equation}
+        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
+        \label{equation:halflife}
+      \end{equation}
+
+  \noindent
+  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
+  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
+  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
+  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process.
+
+% add in ives and garland advice on what is big and what is small here? Or earlier?
+
+\subsection{Bayesian simulations} 
+% (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
     For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
@@
 \section{Just an OU problem?}
-Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
+%Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
+
+%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
+$\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
+It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$.
+Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously.
+\citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989).
+They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes).
+This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.
 
@@
 
-\section{Concluding comments/Conclusions}
-\section{Epilogue - vaguely reproducible science}
-
-%--------------------------------------
-% I've left all this as is for now as we can slot relevant paragraphs into the appropriate places
-
-\section{Discussion}
-
-% This came from the intro but fits better later
-%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
-$\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
-It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$.
-Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously.
-\citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989).
-They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes).
-This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.
-
-Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways.
-Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes.
-But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation.
-The results we have presented here illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used.
-
-In the case of the OU model, we can make a series of specific recommendations for future analyses:
-\begin{enumerate}[(i)]
-  \item At least 200 species should be included.
-The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.
-  \item Likelihood ratio tests are untrustworthy.
-We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances.
-The simulations indicate that likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered.
-  \item Simulate fitted models and compare these to your empirical results.
-A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values.
-These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}.
-This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests.
-Any given tree will therefore generate unique parameter estimates.
-Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}.
-Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.
-  \item Consider plausible alternative hypotheses.
-The results indicate that a simple parsimonious model can be mistaken for an OU process: namely a BM process in which a small amount of error is added to the data.
-Small ($<$ 1) values of $\alpha$ are quite compatible with this interpretation of the data.
-The effects of error become more severe with increasing tree size.
-\end{enumerate}
-
-In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
-It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
-However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
-This is because the variances will themselves be estimates and subject to error.
-Treating such variances as error-free can potentially create biases (Freckleton unpub).
-If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance.
-However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
-
-Our review of the literature revealed that the OU model is frequently described and interpreted as a model of ‘stabilizing selection’.
-The OU model is attractive because it sets effective bounds in probability on the size of species’ traits, but to use the term ‘stabilizing selection’ is inaccurate and misleading.
-As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche.
-Under this formulation, α can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
-
-In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process.
-This is because according to the OU process traits move back towards the optimum as their values become particularly large or small.
-Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig.
-\ref{figure.traitsim}).
-Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.
-This interpretation is particular important for inference of phylogenetic niche conservatism.
-Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
-A more parsimonious explanation would be that there is measurement error in the data.
-This seems particularly likely for climatic niche traits that are often averaged across an entire species range.
-Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.
-
-If ancillary data are available, however, this may not be the case.
-If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.
-A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances in equation (6) depend on both the shared distances between species and the distance of a node to the nearest tip.
-This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G.
-Slater unpub.).
-Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}.
-These implementations should not be used with fossil data.
-
-The results we report are not unique to the OU model.
+ %The results we report are not unique to the OU model.
 As noted above, \citet{grafen1989phylogenetic} pointed out that his $\rho$ parameter would likely be biased, and other transformations \citep[e.g., $\lambda$, $\delta$, ACDC;][]{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa} will exhibit similar behavior \citep[e.g., see][]{freckleton2002phylogenetic}.
@@
 
-In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species.
+\section{Concluding comments/Conclusions}
+% Conclusion from original
+%In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species.
 The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes.
@@
 
-  \subsection{Other evolutionary models}
-    Although our focus is on the OU model we also explored three other commonly used models of trait evolution -- $\kappa$, $\lambda$, and $\delta$ \citep{Pagel:1997aa,Pagel:1999aa} -- fitted using Maximum Likelihood with the same simulated data.
-    These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
-    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the Supplementary Material.
+\section{Epilogue: (vaguely) reproducible science}
+At the symposium that generated this special issue, one of us (NC) gave a talk on reproducibility in science. 
+We have therefore tried to make this paper as reproducible as possible. 
+All simulated phylogenies, data, and R code are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
+However, although our R code is provided, it's messy and difficult to use. 
+We also do not provide an automated way of reproducing the data collection for our literature review.
+Nor do we use tools like Travis CI, Docker, or packrat etc.
+% Add stuff about BayesTraits here. 
+We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 
+With a little effort most people should be able to produce something vaguely reproducible, and move us all slightly closer to truly reproducible science.
 
@@
 
+% Bibliography
 \bibliographystyle{biolinnsoc}
@@
 
+% Appendix
 \section{Appendix}
@@
 
+% Supplementary Material
 \input{ohyou_suppmat}

### 2015-04-21 Manuscript/ohyou_suppmat.tex

@@
 \section{Supplementary Material}
+\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
 
@@
     Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. 
-    We summarise these results in Figure 1 and Table 1, and the full dataset is available below along with the full list of references.
+    We summarise these results in Figures 1 and 2 and Table 1, and the full dataset is available below along with the full list of references.
 
     In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. 1). 
-    Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Supplemental Fig. S1 and Table S1). 
-    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the ``best'' model of trait evolution or to control for phylogenetic non-independence (Supplemental Table S1 and Fig.
-S1).
+    Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Figure 2 and Table S1). 
+    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the ``best'' model of trait evolution or to control for phylogenetic non-independence (Table S1).
 
@@
 All packages mentioned are R packages apart from BayesTraits, COMPARE, MATLAB, Mesquite, PAM, PDAP and PDTREE.}\\
+\bigskip        
 \hline
-
-\textbf{Paper}  &   Year    &   Journal &   Taxon   &   Ntips   &   Use in paper    &   Stats/R package \\
+\textbf{Paper}  &   \textbf{Year}    &   \textbf{Journal} &   \textbf{Taxon}   &   \textbf{Ntips}   &   \textbf{Use in paper}   &   \textbf{Stats/R package} \\
 \hline
@@
 Edwards and Donoghue 2006     &   2006    &   Am Nat  &   plants  &   12  &   ancestral state reconstruction  &   COMPARE \\
-Gvoždík and Damme 2006    &   2006    &   Evolution   &   amphibians  &   10  &   phylogenetic correction &   COMPARE \\
+Gvo\v{z}d\'{i}k and Damme 2006    &   2006    &   Evolution   &   amphibians  &   10  &   phylogenetic correction &   COMPARE \\
 Halsey et al. 2006    &   2006    &   Am Nat  &   birds/mammals   &   90  &   phylogenetic correction &   Custom code \\
@@
 Clabaut et al. 2007   &   2007    &   Evolution   &   fish    &   45  &   phylogenetic correction &   APE \\
-Gomez and Théry 2007  &   2007    &   Am Nat  &   birds   &   40  &   model of evolution  &   OUCH    \\
+Gomez and Th\'{e}ry 2007  &   2007    &   Am Nat  &   birds   &   40  &   model of evolution  &   OUCH    \\
 Hipp 2007 &   2007    &   Evolution   &   plants  &   53  &   model of evolution  &   BayesTraits \\
@@
 Weir et al. 2012  &   2012    &   Evolution   &   birds   &   232 &   model of evolution  &   GEIGER  \\
-Arbour and López-Fernández 2013   &   2013    &   PRSB    &   fish    &   27  &   model of evolution  &   OUCH    \\
+Arbour and L\'{o}pez-Fern\'{a}ndez 2013   &   2013    &   PRSB    &   fish    &   27  &   model of evolution  &   OUCH    \\
 Benesh et al. 2013    &   2013    &   Am Nat  &   helminths   &   143 &   phylogenetic correction &   APE \\
@@
 Christin et al. 2013  &   2013    &   PNAS    &   plants  &   545 &   model of evolution  &   GEIGER/OUCH \\
-Frédérich et al. 2013 &   2013    &   Am Nat  &   fish    &   208 &   model of evolution  &   OUwie   \\
+Fr\'{e}d\'{e}rich et al. 2013 &   2013    &   Am Nat  &   fish    &   208 &   model of evolution  &   OUwie   \\
 Friedman et al. 2013  &   2013    &   Evolution   &   birds   &   15  &   model of evolution  &   OUCH    \\
@@
 Litsios et al. 2013   &   2013    &   Evolution   &   plants  &   382 &   model of evolution  &   OUwie   \\
-López-Fernández et al. 2013   &   2013    &   Evolution   &   fish    &   135 &   model of evolution  &   GEIGER  \\
+L\'{o}pez-Fern\'{a}ndez et al. 2013   &   2013    &   Evolution   &   fish    &   135 &   model of evolution  &   GEIGER  \\
 Machac et al. 2013    &   2013    &   Evolution   &   mammals &   231 &   model of evolution  &   GEIGER  \\
@@
 Moen et al. 2013  &   2013    &   PRSB    &   amphibians  &   44  &   convergent evolution    &   GEIGER  \\
-Pérez i de Lanuza et al. 2013 &   2013    &   JEB &   reptiles    &   42  &   ancestral state reconstruction  &   GEIGER  \\
+P\'{e}rez i de Lanuza et al. 2013 &   2013    &   JEB &   reptiles    &   42  &   ancestral state reconstruction  &   GEIGER  \\
 Pienaar et al. 2013   &   2013    &   ELE &   birds   &   382 &   model of evolution  &   SLOUCH  \\
@@
 H., and H.
-López-Fernández.
+L\'{o}pez-Fern\'{a}ndez.
 2013.
@@
 Galvan, I., and A. P. Moller. 2011. Brain size and the expression of pheomelanin-based colour in birds. Journal of Evolutionary Biology 24:999-1006.\\
-Gomez, D., and M. Théry. 2007. Simultaneous crypsis and conspicuousness in color patterns: comparative analysis of a neotropical rainforest bird community. The American Naturalist 169:S42-S61.\\
+Gomez, D., and M. Th\'{e}ry. 2007. Simultaneous crypsis and conspicuousness in color patterns: comparative analysis of a neotropical rainforest bird community. The American Naturalist 169:S42-S61.\\
 Gomez-Mestre, I., R. A. Pyron, and J. J. Wiens. 2012. Phylogenetic analyses reveal unexpected patterns in the evolution of reproductive modes in frogs. Evolution 66:3687-3700.\\
@@
 Proceedings of the National Academy of Sciences 110:11469-11474.\\
-Gvoždík, L., and R. V. Damme. 2006. Triturus newts defy the running-swimming dilemma. Evolution 60:2110-2121.\\
+Gvo\v{z}d\'{i}k, L., and R. V. Damme. 2006. Triturus newts defy the running-swimming dilemma. Evolution 60:2110-2121.\\
 Halsey, L. G., P. J. Butler, and T. M. Blackburn. 2006. A phylogenetic analysis of the allometry of diving. The American Naturalist 167:276-287.\\
@@
 J.
-Álvarez Pérez, and T.
+\'{A}lvarez P\'{e}rez, and T.
 Garland.
@@
 Evolution in press.\\
-López-Fernández, H., J. H. Arbour, K. O. Winemiller, and R. L. Honeycutt. 2013. Testing for ancient adaptive radiations in Neotropical cichlid fishes. Evolution 67:1321-1337.\\
+L\'{o}pez-Fern\'{a}ndez, H., J. H. Arbour, K. O. Winemiller, and R. L. Honeycutt. 2013. Testing for ancient adaptive radiations in Neotropical cichlid fishes. Evolution 67:1321-1337.\\
 Machac, A., D. Storch, and J. J. Wiens. 2013. Ecological causes of decelerating diversification in carnivoran mammals. Evolution 67:2423–2433.\\
@@
 Smith, N. D. 2012. Body mass and foraging ecology predict evolutionary patterns of skeletal pnematicity in the diverse "waterbird" clade. Evolution 66:1059-1078.\\
-Smith, S. D., C. Ané, and D. A. Baum. 2008. The role of pollinator shifts in the floral diversification of Iochroma (Solanaceae). Evolution 62:793-806.\\
+Smith, S. D., C. An\'{e}, and D. A. Baum. 2008. The role of pollinator shifts in the floral diversification of Iochroma (Solanaceae). Evolution 62:793-806.\\
 Sookias, R.
@@
 Hansen, K.
-Østbye, T.
+\o{O}stbye, T.
 Klepaker, A.
@@
 A.
-Vøllestad.
+V\o{o}llestad.
 2013.
@@
 Wiens, J. J., R. A. Pyron, and D. S. Moen. 2011. Phylogenetic origins of local-scale diversity patterns and the causes of Amazonian megadiversity. Ecology Letters 14:643-652.
-
-  \subsection{Bayesian simulations} % keep? or move it all to supp mat?
-    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
-    The choice of priors for OU models has not been fully explored. 
-    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
-    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
-    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
-S2-S5).
-% add links to this
-    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
-    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
-    Multiple independent chains were run for each analysis to ensure convergence was reached. 
-    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
-    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
-    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
-    We treated Bayes factors $>$ 2 as evidence favouring the OU model.

### 2015-04-21 Manuscript/ohyou.tex
First pass at edits to ms - more to do. Corrected simulation table to d/b

@@
   This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
-  The model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait mean was recast as a fitness optimum on an adaptive landscape. 
+  The model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
   The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see section X).
@@
   In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
-  We use simulations to demonstrate the inherent bias in estimating the $\alpha$ parameter, discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
-  We also show that very small amounts of measurement error in data can have profound effects on the performance of models. 
+  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima).
+We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
   Many of our findings are applicable to other models of evolution (see section X) 
@@
   , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to mention these problems, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}.
+  We are not the first to mention these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
 
@@
 As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche.
-  Under this formulation, $\alpha$ can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optima;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
 
@@
   OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
-  The latter represents evolution under multiple selective regimes, and may be more biologically realistic. 
+  The latter represents evolution under multiple selective regimes, and may be more biologically realistic for many datasets. 
   OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
@@
   The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
-  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate, but our general conclusions also apply to multiple optima OU models.
+  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate.
+For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.
+  %Add citation to Cecile's paper here
 
@@
 
-\section{Known statistical issues with the OU model}
-% Problems with the OU model and where they are reported in the literature?
-% Could rename this section also or move it. Whatever makes most sense.
-
-    \begin{enumerate}
-      \item $\alpha$ has a hard bound at zero but no upper bound, meaning that the Maximum Likelihood profile is asymmetric.
-As values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small.
-This suggests that estimates of $\alpha$ are likely to be biased
-      % make sure we really mean biased here
-      , and that the asymptotic assumptions of Likelihood ratio tests on $\alpha$ are likely to be violated.
-      \item blah
-      \item measurement errors?
-    \end{enumerate}
+\section{Performance of the OU model}
 
@@
 % It's all a bit drafty at present so feel free to edit heavily - you've a better idea than me about the results anyway
-\section{Performance of OU models using Maximum Likelihood simulations}
   To explore some issues with the OU model in more detail, we ran a number of simulations designed to mirror the use of OU models in the literature.
@@
     We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. 
-    Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently \textgreater 1. 
+    Second, for moderate amounts of error (5-10\%), estimates of $\alpha$ are consistently \textgreater 1. 
     Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.
@@
   \item \textbf{Measurement error increases Type I error rates.} 
-  Our results indicate that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
+  Our results show that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
   The effects of measurement error become more severe with increasing tree size.
@@
   These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution. 
-  This is particularly important for inference of phylogenetic niche conservatism. 
+  For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
   Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
-  A more parsimonious explanation would be that there is measurement error in the data.
-  This seems particularly likely for climatic niche traits that are often averaged across an entire species range. 
-  Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.
+  A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.
 
@@
 
-\subsection{Potential solutions and recommendations}
-  Although there are limitations with the OU model, there are also several ways researchers can get around these problems as follows:
+\subsection{Problems, potential solutions and recommendations}
+  Although there are clearly limitations with the OU model, at least some can be dealt with through judicious use of the data, additional analyses and cautious interpretation of model parameters.
+We discuss the main issues that we have identified and highlight possible solutions below.
+Broadly, the recommendations can be divided into approaches to identifying the best fitting model and approaches to interpret model parameters.
 
   \begin{enumerate}[(i)]
-    \item \textbf{Use at least 200 species.} 
-    If you really want to be confident that the OU model is favoured over the Brownian model, you need to fit the models to more than 200 species. 
-    Note that this does not solve the issue of measurement error. in the data.
+    \item \textbf{Tree size} 
+    OU models should not be applied to small trees.
+Our simulations indicate that trees with >200 tips are necessary to obtain acceptable type I error rates.
+However, we are cautious with recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size.
+We note also that large trees are particularly susceptible to issues arising from unaccounted measurement error in the data.
     \item \textbf{Simulate fitted models and compare these to your empirical results.}
@@
     \item \textbf{Consider plausible alternative hypotheses.}
-    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$parameter when an OU model is favoured. 
-    We explain this in detail below.
+    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured.
+Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e.
+very close to Brownian.
+  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
+  Thus this does not reflect any kind of OU process underlying the data. The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX.
+
+  % Rob's old figure 2 back in here with a fairly lengthy figure legend (mainly so we don't need it in the main text)
+
+ To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
+  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.).
+Similarly, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
+
+  The $\alpha$ parameter ranges from 0 to infinity and recognising small or large values may not be intuitive.
+$\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+
+      \begin{equation}
+        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
+        \label{equation:halflife}
+      \end{equation}
+
+  \noindent
+  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
+  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
+  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
+  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process. 
+  As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would clearly lead to similarly biases in the phylogenetic half-life.
+
+% add in ives and garland advice on what is big and what is small here? Or earlier?
+
     \item \textbf{Account for measurement error.} % details
     % This is from the original Discussion
+
+    % I dislike referring to Rob's unpublished work. I think here we need.
+
     %In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
@@
     These implementations should not be used with fossil data.
+
     \item \textbf{Use a Bayesian approach.} See below. %Or cite our preprint?
 \end{enumerate}
-
-  Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
-  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
-  Thus this does not reflect any kind of OU process underlying the data.
-
-  To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
-  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.).
-
-  $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
-
-      \begin{equation}
-        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
-        \label{equation:halflife}
-      \end{equation}
-
-  \noindent
-  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
-  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
-  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
-  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process.
-
-% add in ives and garland advice on what is big and what is small here? Or earlier?
 

### 2015-04-29 Manuscript/ohyou.tex
Updates and destruction of main text

@@
 
+  The majority of PCMs use an explicit evolutionary model to characterize trait evolution (Freckleton et al.
+2011).
 Most model-based methods for characterizing trait evolution are based on the Brownian constant variance model \citep[for exceptions see][]{price1997correlated,harvey2000comparative,freckleton2006detecting}.
@@
   One of the most commonly used Brownian-like models is the Ornstein Uhlenbeck (OU) model. 
-  This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
-  The model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
-  The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see section X).
-  % Add section no here
-  Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
+  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
+  The process operating in comparative data is analogous to but distinct from stabilizing selection.
+The phylogenetic OU model is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing} that is shared across a clade or subset of species.
+  Although widely used, the properties of the OU model, and other direct extensions of the Brownian model, are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
 
-% This paragraph may need modifying once we finish
-  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
-  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima).
-We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
-  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
-  Many of our findings are applicable to other models of evolution (see section X) 
-  % Add section no here
-  , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to mention these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
-
-\section{The Ornstein Uhlenbeck (OU) model}
-
-  According to the Brownian model \citep{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
-
-      \begin{equation}
-        dX(t) = \sigma dW(t)
-      \end{equation}
-
-    \noindent
-    where $W(t)$ is drawn at random from a normal distribution with mean $0$ and variance $\sigma^2$. 
-    The model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant. 
-    Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.
-    The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species. 
-    This means that close relatives will be more similar in their trait values than more distant relatives. It also means that variance in the trait will increase (linearly) in proportion to time. 
-    The model has two parameters, the Brownian rate parameter, $\sigma^2$ and the state of the root at time zero, $X(0)$.
-
-   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
-    The model has the following form:
-
-      \begin{equation}
-        dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
-      \end{equation}
-
-    \noindent
-    Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
-    The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
-    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
-    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see the Appendix.
-
-    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
-    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
-
-  % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
-  \section{Uses of OU models}
 The popularity of the OU model has grown extensively in recent years (Fig.
@@
 
+\section{The Ornstein Uhlenbeck (OU) model}
+
+  According to the Brownian model \citep{cavalli1967,felsenstein1973maximum}, a trait X evolves at random at a rate $\sigma$:
+
+      \begin{equation}
+        dX(t) = \sigma dW(t)
+      \end{equation}
+
+    \noindent
+    where $W(t)$ is drawn at random from a normal distribution with mean $0$ and variance $\sigma^2$. 
+    The model assumes that there is no overall drift in the direction of evolution (hence the expectation of $W(t)$ is zero) and that the rate of evolution is constant. 
+    Because the direction of change in trait values at each step is random, Brownian motion is often described as a ``random walk''.
+    The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species. 
+    This means that close relatives will be more similar in their trait values than more distant relatives. It also means that variance in the trait will increase (linearly) in proportion to time. 
+    The model has two parameters, the Brownian rate parameter, $\sigma^2$ and the state of the root at time zero, $X(0)$.
+
+   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
+    The model has the following form:
+
+      \begin{equation}
+        dX(t) = - \alpha (X(t) - \mu) + \sigma dW(t)
+      \end{equation}
+
+    \noindent
+    Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
+    The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
+    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
+    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see the Appendix.
+
+    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
+    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
+
+% This paragraph may need modifying once we finish
+  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
+  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima).
+We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
+  Many of our findings are applicable to other models of evolution (see section X) 
+  % Add section no here
+  , but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
+  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
+
 \section{Performance of the OU model}
@@
     This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model \citep{}. % Add citations
-    We then estimated the rejection rate for each set of simulations, i.e., the proportion of times OU models are favoured relative to a Brownian motion model, even though the data were simulated under Brownian model.
+    We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
     This gives us a measure of Type I error rate for the OU model.
@@
 
-  \subsection{Simulating phylogenies and data with error}
     We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
@@
 
-  \subsection{Effects of measurement error on OU model performance}
     We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges.
-    Our expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.
+    Our expectation is that the OU model should fit the data better than the Brownian model because the $\alpha$ parameter should account for much of the error.
+In this case, rejection of the Brownian model does not represent type I error because it would be quite correct to reject the Brownian.
+However, the reason for the better fit would be entirely unrelated to any macroevolutionary process.
 
@@
 
-\section{Prospects: limitations, recommendations and potential solutions?}
   \subsection{Limitations of the OU model}
@@
   The effects of measurement error become more severe with increasing tree size.
+
+  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution.
 \end{enumerate}
 
-  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution. 
-  For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
-  Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
-  A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.
+  % As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here. 
+  % For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
+ % Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
+ % A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.
 
@@
 
-\subsection{Problems, potential solutions and recommendations}
-  Although there are clearly limitations with the OU model, at least some can be dealt with through judicious use of the data, additional analyses and cautious interpretation of model parameters.
-We discuss the main issues that we have identified and highlight possible solutions below.
-Broadly, the recommendations can be divided into approaches to identifying the best fitting model and approaches to interpret model parameters.
+\section{Prospects: limitations, recommendations and potential solutions}
 
+  We have highlighted several issues with the OU model.
+These can be broadly divided into two classes: (i) fitting the model and (ii) interpreting model parameters.
+These issues can be at least partially addressed with additional or alternative analytical approaches.
+Below we make recommendations for best practice in fitting model and approaches to interpret model parameters.
+We highlight areas of future research that may be important in alleviating outstanding issues.
+
+  \subsection{Recommendations for model fitting}
     \begin{enumerate}[(i)]
-    \item \textbf{Tree size} 
-    OU models should not be applied to small trees.
+    \item \textbf{Simulate under the null model} OU models should not be applied to small trees.
 Our simulations indicate that trees with >200 tips are necessary to obtain acceptable type I error rates.
@@
 We note also that large trees are particularly susceptible to issues arising from unaccounted measurement error in the data.
-    \item \textbf{Simulate fitted models and compare these to your empirical results.}
+Instead, we suggest using simulations to asses the model fit.
+Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for likelihood ratio tests.
+%We could potentially add a script/workflow for this
+
+    \item \textbf{Consider Bayesian approaches} An alternative but little explored approach is to use Bayesian methods.
+Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree and geiger (the Single Stationary Peak model in fitContinuousMCMC) and stand alone software including BayesTraits.
+%citations needed
+
+	We repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits.
+We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}.
+The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1).
+Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded).
+We treated Bayes factors $>$ 2 as evidence favouring the OU model.
+We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
+Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample.
+Multiple independent chains were run for each analysis to ensure convergence was reached.
+An important challenge for Bayesian approaches is selection of appropriate priors.
+We used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
+For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$.
+    Table XX shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
+    The Bayesian approach results in highly conservative rejection rates regardless for all tree shapes in the absence of measurement error.
+While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power.
+The Bayesian approach also appears to more readily handle low levels of measurement error.
+With 1\% measurement error the Bayesian approach retains acceptable rejection rates (<0.05) for trees of up to 150 tips.
+However, more error or larger trees result in the frequent rejection of the Brownian model.
+As noted above, this is not an issue of type I and it is entirely correct that the Brownian model is rejected.
+In such cases emphasis should shift to how the $\alpha$ parameter is interpreted.
+At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation.
+    \end{enumerate}
+
+  \subsection{Recommendations for interpreting $\alpha$}    
+    \begin{enumerate}[(i)]
+    \item \textbf{Consider plausible alternative hypotheses} The effects of measurement in particular suggest that $\alpha$ must always be interpreted with caution.
+Because the OU model was proposed with evolutionary process in mind these alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ can be overlooked.
+Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured.
+Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e.
+very close to Brownian.
+  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
+  Thus this does not reflect any kind of OU process underlying the data.
+The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX.
+A the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
+
     A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values.
@@
 Any given tree will therefore generate unique parameter estimates.
-Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}.
 Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.
-    \item \textbf{Consider plausible alternative hypotheses.}
-    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured.
-Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e.
-very close to Brownian.
-  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
-  Thus this does not reflect any kind of OU process underlying the data. The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX.
 
-  % Rob's old figure 2 back in here with a fairly lengthy figure legend (mainly so we don't need it in the main text)
-
- To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
-  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.).
-Similarly, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
-
+	\item \textbf{Calculate the phylogenetic half-life} 
   The $\alpha$ parameter ranges from 0 to infinity and recognising small or large values may not be intuitive.
@@
   \noindent
-  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
   If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
@@
 
-    \item \textbf{Account for measurement error.} % details
-    % This is from the original Discussion
-
-    % I dislike referring to Rob's unpublished work. I think here we need.
-
-    %In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
-It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
-However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
-This is because the variances will themselves be estimates and subject to error.
-Treating such variances as error-free can potentially create biases (Freckleton unpub).
-If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance.
-However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
     \item \textbf{If possible, include ancillary data.}
@@
     This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
+    % is this now in Graham's MEE correction?
     Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
     These implementations should not be used with fossil data.
-
-    \item \textbf{Use a Bayesian approach.} See below. %Or cite our preprint?
    \end{enumerate}
 
-\subsection{Bayesian simulations} 
-% (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
-    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
-    The choice of priors for OU models has not been fully explored. 
-    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20.
-    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
-    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs.
-S2-S5).
-% add links to this
-    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
-    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
-    Multiple independent chains were run for each analysis to ensure convergence was reached. 
-    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
-    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
-    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
-    We treated Bayes factors $>$ 2 as evidence favouring the OU model.
+\section{Conclusions and outstanding issues}
 
-\section{Just an OU problem?}
+A recurring theme from our simulations is that interpretation of $\alpha$ is not straightforward in the presence of measurement error.
+It would clearly be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
+Several approaches for accounting for error have been proposed %check references.
+It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
+However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
+This is because the variances will themselves be estimates and subject to error.
+Treating such variances as error-free can potentially create biases (Freckleton unpub).
+An additional challenge common to (especially large) comparative data sets is that measurement error or intraspecific variation is either not available or is inconsistently reported.
+One approach if the variances are unknown would be to simultaneously fit Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} to estimate the non-Brownian component of the trait variance.
+However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
+
+These issues are not limited to OU models.
+Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being mislead by  measurement error.
+The fundamental problem is that rejection of the Brownian model in favour of another does not necessarily say anything about process.
+This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}.
+We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.
+
 %Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
@@
 
-\section{Concluding comments/Conclusions}
 % Conclusion from original

### 2015-04-29 Manuscript/ohyou_SuppInfo.tex
Renamed supp mat file

--- 2015-04-21 Manuscript/ohyou_suppmat.tex
@@
-\section{Supplementary Material}
+\documentclass[a4paper,12pt]{article}
+
+\usepackage[osf]{mathpazo} % palatino
+\usepackage{ms}            % load the template
+\usepackage[round]{natbib} % author-year citations
+\usepackage{graphicx}
+\usepackage{parskip} 
+\usepackage{enumerate} 
+\usepackage{longtable}
+\usepackage{pdflscape}
+\usepackage{fullpage}
+\usepackage{hyperref}
+\usepackage[nomarkers, nolists]{endfloat}  % To push floats to the end   
+\pagenumbering{arabic}
+
+\begin{document}
+
+\raggedright
+\doublespacing
+\setlength{\parindent}{1cm}
+
+\section{Supporting Information}
 \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
 
-  \subsection{Literature review}
-    To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 \citep[when the R package ouch was released;][]{Butler:2004aa} and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''.
+  \subsection{Literature review methods}
+    To get an overview of the use of OU models in ecology, evolution and palaeontology, we used Google Scholar (accessed 13th March 2015) to locate papers published between 2005 (when the R package ouch was released; Butler and King 2004) and 2014 that contained the terms ``Ornstein Uhlenbeck'' and either ``'ecology'', or ``evolution'' and ``biology'' (the ``biology'' term was added to omit physics papers which also use the term ``evolution''), or ``paleo/palaeo''.
     We also recorded the total number of papers containing the terms ``ecology'', or ``evolution'' and ``biology'', or ``paleo/palaeo'', published between 2005 and 2014 and plotted the number of OU papers published each year as a proportion of the total number of papers published (Fig.
-\ref{figure.literature}).
+1 in the main text).
 
@@
     Where papers had multiple study groups, statistics/R packages or reasons for fitting OU models we counted them in each relevant category. 
-    We summarise these results in Figures 1 and 2 and Table 1, and the full dataset is available below along with the full list of references.
+    We summarise these results in Figures 1 and 2 and Table 1 in the main text, and the full dataset is available in Table \ref{table:litreview} along with the full list of references.
 
-    In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. 1). 
-    Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Figure 2 and Table S1). 
-    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the ``best'' model of trait evolution or to control for phylogenetic non-independence (Table S1).
+  \subsection{Literature review results}
+    In total, our literature search found 3720 papers published between 2005 and 2014, and the number has increased substantially since 2005 (Fig. 1 in main text). 
+    Most papers fit OU models to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58, Figure 2 in main text and Table \ref{table:litreview}). 
+    The majority of papers fit OU models using R packages, particularly GEIGER and, although other uses are becoming more common, most papers use OU models in an effort to discern the ``best'' model of trait evolution or to control for phylogenetic non-independence (Table \ref{table:litreview}).
 
@@
 \begin{center}
+
+\LTcapwidth=25cm
 \begin{longtable}{p{6cm}llllll}
@@
 \bigskip        
-\hline
+%\hline
 \textbf{Paper}  &   \textbf{Year}    &   \textbf{Journal} &   \textbf{Taxon}   &   \textbf{Ntips}   &   \textbf{Use in paper}   &   \textbf{Stats/R package} \\
@@
 \subsection{Literature review references}
-
+\label{section:refs} 
 \indent
@@
 Wiens, J. J., R. A. Pyron, and D. S. Moen. 2011. Phylogenetic origins of local-scale diversity patterns and the causes of Amazonian megadiversity. Ecology Letters 14:643-652.
+
+\end{document}

### 2015-04-29 Analyses/OhYouFigures.R
Changed figures as journal wants PDF

@@ -9,7 +9,7 @@ ds$evo <- (ds$evolution / ds$evolution.total) * 100
 ds$pal <- (ds$paleo / ds$paleo.total) * 100
 
 # Plot figure 1
-png("Manuscript/Figures/PapersThruTime.png")
+pdf("Manuscript/Figures/OhYou_Figure1.pdf")
 par(bty = "l")
 plot(evo ~ Year, data = ds, las = 1, pch = 16, cex = 1.2, cex.lab = 1.2, cex.axis = 1.2,
      ylab = "OU publications (% of total papers)", xlab = "Year published", ylim = c(0, 0.7))
@@ -34,7 +34,7 @@ ds2$General.use <- ds2$General.use[, drop = TRUE]
 summary.table <- tapply(ds2$Paper, list(ds2$General.use, ds2$Noptima), length)
 
 # Figure 3
-png("Manuscript/Figures/NTaxa.png")
+pdf("Manuscript/Figures/OhYou_Figure2.pdf")
   par(bty = "l")
   hist(ds1$Ntips[which(ds1$Ntips <= 1000)], las = 1, cex.lab = 1.2, cex.axis = 1.2,
        ylab = "Number of papers", xlab = "Number of taxa", ylim = c(0, 80), main = "")

### 2015-04-29 Manuscript/ohyou.tex
Small edits and change to figures

@@
   These approaches have been used, for example, to infer potential rates of species responses to climate change \citep{Quintero:2013aa}, test the role of ecological niche as a driver of morphological evolution \citep{pienaar2013macroevolution}, and test for constraints in adaptive radiations \citep{blackburn2013adaptive}.
-  %Unfortunately, although widely used, the properties of some phylogenetic comparative methods are poorly understood leading to the potential for inappropriate use and misinterpretation of results.
 
@@
 
+% NC: I've moved this up as I think the "uses" section forms part of the intro to the OU model. I'm not convinced uses should come before the model itself but I don't think it matters at this point.
+% NC:  This paragraph may need modifying once we finish
+  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
+  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima).
+We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
+  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
+  Many of our findings are applicable to other models of evolution (see section X) 
+  % Add section no here
+  , but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
+  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
+
+\section{Uses of the OU model}
   The popularity of the OU model has grown extensively in recent years (Fig.
-\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material).
-% Need to add ref to supp mat
+\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supporting Information).
   This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
@@
   It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading.
-As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche.
-  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optima;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
+As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optimum for that niche.
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
 
@@
       \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure1.pdf}
       \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supplementary Material for details.
+See Supporting Information for details.
 %linkto supp mat
@@
 
-  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material). 
+  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supporting Information). 
   Multiple models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
@@
   OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
-
-  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supplementary Material).  
+  % add citation
+  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supporting Information).  
   Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
@@
 
-  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supplementary Material) therefore, we focus on this use of the OU model in our simulations below.
+  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supporting Information) therefore, we focus on this use of the OU model in our simulations below.
   The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
-  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supplementary Material) and easier to simulate.
+  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.
 For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.
-  %Add citation to Cecile's paper here
 
@@
       \begin{center}
-        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supplementary Material for details.}
+        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supporting information for details.}
         \bigskip  
@@
 
-\section{The Ornstein Uhlenbeck (OU) model}
+\section{OU model outline}
 
@@
 
-% This paragraph may need modifying once we finish
-  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
-  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima).
-We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
-  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
-  Many of our findings are applicable to other models of evolution (see section X) 
-  % Add section no here
-  , but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
-  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
-
 \section{Performance of the OU model}
 
-% I've separated this out a bit to try and make the progression more logical
-% It's all a bit drafty at present so feel free to edit heavily - you've a better idea than me about the results anyway
   To explore some issues with the OU model in more detail, we ran a number of simulations designed to mirror the use of OU models in the literature.
@@
     To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
-    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model \citep{}. % Add citations
+    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model.
     We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
@@
 
-% I ended up just LaTeXing the tables for my own amusement
       \begin{table}[!htbp]
@@
     This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
-    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supplemental Material). % link to supp mat
+    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supporting Information).
 
@@
       \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/NTaxa.png}
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
       \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
 Two studies with \textgreater 3000 taxa have been omitted for clarity.
-See Supplementary Material for details.
-%linkto supp mat
+See Supporting Information for details.
       }
@@
 \section{Effects of measurement error on OU model performance}
-  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{}.
-% add citation to Matt's model adequacy stuff
+  Measurement error can strongly effect the results of comparative analyses \citep{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep[see][]{pennell2015model}.
   Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
@@
   We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
-    % should we say "as above" here or be specific i.e. copy paste from above?
   Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees.
@@
 
-    We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges.
+  We compared the fit of a Brownian model to that of an OU model using Likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges.
   Our expectation is that the OU model should fit the data better than the Brownian model because the $\alpha$ parameter should account for much of the error.
-In this case, rejection of the Brownian model does not represent type I error because it would be quite correct to reject the Brownian.
+In this case, rejection of the Brownian model does not represent Type I error because it would be quite correct to reject the Brownian.
 However, the reason for the better fit would be entirely unrelated to any macroevolutionary process.
@@
 With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species.
-    This is important for the interpretation of the OU model. 
-    We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. 
+  This is important for the interpretation of the OU model; we cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. 
   Second, for moderate amounts of error (5-10\%), estimates of $\alpha$ are consistently \textgreater 1. 
@@
   \subsection{Limitations of the OU model}
-  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways.
-  Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. 
+  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to be aware that such models are statistically complex and may behave in unexpected ways.
+  Transformations of the variance-covariance matrix in the Brownian model \citep[e.g.
+$\lambda$;][]{Pagel:1997aa} are an attractive and computationally simple way to modify the basic model to include evolutionary processes.
   But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. 
-  The results of our simulations illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supplementary Table S1).
+  The results of our simulations illustrate that such problems can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supporting Information Table S1).
 
@@
   The effects of measurement error become more severe with increasing tree size.
-
-  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution.
+  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution, particularly where there is likely to be intraspecific variation or measurement error in the data.
 \end{enumerate}
 
-  % As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here. 
+  % GT:  As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here.
+
  % For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
@@
  % A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.
-
-% Not sure this is needed anywhere
-%In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process.
-This is because according to the OU process traits move back towards the optimum as their values become particularly large or small.
-Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig.
-\ref{figure.traitsim}).
-Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.
 
@@
     \begin{enumerate}[(i)]
-    \item \textbf{Simulate under the null model} OU models should not be applied to small trees.
-Our simulations indicate that trees with >200 tips are necessary to obtain acceptable type I error rates.
+      \item \textbf{Simulate under the null model.}\\
+      OU models should not be applied to small trees. 
+      Our simulations indicate that trees with \textgreater 200 tips are necessary to obtain acceptable Type I error rates. 
       However, we are cautious with recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size. 
@@
       Instead, we suggest using simulations to asses the model fit.
-Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for likelihood ratio tests.
+Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for Likelihood ratio tests.
 %We could potentially add a script/workflow for this
+      \citealp{boettiger2012your} discuss this at length and we recommend reading their paper for a full overview of this approach.
+      % NC: Though their method is hard to apply it does cover all the points we discuss here so we chould make it clearer (as requested by Syst Biol reviewers)
 
-    \item \textbf{Consider Bayesian approaches} An alternative but little explored approach is to use Bayesian methods.
-Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree and geiger (the Single Stationary Peak model in fitContinuousMCMC) and stand alone software including BayesTraits.
-%citations needed
+      \item \textbf{Consider Bayesian approaches.}\\
+      An alternative but less explored approach is to use Bayesian methods. 
+      Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree \citep{FitzJohn:2012aa} and GEIGER \citep[the Single Stationary Peak model in \texttt{fitContinuousMCMC};][]{Harmon:2008aa} and stand alone software including BayesTraits \citep{pagel2013bayestraits}.
 
-	We repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits.
+	    We (AM and CV) repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits \citep{pagel2013bayestraits}. 
       We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}. 
       The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
-Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded).
-We treated Bayes factors $>$ 2 as evidence favouring the OU model.
+      Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). W
+      e treated Bayes factors $>$ 2 as evidence favouring the OU model.
 We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
@@
       For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$.
+
       Table XX shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
@@
       The Bayesian approach also appears to more readily handle low levels of measurement error. 
-With 1\% measurement error the Bayesian approach retains acceptable rejection rates (<0.05) for trees of up to 150 tips.
+      With 1\% measurement error the Bayesian approach retains acceptable rejection rates (\textless 0.05) for trees of up to 150 tips. 
       However, more error or larger trees result in the frequent rejection of the Brownian model. 
-As noted above, this is not an issue of type I and it is entirely correct that the Brownian model is rejected.
+      As noted above, this is not an issue of Type I and it is entirely correct that the Brownian model is rejected. 
       In such cases emphasis should shift to how the $\alpha$ parameter is interpreted. 
       At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation. 
+      We also recommend reading \citet{pennell2015model} as their paper presents very similar conclusions with respect to Bayesian methods.
     \end{enumerate}
@@
     \begin{enumerate}[(i)]
-    \item \textbf{Consider plausible alternative hypotheses} The effects of measurement in particular suggest that $\alpha$ must always be interpreted with caution.
+    \item \textbf{Consider plausible alternative hypotheses.}\\
+    The effects of measurement error in particular suggest that $\alpha$ must always be interpreted with caution. 
     Because the OU model was proposed with evolutionary process in mind these alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ can be overlooked.
     Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
-Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e.
-very close to Brownian.
+    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
     In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
@@
     The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX. 
-A the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
+    At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
 
@@
 
-	\item \textbf{Calculate the phylogenetic half-life} 
-  The $\alpha$ parameter ranges from 0 to infinity and recognising small or large values may not be intuitive.
+	  \item \textbf{Calculate the phylogenetic half-life.} \\
+    The $\alpha$ parameter ranges from zero to infinity and recognising small or large values may not be intuitive. 
     $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
@@
 
-    \item \textbf{If possible, include ancillary data.}
+    \item \textbf{If possible, include ancillary data.}\\
     If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
@@
 \section{Conclusions and outstanding issues}
-
   A recurring theme from our simulations is that interpretation of $\alpha$ is not straightforward in the presence of measurement error.
@@
 
-% Supplementary Material
-\input{ohyou_suppmat}
+\newpage
+\section{Figure legends}
+
+Figure 1:
+
+Figure 2: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+Two studies with \textgreater 3000 taxa have been omitted for clarity.
+See Supporting Information for details.
+
+\newpage
+\section{Tables}
 

### 2015-05-01 R/OUtables.R
Added tables with BayesTraits result from analyses with exp 10 prior on alpha

@@ -99,3 +99,121 @@ write.csv(OUresVarRateTable, "Manuscript/tables/OU_var_rate_noerror.csv")
 
 
 
+####################
+## Bayestraits tables
+OUBayes <- read.csv("Data/MCMC_ExpLam10.csv", stringsAsFactors=FALSE)
+
+
+OUBayesTable <- as.data.frame(matrix(nrow=28, ncol=4))
+colnames(OUBayesTable) <- c("Tree type", "Tree size", "Rejection rate", "Median alpha (95% HPD)")
+
+OUBayesTable[1:7,"Tree type"] <- "b/d = 0"
+OUBayesTable[8:14,"Tree type"] <- "b/d = 0.25"
+OUBayesTable[15:21,"Tree type"] <- "b/d = 0.5"
+OUBayesTable[22:28,"Tree type"] <- "b/d = 0.75"
+
+OUBayesTable[seq(1, 28, 7), "Tree size"] <- 25
+OUBayesTable[seq(2, 28, 7), "Tree size"] <- 50
+OUBayesTable[seq(3, 28, 7), "Tree size"] <- 100
+OUBayesTable[seq(4, 28, 7), "Tree size"] <- 150
+OUBayesTable[seq(5, 28, 7), "Tree size"] <- 200
+OUBayesTable[seq(6, 28, 7), "Tree size"] <- 500
+OUBayesTable[seq(7, 28, 7), "Tree size"] <- 1000
+
+
+nms <- c("yule25", "yule50", "yule100", "yule150", "yule200", "yule500", "yule1000", "bdLow25", "bdLow50", "bdLow100", "bdLow150", "bdLow200", "bdLow500", "bdLow1000", "bdMid25", "bdMid50", "bdMid100", "bdMid150", "bdMid200", "bdMid500", "bdMid1000", "bdHigh25", "bdHigh50", "bdHigh100", "bdHigh150", "bdHigh200", "bdHigh500", "bdHigh1000")
+
+for (i in 1:28) {
+x <- OUBayes[OUBayes[,"Model"] == paste(nms[i]),]
+OUBayesTable[i,"Rejection rate"] <- round(sum(x[,"Exp_L10_2Bfactor"] > 2, na.rm=T)/sum(!is.na(x[,"Exp_L10_2Bfactor"])), 3)
+
+alpha <- quantile(x[,"Exp_L10_Alpha_Mode"], probs=c(0.5, 0.025, 0.975))
+
[…]

### 2015-05-01 Manuscript/ohyou.tex

@@
   We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
-  Many of our findings are applicable to other models of evolution (see section X) 
-  % Add section no here
-  , but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
+  Many of our findings are applicable to other models of evolution, but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
   We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
@@
 
-    \begin{figure}[h]
+    \begin{figure}[!htbp]
       \centering
@@
 See Supporting Information for details.
-%linkto supp mat
       }
@@
 \end{figure}
+% Weird issue with endfloat means these need to have no white space
 
@@
 
-    \begin{table}[h]
+    \begin{table}[!htbp]
       \begin{center}
@@
 \end{table}
+% Weird issue with endfloat means these need to have no white space
 
@@
 \end{table} 
+% Weird issue with endfloat means these need to have no white space
 
@@
 \end{figure} 
+% Weird issue with endfloat means these need to have no white space
 
-    We provide recommendations related to these findings below. % name of section?
+    We provide recommendations related to these findings below.
 
@@
 \end{table} 
+% Weird issue with endfloat means these need to have no white space
     \end{landscape}
@@
 
-Figure 1:
+\noindent
+\textbf{Figure 1}: The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+See Supporting Information for details.
 
-Figure 2: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+\noindent
+\textbf{Figure 2}: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
 Two studies with \textgreater 3000 taxa have been omitted for clarity.

### 2015-05-01 Manuscript/ohyou.tex
Minor edits to final sections

@@
   % GT:  As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here. 
+  % NC: OK happy to leave it out
 
@@
 
-\section{Prospects: limitations, recommendations and potential solutions}
+\section{Recommendations and potential solutions}
 
-  We have highlighted several issues with the OU model.
+  We have highlighted several issues with the OU model above.
 These can be broadly divided into two classes: (i) fitting the model and (ii) interpreting model parameters.
@@
     \begin{enumerate}[(i)]
-      \item \textbf{Simulate under the null model.}\\
+      \item \textbf{Simulate under the null model.}
       OU models should not be applied to small trees. 
       Our simulations indicate that trees with \textgreater 200 tips are necessary to obtain acceptable Type I error rates. 
-      However, we are cautious with recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size. 
+      However, we are cautious about recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size. 
       We note also that large trees are particularly susceptible to issues arising from unaccounted measurement error in the data. 
-      Instead, we suggest using simulations to asses the model fit.
+      Instead, we suggest using simulations to assess the model fit. 
       Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for Likelihood ratio tests.
 %We could potentially add a script/workflow for this
-      \citealp{boettiger2012your} discuss this at length and we recommend reading their paper for a full overview of this approach.
-      % NC: Though their method is hard to apply it does cover all the points we discuss here so we chould make it clearer (as requested by Syst Biol reviewers)
+      \citet{boettiger2012your} discuss this at length and we recommend reading their paper for a full overview of this approach.
+      % NC: Though their method is hard to apply it does cover all the points we discuss here so we chould make it clearer (as requested by Syst Biol reviewers).
+Also I'd forgotten that the new geiger code means their package is deprecated and can't be used anyway.
 
-      \item \textbf{Consider Bayesian approaches.}\\
+      \item \textbf{Consider Bayesian approaches.}
       An alternative but less explored approach is to use Bayesian methods. 
@@
       The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
-      Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). W
-      e treated Bayes factors $>$ 2 as evidence favouring the OU model.
-We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
+      Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). 
+      We treated Bayes factors \textgreater 2 as evidence favouring the OU model. We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
       Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
@@
 
-      Table XX shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
+      Table \ref{tables:bayesian} shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
       The Bayesian approach results in highly conservative rejection rates regardless for all tree shapes in the absence of measurement error. 
@@
       At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation. 
-      We also recommend reading \citet{pennell2015model} as their paper presents very similar conclusions with respect to Bayesian methods.
+      We also recommend reading \citet{pennell2015model} as their paper presents similar conclusions with respect to Bayesian methods.
     \end{enumerate}
+
+    \begin{table}[!htbp]
+        \begin{center}
+        \caption{Legend for Bayesian table.} % need to enter legend here.
+        \bigskip
+        \input{Tables/OU_bd_noerror.tex} % need to input actual table here.
+        \label{tables:bayesian}
+        \end{center}
+\end{table} 
+% Weird issue with endfloat means these need to have no white space
 
@@
     \begin{enumerate}[(i)]
-    \item \textbf{Consider plausible alternative hypotheses.}\\
+    \item \textbf{Consider plausible alternative hypotheses.}
+    Because the OU model was proposed with evolutionary process in mind, alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ are often overlooked.
     The effects of measurement error in particular suggest that $\alpha$ must always be interpreted with caution. 
-    Because the OU model was proposed with evolutionary process in mind these alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ can be overlooked.
     Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
@@
     Thus this does not reflect any kind of OU process underlying the data. 
-    The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX. 
+    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in Figure XX. % NC: Not sure what this figure is...
     At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
 
-    A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. 
+    A good strategy for data exploration would be to simulate data under Brownian and the favoured OU model to generate distributions of parameters under known values. 
     These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. 
@@
     Any given tree will therefore generate unique parameter estimates. 
-    Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.
+    Generating data under the favoured OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.
 
-	  \item \textbf{Calculate the phylogenetic half-life.} \\
-    The $\alpha$ parameter ranges from zero to infinity and recognising small or large values may not be intuitive. 
-    $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
+	  \item \textbf{Calculate the phylogenetic half-life.}
+    The $\alpha$ parameter ranges from zero to infinity, thus recognising ``small'' or ``large'' values may not be intuitive. 
+    $\alpha$ can sometimes be interpreted more easily by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
 
@@
     However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process. 
-    As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would clearly lead to similarly biases in the phylogenetic half-life.
+    As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would lead to similar biases in $t_\frac{1}{2}$.
 
@@
 
-    \item \textbf{If possible, include ancillary data.}\\
-    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
+    \item \textbf{If possible, include ancillary data.}
+    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.
     This is implemented in GEIGER \citep{Harmon:2008aa}.
-    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances (see Appendix \ref{equation:OUcov} depend on both the shared distances between species and the distance of a node to the nearest tip.
-    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
-    % is this now in Graham's MEE correction?
+    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances depend on both the shared distances between species and the distance of a node to the nearest tip (see equation \ref{equation:OUcov} in the Appendix).
+    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like \citep{slater2014correction}.
     Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
@@
 \section{Conclusions and outstanding issues}
+% NC: I'm not so sure about this paragraph.
+It reads like a Freckleton paper and I'm not sure how helpful it is.
+I'd be tempted to just cut it and the subheading, and end with the paragraph on how it applies to all models.
+What do you think?
   A recurring theme from our simulations is that interpretation of $\alpha$ is not straightforward in the presence of measurement error.
-It would clearly be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
-Several approaches for accounting for error have been proposed %check references.
-  It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
+It would clearly be very useful to have estimates of measurement error for all species traits, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
+Several approaches for accounting for error have been proposed.
+%check references.
+  It would be tempting to create a vector of variances that could be combined with V
+  % NC: this is a leftover from the previous version - is V the variance covariance matrix?
+  to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
 However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
-This is because the variances will themselves be estimates and subject to error.
-Treating such variances as error-free can potentially create biases (Freckleton unpub).
+  This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton pers. comm.). 
   An additional challenge common to (especially large) comparative data sets is that measurement error or intraspecific variation is either not available or is inconsistently reported. 
-One approach if the variances are unknown would be to simultaneously fit Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} to estimate the non-Brownian component of the trait variance.
+  One approach if the variances are unknown would be to simultaneously fit Pagel's $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} to estimate the non-Brownian component of the trait variance. 
   However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
 
+% NC: Slightly modified i think this is a good end paragraph. Perhaps with a bit more form the original conclusion below.
   These issues are not limited to OU models. 
   Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being mislead by measurement error. 
-The fundamental problem is that rejection of the Brownian model in favour of another does not necessarily say anything about process.
+  The fundamental problem is that rejection of the Brownian model in favour of another model does not necessarily say anything about process. 
   This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}.
@@
   All simulated phylogenies, data, and R code are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
-  However, although our R code is provided, it's messy and difficult to use. 
+  However, although our R code is provided, it is messy and difficult to use. 
   We also do not provide an automated way of reproducing the data collection for our literature review.
-  Nor do we use tools like Travis CI, Docker, or packrat etc.
-% Add stuff about BayesTraits here. 
+  Nor do we use tools like Travis CI (travis-ci.org), Docker (www.docker.com), or packrat \citep{Ushey:2015aa} etc. to increase the reproducbility of our analyses.
+  Additionally, we use BayesTraits to run our Bayesian analyses. 
+  Although free to download, this software is not an Open Source, which has implications for reproducibility.
+  % NC: I'm not sure what else to write here without pissing off Chris etc. Is it possible to put their scripts on GitHub? Can you ask? 
   We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 

### 2015-05-01 Manuscript/ohyou.tex
Added Bayesian tables

@@
       While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power. 
-      The Bayesian approach also appears to more readily handle low levels of measurement error. 
+      The Bayesian approach also appears to more readily handle low levels of measurement error (Table \ref{tables:bayesian_error}). 
       With 1\% measurement error the Bayesian approach retains acceptable rejection rates (\textless 0.05) for trees of up to 150 tips. 
@@
         \begin{center}
-        \caption{Legend for Bayesian table.} % need to enter legend here.
+        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees using Bayesian methods.
+Tree type refers to the extinction fraction for the birth-death trees.
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.
+} % NC: might need to modify legend here.
         \bigskip
-        \input{Tables/OU_bd_noerror.tex} % need to input actual table here.
+        \input{Tables/OU_bd_noerror_BayesTraits.tex} 
         \label{tables:bayesian}
@@
 % Weird issue with endfloat means these need to have no white space
+
+    \begin{landscape}
+      \begin{table}[!htbp]
+        \begin{center}
+        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.) using Bayesian methods.
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.} %NC: Might need to modify legend
+        \bigskip
+        \input{Tables/OU_yule_error_BayesTraits.tex}
+        \label{tables:bayesian_error}
+        \end{center}
+\end{table} 
+% Weird issue with endfloat means these need to have no white space
+    \end{landscape}
 

### 2015-05-11 Manuscript/ohyou.tex
Added abstract

@@
 
-\title{A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies}
+\title{A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies}
 \author{
@@
 
-\runninghead{A cautionary note on macroevolutionary models}
-\keywords{Bayesian method - comparative methods - delta - kappa - lambda - OU - phylogeny - stabilizing selection - selection} %meant to be after abstract
+\runninghead{A cautionary note on Ornstein Uhlenbeck models}
+\keywords{OU - comparative methods - phylogeny - stabilising selection - macroevolutionary models} %meant to be after abstract
 
@@
 \section{Abstract 100-200}
-  %Phylogenetic comparative methods are increasingly used to give new insight into variation, causes and consequences of trait variation among species.
+  Phylogenetic comparative methods are increasingly used to give new insights into the causes and consequences of trait variation among species. 
   The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
-However, the parameters of these models have been hypothesised to be biased and only asymptotically behave in a statistically predictable way as datasets become large.
-This does not seem to be widely appreciated.
-We show that a commonly used model in evolutionary biology (the Ornstein-Uhlenbeck model) is biased over a wide range of conditions.
-Many studies fitting this model use datasets that are small and prone to substantial biases.
-Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model.
+  However, the properties of these models are often poorly understood which can lead to the misinterpretation of results.
+  Here we focus on one of these models - the Ornstein-Uhlenbeck (OU) model. 
+  We introduce the model, its general properties and some issues with its use in ecology, evolution and palaeontology.
+  Using simulations, we demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$, especially where datasets are small. 
+  We also show that very small amounts of error in data can have profound effects on the performance of OU models.
+  Finally, we make recommendations for best practice in fitting OU models in phylogenetic comparative analyses, and for interpreting the $\alpha$ parameter.
 

### 2015-05-11 Manuscript/ohyou.tex
Removed abstract word count

@@
 
-\section{Abstract 100-200}
+\section{Abstract}
   Phylogenetic comparative methods are increasingly used to give new insights into the causes and consequences of trait variation among species. 

### 2015-05-11 Manuscript/ohyou_wordconvert.tex
Merged GHTs edits

--- 2015-05-11 Manuscript/ohyou.tex
@@
 \usepackage{hyperref}
-%\usepackage[nomarkers, nolists]{endfloat}  % To push floats to the end   
+\usepackage[nomarkers, nolists]{endfloat}  % To push floats to the end   
 \pagenumbering{arabic}
@@
 
-    \begin{figure}[!htbp]
-      \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure1.pdf}
-      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+    %\begin{figure}[!htbp]
+     % \centering
+      %\includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure1.pdf}
+      %\caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
 See Supporting Information for details.
-      }
-      \label{figure.literature}
-\end{figure}
+      %}
+      %\label{figure.literature}
+%\end{figure}
 % Weird issue with endfloat means these need to have no white space
@@
 
-    \begin{figure}[!htbp]
-      \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
-      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+ %   \begin{figure}[!htbp]
+  %    \centering
+   %   \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
+    %  \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
 Two studies with \textgreater 3000 taxa have been omitted for clarity.
 See Supporting Information for details.
-      }
-      \label{figure:ntaxa}
-\end{figure} 
+     % }
+      %\label{figure:ntaxa}
+%\end{figure} 
 % Weird issue with endfloat means these need to have no white space
@@
 \textbf{Figure 1}: The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supporting Information for details.
+See Supporting Information for details.\\
 

### 2015-05-11 Manuscript/ohyou.tex
Fixed Hoehna citation

@@
 \vfill
-%\paragraph{Word-count:} X words
 
 \runninghead{A cautionary note on Ornstein Uhlenbeck models}
-\keywords{OU - comparative methods - phylogeny - stabilising selection - macroevolutionary models} %meant to be after abstract
+\keywords{OU - comparative methods - phylogeny - stabilising selection - macroevolutionary models}
 
@@
 \section{Abstract}
-  Phylogenetic comparative methods are increasingly used to give new insights into the causes and consequences of trait variation among species. 
+  Phylogenetic comparative methods are increasingly used to give new insights into the dynamics of trait evolution in deep time. 
   The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
   However, the properties of these models are often poorly understood which can lead to the misinterpretation of results.
-  Here we focus on one of these models - the Ornstein-Uhlenbeck (OU) model. 
-  We introduce the model, its general properties and some issues with its use in ecology, evolution and palaeontology.
-  Using simulations, we demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$, especially where datasets are small. 
-  We also show that very small amounts of error in data can have profound effects on the performance of OU models.
-  Finally, we make recommendations for best practice in fitting OU models in phylogenetic comparative analyses, and for interpreting the $\alpha$ parameter.
+  Here we focus on one of these models - the Ornstein Uhlenbeck (OU) model.
+  We show that the OU model is frequently incorrectly favoured over simpler models, and that many studies fitting this model use datasets that are small and prone to Type I error.   
+  We also show that very small amounts of error in datasets can have profound effects on the performance of OU models.
+  Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model. 
+  We conclude by making recommendations for best practice in fitting OU models in phylogenetic comparative analyses, and for interpreting the parameters of the OU model.
 
@@
   One of the most commonly used Brownian-like models is the Ornstein Uhlenbeck (OU) model. 
-  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
-  The process operating in comparative data is analogous to but distinct from stabilizing selection.
+  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilising selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
+  The process operating in comparative data is analogous to but distinct from stabilising selection.
 The phylogenetic OU model is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing} that is shared across a clade or subset of species.
@@
 
-% NC: I've moved this up as I think the "uses" section forms part of the intro to the OU model. I'm not convinced uses should come before the model itself but I don't think it matters at this point.
-% NC:  This paragraph may need modifying once we finish
   In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
@@
   For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
-  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading.
+  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilising selection'', this is inaccurate and misleading.
 As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optimum for that niche.
-  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive} but not as an estimate of stabilising selection in the population genetics sense.
 
@@
       \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure1.pdf}
-      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
 See Supporting Information for details.
@@
   The latter represents evolution under multiple selective regimes, and may be more biologically realistic for many datasets. 
-  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{} (Table \ref{table.uses})
-  % add citation
+  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{christin2013anatomical}(Table \ref{table.uses})
   OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supporting Information).  
@@
     We simulated 1000 phylogenies for each combination of tips and models resulting in 56,000 simulated phylogenies in total. 
-    Trees were simulated using the R package TESS \cite{hohna2013fast}. 
+    Trees were simulated using the R package TESS \citep{hohna2013fast}. 
     We then simulated the evolution of a single trait under a Brownian motion model on each phylogeny using the R package MOTMOT \citep{Thomas:2011aa}. 
@@
     For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
+    Indeed, for such trees the confidence limits on parameter estimates are broad (\ref{figure:ouprofiles})
     This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
@@
       \centering
-      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
+      \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
+      \caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ``true'' value of $\alpha$ is 0.
+      }
+      \label{figure:ouprofiles}
+\end{figure}
+% Weird issue with endfloat means these need to have no white space
+
+    \begin{figure}[!htbp]
+      \centering
+      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure3.pdf}
       \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
@@
 
-  % GT:  As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here. 
-  % NC: OK happy to leave it out
-
- % For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
- % Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}.
- % A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.
-
 \section{Recommendations and potential solutions}
@@
       Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for Likelihood ratio tests.
-%We could potentially add a script/workflow for this
       \citet{boettiger2012your} discuss this at length and we recommend reading their paper for a full overview of this approach.
-      % NC: Though their method is hard to apply it does cover all the points we discuss here so we chould make it clearer (as requested by Syst Biol reviewers).
-Also I'd forgotten that the new geiger code means their package is deprecated and can't be used anyway.
 
@@
 Tree type refers to the extinction fraction for the birth-death trees.
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.
-} % NC: might need to modify legend here.
+The rejection rate is the proportion of OU models favoured relative to a Brownian motion model based on Bayes factors \textgreater 2.}
         \bigskip
@@
     Thus this does not reflect any kind of OU process underlying the data. 
-    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in Figure XX. % NC: Not sure what this figure is...
+    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in (\ref{figure:ouhistory}). 
     At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
+
+     \begin{figure}[!htbp]
+      \centering
+      \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure4.pdf}
+      \caption{Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein Uhlenbeck model.
+The covariance between species' traits values is scaled by the intra-specific trait variance (i.e.
+equal to correlation between species' traits).
+This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height: $\frac{t_{ij}}{T}$).
+      }
+      \label{figure:ouhistory}
+\end{figure}
 
@@
 
-% add in ives and garland advice on what is big and what is small here? Or earlier?
-
     \item \textbf{If possible, include ancillary data.}
@@
 \section{Conclusions and outstanding issues}
-% NC: I'm not so sure about this paragraph.
-It reads like a Freckleton paper and I'm not sure how helpful it is.
-I'd be tempted to just cut it and the subheading, and end with the paragraph on how it applies to all models.
-What do you think?
-  A recurring theme from our simulations is that interpretation of $\alpha$ is not straightforward in the presence of measurement error.
+  A recurring theme from our simulations is that interpretation of OU models is not straightforward.
+More focus is needed on the interpretation of $\alpha$ rather than simply model fit.
+Even when the OU model is favoured, $\alpha$ may be so small as to be indistinguishable from Brownian motion in any biological sense.
 It would clearly be very useful to have estimates of measurement error for all species traits, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
-Several approaches for accounting for error have been proposed.
+Several approaches for accounting for error have been proposed and warrant further study for OU models and more widely for models of trait evolution.
 %check references.
-  It would be tempting to create a vector of variances that could be combined with V
-  % NC: this is a leftover from the previous version - is V the variance covariance matrix?
-  to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}.
-However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling.
-  This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton pers. comm.). 
-  An additional challenge common to (especially large) comparative data sets is that measurement error or intraspecific variation is either not available or is inconsistently reported. 
-  One approach if the variances are unknown would be to simultaneously fit Pagel's $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} to estimate the non-Brownian component of the trait variance. 
-  However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.
 
-% NC: Slightly modified i think this is a good end paragraph. Perhaps with a bit more form the original conclusion below.
-  These issues are not limited to OU models. 
-  Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being mislead by measurement error. 
+  Indeed, the problems that we report are not limited to OU models. 
+  Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being misled by measurement error. 
   The fundamental problem is that rejection of the Brownian model in favour of another model does not necessarily say anything about process. 
@@
   We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.
-
-%Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
-
-%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes.
-$\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood.
-It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$.
-Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously.
-\citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989).
-They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes).
-This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.
-
- % The bias % check we mean bias
- % reported above are not unique to the OU model.
-The parameter $\delta$, which tests for accelerating or decelerating rates of evolution \citep{Pagel:1997aa,Pagel:1999aa} also suffers from elevated Type I error and is upwardly biased, favoring models that imply temporally increasing evolutionary rates (Supplemental Fig.
-S8).
-In contrast, although the speciational model ($\kappa$) has slightly elevated Type I errors, the parameter estimates are unbiased for the majority of tree sizes and shapes (Supplemental Fig.
-S6).
-As previously shown (Freckleton et al.
-2002), the $\lambda$ model is conservative with low Type I error rates (Supplemental Fig.
-S7).
-As with the OU model, the $\delta$ models are all strongly favored over Brownian motion when there is error in the species data (Supplemental Figs.
-S9-S11).
-
- %The results we report are not unique to the OU model.
-As noted above, \citet{grafen1989phylogenetic} pointed out that his $\rho$ parameter would likely be biased, and other transformations \citep[e.g., $\lambda$, $\delta$, ACDC;][]{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa} will exhibit similar behavior \citep[e.g., see][]{freckleton2002phylogenetic}.
-\citet{freckleton2002phylogenetic} explored the behavior of Pagel’s $\lambda$ and found that this statistic was biased but in manner that was likely to be conservative in the rejection of the Brownian model (also see Supplemental Fig.
-S7).
-Some transformations may be unbiased: for example, Pagel’s $\kappa$ is a transformation of branch lengths, and is essentially the “standard diagnostic” that is used in testing the assumptions of phylogenetic contrasts, and known to behave appropriately \citep[Supplemental Fig.
-S6;][]{garland1992procedures}.
-
-% Conclusion from original
-%In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species.
-The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes.
-Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process.
-This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}.
-We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.
-The results presented above highlight that in the case of the OU model, this is especially important.
 
@@
   Although free to download, this software is not an Open Source, which has implications for reproducibility.
-  % NC: I'm not sure what else to write here without pissing off Chris etc. Is it possible to put their scripts on GitHub? Can you ask? 
   We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 
@@
 
-  The Ornstein-Uhlenbeck (OU) model describes a mean-reverting process and has the following form, adding an extra term to the Brownian model:
+  The Ornstein Uhlenbeck (OU) model describes a mean-reverting process and has the following form, adding an extra term to the Brownian model:
 
@@
 \noindent
-\textbf{Figure 1}: The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+\textbf{Figure 1}: The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
+See Supporting Information for details.\\
+
+\noindent
+\textbf{Figure 2}: Examples of profile likelihoods for selected simulated datasets. In all cases the ``true'' value of $\alpha$ is 0.
+
+\noindent
+\textbf{Figure 3}: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+Two studies with \textgreater 3000 taxa have been omitted for clarity.
 See Supporting Information for details.
@@
 \noindent
-\textbf{Figure 2}: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
-Two studies with \textgreater 3000 taxa have been omitted for clarity.
-See Supporting Information for details.
+\textbf{Figure 4}: Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein Uhlenbeck model.
+The covariance between species' traits values is scaled by the intra-specific trait variance (i.e.
+equal to correlation between species' traits).
+This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height: $\frac{t_{ij}}{T}$).
 

### 2015-05-11 Manuscript/ohyou.tex
Fixed referencing errors

@@
     For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
-    Indeed, for such trees the confidence limits on parameter estimates are broad (\ref{figure:ouprofiles})
+    Indeed, for such trees the confidence limits on parameter estimates are broad (Figure \ref{figure:ouprofiles})
     This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
@@
     Thus this does not reflect any kind of OU process underlying the data. 
-    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in (\ref{figure:ouhistory}). 
+    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in (Figure \ref{figure:ouhistory}). 
     At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.

### 2015-05-15 Manuscript/ohyou.tex
dealt with  RF's comments

@@
 We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses.
-  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
-  Many of our findings are applicable to other models of evolution, but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.
+  We also show that very small amounts of intraspecific trait variation (including measurement error) can profoundly affect on the performance of models. 
+  These findings will be applicable to other models of evolution, but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
+  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}. 
+  However widespread use of the model is clear evidence that many are unaware of the potential problems.
+We use a simulation approach to summarize the problems and to generate practical recommendations of how to deal with them.
 
@@
   It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilising selection'', this is inaccurate and misleading. 
-As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optimum for that niche.
-  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive} but not as an estimate of stabilising selection in the population genetics sense.
+  As formulated by \citet{hansen1997stabilizing}, a trait has a primary optimum that is the mean of individual species optimum for that trait. 
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive}. 
+  However, this is not an estimate of stabilising selection in the population genetics sense.
 
@@
 
-  The OU model is most commonly used to model the evolution of a continuous character (Table \ref{table.uses}; Supporting Information). 
-  Multiple models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
+  The OU model is most commonly used to model the evolution of a single continuous character (Table \ref{table.uses}; Supporting Information). 
+  Usually several models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
   OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
@@
   The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
-  This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
+  This is rarely interpreted as corresponding to any kind of process,
+  ; instead it improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
+  It is not clear that it was originally intended that the OU model would be used in such a context.
   Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.
@@
   \item \textbf{Likelihood ratio tests are untrustworthy.}
-  We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
-  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered.
+  All likelihood ratio tests depend on asymptotic properties, so there is no guarantee that any parameter such as $\alpha$ will behave appropriately when tested. 
+  In the case of $\alpha$ we would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
+  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as simulation and MCMC, should be considered.
   \item \textbf{Measurement error increases Type I error rates.} 
@@
       Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for Likelihood ratio tests.
-      \citet{boettiger2012your} discuss this at length and we recommend reading their paper for a full overview of this approach.
+      \citet{boettiger2012your} discuss this at length and other papers use similar approaches \citep[e.g.,][]{freckleton2002phylogenetic, martins1991phylogenetic}.
 
@@
       At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation. 
-      We also recommend reading \citet{pennell2015model} as their paper presents similar conclusions with respect to Bayesian methods.
+      See \citet{pennell2015model} for more details on using Bayesian methods in evolutionary models.
     \end{enumerate}

### 2015-05-15 Manuscript/ohyou.tex
dealt with  CV and AM's comments plus some additions

@@
   The OU model is most commonly used to model the evolution of a single continuous character (Table \ref{table.uses}; Supporting Information). 
-  Usually several models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
+  Usually several models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body,cardillo2015geographic,slater2015iterative} are fit to the same continuous character and model selection is then used to determine which model best fits the data.
   OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
@@
     Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
+    Note that $\alpha$ scales with tree height, so it needs to be interpreted relative to this. 
+    Generally the simplest solution is to rescale tree heights to 1 \citep[e.g.,][simulations below]{ives2010phylogenetic}. 
+    \citet{ives2010phylogenetic} suggest interpreting $log( \alpha)$ (after rescaling trees to a height of 1) rather than raw $\alpha$.
+They equate $-log( \alpha) = -4$ as a very low, almost Brownian value, and $-log( \alpha) = 4$ as a very high value.
+    Others \citep[e.g.][]{slater2015iterative,slater2013robust,hansen2012interpreting} prefer to use the ``phylogenetic half-life'': $t_\frac{1}{2}$ (see ``Recommendations for interpreting $\alpha$'' below).
 
@@
       Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). 
-      We treated Bayes factors \textgreater 2 as evidence favouring the OU model. We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
+      We treated Bayes factors \textgreater 2 as evidence favouring the OU model \citep{kass1995bayes}. 
+      We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in.
       Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
@@
       The Bayesian approach results in highly conservative rejection rates regardless for all tree shapes in the absence of measurement error. 
-      While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power. 
+      While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power, although testing would be required to confirm this.
       The Bayesian approach also appears to more readily handle low levels of measurement error (Table \ref{tables:bayesian_error}). 
@@
     Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
-    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
+    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero (e.g., examples in \citealp{Harmon:2010aa}), i.e.
+very close to Brownian.
     In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
@@
     If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
+    $t_\frac{1}{2}$ equal to the height of the phylogeny is a moderate value \citep{hansen2012interpreting}.
     We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
@@
     \item \textbf{If possible, include ancillary data.}
-    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.
+    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}, provided their placement is well known.
     This is implemented in GEIGER \citep{Harmon:2008aa}.
@@
 It would clearly be very useful to have estimates of measurement error for all species traits, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}.
-Several approaches for accounting for error have been proposed and warrant further study for OU models and more widely for models of trait evolution.
-%check references.
+Several approaches for accounting for error have been proposed and warrant wider implementation for OU models and other models of trait evolution \citep[e.g.,][]{lynch1991methods,hansen2012interpreting,martins1997phylogenies,rohlfs2013modeling,ives2007within}, however it is outside our scope to explore these here.
 
@@
 
-\section{Epilogue: (vaguely) reproducible science}
-  At the symposium that generated this special issue, one of us (NC) gave a talk on reproducibility in science. 
-  We have therefore tried to make this paper as reproducible as possible. 
+\section{Epilogue: Open and (vaguely) reproducible science}
+  At the symposium that generated this special issue, one of us (NC) gave a talk on Open Science and reproducibility. 
+  We have therefore tried to make this paper as open and reproducible as possible. 
   All simulated phylogenies, data, and R code are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
@@
   Additionally, we use BayesTraits to run our Bayesian analyses. 
-  Although free to download, this software is not an Open Source, which has implications for reproducibility.
+  Although free to download, this software is not an Open Source, which has implications for long-term development of the methods.
   We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 
-  With a little effort most people should be able to produce something vaguely reproducible, and move us all slightly closer to truly reproducible science.
+  With a little effort most people should be able to produce something vaguely reproducible, and to provide their data and code, moving us all slightly closer to truly reproducible and Open Science.
 
@@
 GHT was supported by a Royal Society University Research Fellowship.
+CV was supported by a Leverhulme Trust Research Project Grant RPG-2013-185.
 AM was supported by BBSRC grant BB/K004344/1 and the computing time was funded by European Research Council Grant no.
-268744, Mother Tongue.
+268744, ``MotherTongue''.
 

### 2015-05-22 Manuscript/ohyou.tex
Added GHT d final edits

@@
   Here we focus on one of these models - the Ornstein Uhlenbeck (OU) model.
-  We show that the OU model is frequently incorrectly favoured over simpler models, and that many studies fitting this model use datasets that are small and prone to Type I error.   
+  We show that the OU model is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, and that many studies fitting this model use datasets that are small and prone to Type I error.
   We also show that very small amounts of error in datasets can have profound effects on the performance of OU models.
@@
   \subsection{Performance of $\alpha$ and Likelihood ratio tests}
-    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}.
+    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT (\citealp{Thomas:2011aa}; \href{https://github.com/ghthomas/motmot}{github.com/ghthomas/motmot}).
     This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model.
@@
 \section{Effects of measurement error on OU model performance}
-  Measurement error can strongly effect the results of comparative analyses \citep{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep[see][]{pennell2015model}.
+  Measurement error can strongly affect the results of comparative analyses \citep{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep[see][]{pennell2015model}.
   Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.
@@
   \item \textbf{Likelihood ratio tests are untrustworthy.}
-  All likelihood ratio tests depend on asymptotic properties, so there is no guarantee that any parameter such as $\alpha$ will behave appropriately when tested. 
+  All Likelihood ratio tests depend on asymptotic properties, so there is no guarantee that any parameter such as $\alpha$ will behave appropriately when tested. 
   In the case of $\alpha$ we would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
@@
       In such cases emphasis should shift to how the $\alpha$ parameter is interpreted. 
-      At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation. 
+      Bayesian analysis is a promising approach for OU model fitting. 
+      However, further testing of the Bayesian approach to simulated data sets under a wide range of values of $\alpha$ is necessary to fully characterize performance. 
       See \citet{pennell2015model} for more details on using Bayesian methods in evolutionary models.
@@
 Tree type refers to the extinction fraction for the birth-death trees.
+The value of $\alpha$ is the median across simulated data sets based on modal estimates from the posterior distribution.
 The rejection rate is the proportion of OU models favoured relative to a Brownian motion model based on Bayes factors \textgreater 2.}
@@
         \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.) using Bayesian methods.
+The value of $\alpha$ is the median across simulated data sets based on modal estimates from the posterior distribution.
 The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.} %NC: Might need to modify legend
@@
     Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
-    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero (e.g., examples in \citealp{Harmon:2010aa}), i.e.
-very close to Brownian.
+    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very small and biologically indistinguishable from Brownian (e.g., examples in \citealp{Harmon:2010aa}).
     In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
@@
   We also do not provide an automated way of reproducing the data collection for our literature review. 
-  Nor do we use tools like Travis CI (travis-ci.org), Docker (www.docker.com), or packrat \citep{Ushey:2015aa} etc. to increase the reproducbility of our analyses.
+  Nor do we use tools like Travis CI (\href{https://travis-ci.org}{travis-ci.org}), Docker (\href{www.docker.com}{www.docker.com}), or packrat \citep{Ushey:2015aa} etc.
+to increase the reproducibility of our analyses.
   Additionally, we use BayesTraits to run our Bayesian analyses. 
-  Although free to download, this software is not an Open Source, which has implications for long-term development of the methods.
-  We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 
+  BayesTraits is free to download as a binary executable for various platforms but it is not Open Source. 
+  This means that while BayesTraits analyses are reproducible the software has limitations with respect to long-term development of the code. 
+  We included this epilogue to highlight that fully reproducible and Open Science is really hard, but we are trying to improve! 
   With a little effort most people should be able to produce something vaguely reproducible, and to provide their data and code, moving us all slightly closer to truly reproducible and Open Science.
@@
 \section{Acknowledgments}
-  Thanks to Mark Pagel, Matt Pennell, Graham Slater and Rich FitzJohn for fruitful discussions about OU models, and three anonymous reviewers for helpful comments on a previous version of this paper.
+  Thanks to Rich FitzJohn, Mark Pagel, Matt Pennell and Graham Slater for fruitful discussions about OU models, and three anonymous reviewers for helpful comments on a previous version of this paper. 
   NC was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696. 
@@
   CV was supported by a Leverhulme Trust Research Project Grant RPG-2013-185. 
-AM was supported by BBSRC grant BB/K004344/1 and the computing time was funded by European Research Council Grant no.
-268744, ``MotherTongue''.
+  AM was supported by BBSRC grant BB/K004344/1 and the computing time was funded by European Research Council Grant no. 268744, ``MotherTongue''.
 

### 2015-06-23 Manuscript/Revision/response.tex
Initial commit of response letter

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[colorlinks = true, urlcolor = blue]{hyperref} %hyperlinks
\usepackage[osf]{mathpazo}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\signature{Natalie Cooper \\ Gavin Thomas}
\address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ ncooper@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}

Referees' Comments to Authors:
Referee: 1

Comments to the Author
The authors discuss some potential pitfalls in model choice and subsequent interpretation regarding the widely used OU model for analysing a single continuous trait on a phylogenetic tree. Due to the heavy usage of this model (as evidenced by the authors own literature survey) this is clearly an important paper, and the authors provide a broad array of simulations, which span a sufficient number of scenarios to characterise some major biases of the typical likelihood based model-fitting approach. The authors' interpretation of the results is measured and sensible (neither overly doom laden nor free of potential solutions). They have also strived to make their study repeatable whilst acknowledging the challenges in doing so and the limitations of their chosen approach. I have no problem supporting the publication of this paper, although I have a few notes (below) on potential improvements. These overwhelmingly ask for additional clarity on specific points and so I see them as minor.

L22 - Except these are exclusively for continuous data. Maybe use "core" instead of foundation, or a caveat like "at least for continuous data".
L28 (and L199) - I'm not sure if Type 1 error is the correct term to use as this is (to my mind) a frequentists term. I realise that we are (in a sense) rejecting the null of Brownian motion (albeit falsely) and that the LRT gives us a p-value to do so. However, the notion isn't necessarily generalisable to, e.g., three or more models. Perhaps the authors could just state explicitly what they mean by Type I errors to avoid confusion.
L89 - The year range given here and that in the figure/figure caption are different.
L102 - This explanation seems a little incomplete. Why isn't it?
L124 - This also seems incomplete. Maybe insert "merely"?
L156 - I think it would be helpful to discuss mu a bit more. Questions I would like to see answered include: Presumably this is the trait "optimum"? What does it mean to have an optimum outside of the range of the tip values? What if mu is unrealistic, e.g., continuous measure is wingspan and mu is 1000km? If mu is large/far from the tip mean does that mean alpha can be small and the model still be a better fit? etc.
L167 - I'm not totally clear what is meant by tree height (I think I know, but would prefer it to be defined explicitly), nor how or why it scales with alpha.
L188 - Lol.
L219 - Why did the authors chose a Yule model here and a birth-death earlier? Does it mean one is ultrametric and the other not? (Maybe this relates to the fossil point below.)
L263 - This paragraph could do with some unpacking. What asymptotic properties? How is alpha bounded? etc.
L267 - Does this extend, then, to AIC(c)s? (As many people use AIC instead of the LRT and fit more than two models this could be important.)
L283 - "in fitting model" missing "the"?.
L299 - Citations should be in date order.
L325 - Do you mean "regardless OF"?
L330 - Potential pregnancy (missing period).
L342 - "processES"?
L367 - Although in practice are there not ways to set limits on the min-max value of alpha?
L376 - Maybe clarify exactly why (I think I have an idea!).
L382 - Could or should? (Conflict of interest: I am a palaeobiologist!) It seems like fossils could really help as by definition they have smaller values of T, and hence are closer to the root, and hence are better at constraining the root, and hence (potentially) help with model choice in general. (E.g., AFAIK you can only really favour a trend model if you have fossils in the tree.) In addition, when the authors simulate the birth-death models are they using non-ultrametric trees? Or just pruning extinct tips? If the latter then they aren't really testing whether fossils can help, but maybe this could be mentioned as a possible future simulation study.
L385 - Maybe I'm missing something here, but to my mind a non-ultrametric tree is not special in terms of co-variance. Rather it is the variances that are unusual. I.e., the diagonal of the VCV will not just be the same number (same T) for all species. In other words if you modified a non-ultrametric tree to make it ultrametric by making all tips range through to the "present" you would only be modifying the diagonal.
L414 - Epilogue is fine in terms of content, but language is very informal ("vaguely", "messy", "!"). Maybe this isn't an issue, but does read to me as a definite shift in tone from the rest of the paper.
Figure 2 caption - I think this caption could do more to explain the plot. What is meant by "rooty"? What is meant by "tippy"? What does each line represent? What are the dashed lines? etc.
Figure 3 - Would be nice to have more "breaks" in the histogram to help visualise how the data is really distributed. The binning seems too coarse at the moment.

Referee: 2

Comments to the Author
In this article, Cooper and colleagues explore through simulations one of phylogenetic comparative biologists worst kept but most ignored secrets — that the OU model (or, more specifically the single stationary peak version of the OU model) has poor statistical performance. Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU α - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU. The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.

Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs. BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I’m sure this one will be much read and cited.

However, while I understand the authors’ frustrations with the fitting and interpretations of single peak OU processes that are common in the literature, I think they miss an opportunity to really emphasize their concerns here and demonstrate to readers that it’s easy to correct oneself when support for OU comes back in comparative analyses. Though the authors do note that a best fitting OU model could be nothing more than a noisy BM processes based on parameter estimates, a closer look at their simulation results reveals that this is also the case here. For example, on their simulated trees rescaled to unit height, the highest median alpha estimated for the BM datasets is 0.165 for the 25 tip Yule tree. This translates into a phylogenetic half life of 4.2 time units however, 4 times the age of the simulated clade, which no sane macroevolutionary biologist would interpret as being too different from Brownian motion at this time scale! Furthermore, the upper 95% HPDs (Quantiles surely?) are, in many cases low enough to result in questionable acceptance of OU over BM for any of the simulations.

I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives. If the aim here is to caution readers that a best-fitting OU process does not always translate to evolution about a  constrained peak, then showing this with parameter estimates from the simulations would be a useful contribution. This kind of constructive comment would also help those readers who would view this only as a critique of OU models.

As i thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works. I suspect (from my own dealings with authors during reviews) that many do not have an intuitive feel for how the model works still. Based on this, I’d suggest that the authors move their appendix to the main ms and integrate it with the OU model outline on lines 139-174. It may additionally be useful to illustrate the process, both as a diffusion process and as its effects on the phylogenetic variance covariance structure. These are suggestions only, but I suspect that providing a more complete background to the models would further aid naive readers / students who are to be the main beneficiaries of this paper.

Minor comments

SSP vs single peak OU.

I think it would be useful to clarify early on in the ms that the scenario being explored is specifically a single stationary peak model (a la Harmon et al. 2010), rather than a single optimum model. Although I would agree that most users of these methods are more familiar with applying the latter term, there is a small distinction. A single peak OU process may have the root state the same as or different from the long term optimum, whereas the SSP model has both the same. Furthermore, interpretation of these two models with α ≅ 0 is slightly different. For SSP, α ≅ 0 approximates BM, but for a single peak model where the peak differs from the root state, α ≅ 0 collapses to BM with a trend (e.g., see Hansen 1996, Benson et al. 2014). Because the long term mean is introduced as a parameter of the OU process in eq 2 (line 156), it is perhaps worth pointing out this distinction at some point thereafter. Of course, the case where the root state and long term mean are not equal cannot be fitted when fossil data / strong root or internal node priors are lacking and so is unavailable to most PCM users anyway. But still, for completeness I think this is worth pointing out.

line 114. Perhaps my own bugaboo, but I hate the term “to control for phylogeny”. Perhaps replace this with “to model phylogenetically structured residual error in evolutionary correlations”. Revell (2010) might also make a nice citation here.

Line 172: at the risk of losing a citation (although i think you’ll retain me elsewhere) Slater and Pennell used a rate half-life in the context of the exponential decline in rate under the ACDC model, not a phylogenetic half-life for the exponential approach of a trait to its long term mean under the OU model.

The stepping stone analyses and results are very interesting. In Baele et al. (2012), which introduced path sampling for Bayesian demographic model selection using the BEAST software, the authors showed via simulation that Path Sampling and Stepping Stones performed much better than harmonic mean estimators or other model choice metrics at selecting a time homogeneous model over exponential growth (and also at choosing weak exponential growth over constant population size models). There seems a nice parallel here and this suggests that these kinds of approaches should probably be better integrated into comparative methods,.

Lines 263-269: it may be more useful to be explicit here that LRTs assume that the LR statistic is asymptotically chi-square distributed and that this is unlikely the case here (perhaps you could show this using the simulations?). Link this to lines 288-299 to show how to obtain a simulated distribution for the LRT.

lines 404-407: This works the other way too; not incorporating measurement error could lead one to reject a “high phylogenetic signal” model like EB in favor of BM when EB is intact the true model.

Figure 2 legend what are the black and red dashed lines?

Signed: Graham Slater

References

Baele, G., Lemey, P., Bedford, T., Rambaut, A., Suchard, M. A., & Alekseyenko, A. V. (2012). Improving the accuracy of demographic and molecular clock model comparison while accommodating phylogenetic uncertainty. Molecular biology and evolution, 29(9), 2157-2167.

Benson, R. B., Frigot, R. A., Goswami, A., Andres, B., & Butler, R. J. (2014). Competition and constraint drove Cope's rule in the evolution of giant flying reptiles. Nature communications, 5.

Hansen, T. F. (1997). Stabilizing selection and the comparative analysis of adaptation. Evolution, 1341-1351.

Revell, L. J. (2010). Phylogenetic signal and linear regression on species data. Methods in Ecology and Evolution, 1(4), 319-329.

\end{letter}
\end{document}

### 2015-06-23 Manuscript/Revision/response.tex
Fixed minor comments by reviewers

@@
-\documentclass[11pt]{letter}
-\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
+\documentclass[12pt]{letter}
+\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
 \usepackage{color}
@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper \\ Gavin Thomas}
+\signature{Natalie Cooper \\ Gavin Thomas \\ and colleagues}
 \address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ ncooper@tcd.ie}
@@
 \begin{letter}{}
+\opening{Dear Dr Goswami}
 
-Referees' Comments to Authors:
-Referee: 1
+Please find enclosed our resubmission of our manuscript ``A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies".
+We include below details of how we have dealt with the reviewers' comments.
+The reviewers' comments are in blue and our response is in black.
 
-Comments to the Author
+\textcolor{blue}{\textbf{Referee: 1}\\
 The authors discuss some potential pitfalls in model choice and subsequent interpretation regarding the widely used OU model for analysing a single continuous trait on a phylogenetic tree.
@@
 I have no problem supporting the publication of this paper, although I have a few notes (below) on potential improvements.
-These overwhelmingly ask for additional clarity on specific points and so I see them as minor.
+These overwhelmingly ask for additional clarity on specific points and so I see them as minor.}
 
-L22 - Except these are exclusively for continuous data. Maybe use "core" instead of foundation, or a caveat like "at least for continuous data".
-L28 (and L199) - I'm not sure if Type 1 error is the correct term to use as this is (to my mind) a frequentists term.
+We thank the reviewer for their positive comments.
+
+\textcolor{blue}{L22 - Except these are exclusively for continuous data. Maybe use "core" instead of foundation, or a caveat like "at least for continuous data".}
+
+Changed to core.
+
+\textcolor{blue}{L28 (and L199) - I'm not sure if Type 1 error is the correct term to use as this is (to my mind) a frequentists term.
 I realise that we are (in a sense) rejecting the null of Brownian motion (albeit falsely) and that the LRT gives us a p-value to do so.
 However, the notion isn't necessarily generalisable to, e.g., three or more models.
-Perhaps the authors could just state explicitly what they mean by Type I errors to avoid confusion.
-L89 - The year range given here and that in the figure/figure caption are different.
-L102 - This explanation seems a little incomplete. Why isn't it?
-L124 - This also seems incomplete. Maybe insert "merely"?
-L156 - I think it would be helpful to discuss mu a bit more.
+Perhaps the authors could just state explicitly what they mean by Type I errors to avoid confusion.}
+
+Fair point. We mentioned what we mean by Type I error slightly above this point, so have replaced it with "and thus more prone to this problem" rather than repeating it in the abstract.
+
+Later we have clarified what we mean by Type I error.
+
+\textcolor{blue}{L89 - The year range given here and that in the figure/figure caption are different.}
+
+Yes this is correct.
+We meant to show that there have been lots of papers just within the last few years, and the figure was to show how these have increased over a slightly longer timeframe.
+We have rephrased to make this distinction clearer.
+
+\textcolor{blue}{L102 - This explanation seems a little incomplete. Why isn't it?}
+
+******
+
+\textcolor{blue}{L124 - This also seems incomplete. Maybe insert "merely"?}
+
+Added merely.
+
+\textcolor{blue}{L156 - I think it would be helpful to discuss mu a bit more.
 Questions I would like to see answered include: Presumably this is the trait "optimum"?
@@
 If mu is large/far from the tip mean does that mean alpha can be small and the model still be a better fit?
-etc.
-L167 - I'm not totally clear what is meant by tree height (I think I know, but would prefer it to be defined explicitly), nor how or why it scales with alpha.
-L188 - Lol.
-L219 - Why did the authors chose a Yule model here and a birth-death earlier? Does it mean one is ultrametric and the other not? (Maybe this relates to the fossil point below.)
-L263 - This paragraph could do with some unpacking. What asymptotic properties? How is alpha bounded? etc.
-L267 - Does this extend, then, to AIC(c)s? (As many people use AIC instead of the LRT and fit more than two models this could be important.)
-L283 - "in fitting model" missing "the"?.
-L299 - Citations should be in date order.
-L325 - Do you mean "regardless OF"?
-L330 - Potential pregnancy (missing period).
-L342 - "processES"?
-L367 - Although in practice are there not ways to set limits on the min-max value of alpha?
-L376 - Maybe clarify exactly why (I think I have an idea!).
-L382 - Could or should?
+etc.}
+
+****
+
+\textcolor{blue}{L167 - I'm not totally clear what is meant by tree height (I think I know, but would prefer it to be defined explicitly), nor how or why it scales with alpha.}
+
+****
+\textcolor{blue}{L188 - Lol.}
+
+:)
+
+\textcolor{blue}{L219 - Why did the authors chose a Yule model here and a birth-death earlier? Does it mean one is ultrametric and the other not? (Maybe this relates to the fossil point below.)}
+
+****
+
+\textcolor{blue}{L263 - This paragraph could do with some unpacking. What asymptotic properties? How is alpha bounded? etc.}
+
+***
+\textcolor{blue}{L267 - Does this extend, then, to AIC(c)s? (As many people use AIC instead of the LRT and fit more than two models this could be important.)}
+
+***
+
+\textcolor{blue}{L283 - "in fitting model" missing "the"?.}
+
+Added "the".
+
+\textcolor{blue}{L299 - Citations should be in date order.}
+
+Corrected.
+
+\textcolor{blue}{L325 - Do you mean "regardless OF"?}
+
+Corrected.
+
+\textcolor{blue}{L330 - Potential pregnancy (missing period).}
+
+Haha corrected.
+
+\textcolor{blue}{L342 - "processES"?}
+
+Corrected.
+
+\textcolor{blue}{L367 - Although in practice are there not ways to set limits on the min-max value of alpha?}
+
+****
+
+\textcolor{blue}{L376 - Maybe clarify exactly why (I think I have an idea!).}
+
+***
+
+\textcolor{blue}{L382 - Could or should?
 (Conflict of interest: I am a palaeobiologist!) It seems like fossils could really help as by definition they have smaller values of T, and hence are closer to the root, and hence are better at constraining the root, and hence (potentially) help with model choice in general.
@@
 Or just pruning extinct tips?
-If the latter then they aren't really testing whether fossils can help, but maybe this could be mentioned as a possible future simulation study.
-L385 - Maybe I'm missing something here, but to my mind a non-ultrametric tree is not special in terms of co-variance.
+If the latter then they aren't really testing whether fossils can help, but maybe this could be mentioned as a possible future simulation study.}
+
+***
+
+\textcolor{blue}{L385 - Maybe I'm missing something here, but to my mind a non-ultrametric tree is not special in terms of co-variance.
 Rather it is the variances that are unusual.
 I.e., the diagonal of the VCV will not just be the same number (same T) for all species.
-In other words if you modified a non-ultrametric tree to make it ultrametric by making all tips range through to the "present" you would only be modifying the diagonal.
-L414 - Epilogue is fine in terms of content, but language is very informal ("vaguely", "messy", "!").
-Maybe this isn't an issue, but does read to me as a definite shift in tone from the rest of the paper.
-Figure 2 caption - I think this caption could do more to explain the plot. What is meant by "rooty"? What is meant by "tippy"? What does each line represent? What are the dashed lines? etc.
-Figure 3 - Would be nice to have more "breaks" in the histogram to help visualise how the data is really distributed. The binning seems too coarse at the moment.
+In other words if you modified a non-ultrametric tree to make it ultrametric by making all tips range through to the "present" you would only be modifying the diagonal.}
 
-Referee: 2
+***
 
-Comments to the Author
-In this article, Cooper and colleagues explore through simulations one of phylogenetic comparative biologists worst kept but most ignored secrets — that the OU model (or, more specifically the single stationary peak version of the OU model) has poor statistical performance.
-Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU α - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU.
-The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.
+\textcolor{blue}{L414 - Epilogue is fine in terms of content, but language is very informal ("vaguely", "messy", "!").
+Maybe this isn't an issue, but does read to me as a definite shift in tone from the rest of the paper.}
+
+We have made this a little less informal in tone.
+
+\textcolor{blue}{Figure 2 caption - I think this caption could do more to explain the plot.
+What is meant by "rooty"?
+What is meant by "tippy"?
+What does each line represent?
+What are the dashed lines?
+etc.}
+
+***
+
+\textcolor{blue}{Figure 3 - Would be nice to have more "breaks" in the histogram to help visualise how the data is really distributed. The binning seems too coarse at the moment.}
+
+We have added more breaks to this figure.
+
+\textcolor{blue}{
+\textbf{Referee: 2 (Graham Slater)}\\
+In this article, Cooper and colleagues explore through simulations one of phylogenetic comparative biologists worst kept but most ignored secrets - that the OU model (or, more specifically the single stationary peak version of the OU model) has poor statistical performance.
+Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU $\alpha$ - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU.
+The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.}
 
 Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs.
-BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I’m sure this one will be much read and cited.
+BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I'm sure this one will be much read and cited.
 
-However, while I understand the authors’ frustrations with the fitting and interpretations of single peak OU processes that are common in the literature, I think they miss an opportunity to really emphasize their concerns here and demonstrate to readers that it’s easy to correct oneself when support for OU comes back in comparative analyses.
+We thank the reviewer for their positive comments.
+
+\textcolor{blue}{However, while I understand the authors’ frustrations with the fitting and interpretations of single peak OU processes that are common in the literature, I think they miss an opportunity to really emphasize their concerns here and demonstrate to readers that it's easy to correct oneself when support for OU comes back in comparative analyses.
 Though the authors do note that a best fitting OU model could be nothing more than a noisy BM processes based on parameter estimates, a closer look at their simulation results reveals that this is also the case here.
@@
 This translates into a phylogenetic half life of 4.2 time units however, 4 times the age of the simulated clade, which no sane macroevolutionary biologist would interpret as being too different from Brownian motion at this time scale!
-Furthermore, the upper 95% HPDs (Quantiles surely?) are, in many cases low enough to result in questionable acceptance of OU over BM for any of the simulations.
+Furthermore, the upper 95\% HPDs (Quantiles surely?) are, in many cases low enough to result in questionable acceptance of OU over BM for any of the simulations.}
 
-I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives.
+****
+
+\textcolor{blue}{I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives.
 If the aim here is to caution readers that a best-fitting OU process does not always translate to evolution about a  constrained peak, then showing this with parameter estimates from the simulations would be a useful contribution.
-This kind of constructive comment would also help those readers who would view this only as a critique of OU models.
+This kind of constructive comment would also help those readers who would view this only as a critique of OU models.}
 
-As i thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works.
+****
+
+\textcolor{blue}{As i thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works.
 I suspect (from my own dealings with authors during reviews) that many do not have an intuitive feel for how the model works still.
@@
 It may additionally be useful to illustrate the process, both as a diffusion process and as its effects on the phylogenetic variance covariance structure.
-These are suggestions only, but I suspect that providing a more complete background to the models would further aid naive readers / students who are to be the main beneficiaries of this paper.
+These are suggestions only, but I suspect that providing a more complete background to the models would further aid naive readers / students who are to be the main beneficiaries of this paper.}
 
-Minor comments
+***
 
-SSP vs single peak OU.
-
+\textcolor{blue}{\textbf{Minor comments}\\
+\textbf{SSP vs single peak OU.}\\
 I think it would be useful to clarify early on in the ms that the scenario being explored is specifically a single stationary peak model (a la Harmon et al.
@@
 Of course, the case where the root state and long term mean are not equal cannot be fitted when fossil data / strong root or internal node priors are lacking and so is unavailable to most PCM users anyway.
-But still, for completeness I think this is worth pointing out.
+But still, for completeness I think this is worth pointing out.}
 
-line 114.
+****
+
+\textcolor{blue}{line 114.
 Perhaps my own bugaboo, but I hate the term “to control for phylogeny”.
 Perhaps replace this with “to model phylogenetically structured residual error in evolutionary correlations”.
-Revell (2010) might also make a nice citation here.
+Revell (2010) might also make a nice citation here.}
 
-Line 172: at the risk of losing a citation (although i think you’ll retain me elsewhere) Slater and Pennell used a rate half-life in the context of the exponential decline in rate under the ACDC model, not a phylogenetic half-life for the exponential approach of a trait to its long term mean under the OU model.
+***
 
-The stepping stone analyses and results are very interesting.
+\textcolor{blue}{Line 172: at the risk of losing a citation (although i think you’ll retain me elsewhere) Slater and Pennell used a rate half-life in the context of the exponential decline in rate under the ACDC model, not a phylogenetic half-life for the exponential approach of a trait to its long term mean under the OU model.}
+
+Sorry for the error, we have removed the citation (but it's in there later anyway!)
+
+\textcolor{blue}{The stepping stone analyses and results are very interesting.
 In Baele et al.
 (2012), which introduced path sampling for Bayesian demographic model selection using the BEAST software, the authors showed via simulation that Path Sampling and Stepping Stones performed much better than harmonic mean estimators or other model choice metrics at selecting a time homogeneous model over exponential growth (and also at choosing weak exponential growth over constant population size models).
-There seems a nice parallel here and this suggests that these kinds of approaches should probably be better integrated into comparative methods,.
+There seems a nice parallel here and this suggests that these kinds of approaches should probably be better integrated into comparative methods.}
 
-Lines 263-269: it may be more useful to be explicit here that LRTs assume that the LR statistic is asymptotically chi-square distributed and that this is unlikely the case here (perhaps you could show this using the simulations?).
-Link this to lines 288-299 to show how to obtain a simulated distribution for the LRT.
+***
 
-lines 404-407: This works the other way too; not incorporating measurement error could lead one to reject a “high phylogenetic signal” model like EB in favor of BM when EB is intact the true model.
+\textcolor{blue}{Lines 263-269: it may be more useful to be explicit here that LRTs assume that the LR statistic is asymptotically chi-square distributed and that this is unlikely the case here (perhaps you could show this using the simulations?).
+Link this to lines 288-299 to show how to obtain a simulated distribution for the LRT.}
 
-Figure 2 legend what are the black and red dashed lines?
+***
 
-Signed: Graham Slater
+\textcolor{blue}{lines 404-407: This works the other way too; not incorporating measurement error could lead one to reject a “high phylogenetic signal” model like EB in favor of BM when EB is intact the true model.}
 
-References
+***
+
+\textcolor{blue}{Figure 2 legend what are the black and red dashed lines?}
+
+***
 
@@
 
+\closing{Yours sincerely,}
+
 \end{letter}

### 2015-06-23 Analyses/OhYouFigures.R
Added extra breaks to Fig 3 in response to reviewer comment

@@ -34,8 +34,9 @@ ds2$General.use <- ds2$General.use[, drop = TRUE]
 summary.table <- tapply(ds2$Paper, list(ds2$General.use, ds2$Noptima), length)
 
 # Figure 3
-pdf("Manuscript/Figures/OhYou_Figure2.pdf")
+pdf("Manuscript/Figures/OhYou_Figure3.pdf")
   par(bty = "l")
   hist(ds1$Ntips[which(ds1$Ntips <= 1000)], las = 1, cex.lab = 1.2, cex.axis = 1.2,
-       ylab = "Number of papers", xlab = "Number of taxa", ylim = c(0, 80), main = "")
+       ylab = "Number of papers", xlab = "Number of taxa", ylim = c(0, 30), main = "",
+       breaks = seq(from = 0, to = 800, by = 25))
 dev.off()

### 2015-06-23 Manuscript/Revision/response.tex
Few more edits added

@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper \\ Gavin Thomas \\ and colleagues}
+\signature{Natalie Cooper and colleagues}
 \address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ ncooper@tcd.ie}
@@
 
-Please find enclosed our resubmission of our manuscript ``A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies".
+Please find enclosed our resubmission of our manuscript ``A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies" (BJLS-4095).
 We include below details of how we have dealt with the reviewers' comments.
@@
 
-******
+In population genetics stabilising selection describes selection within a population of a species towards a fitness optimum.
+This is quite a different process biologically to the branching evolution of trait along a phylogeny and among different species towards an optimum trait value.
+We have added some explanation here to hopefully make this clearer.
 
@@
 
-****
+Because of the way that mu is estimated it will never fall outside the tip values or be wildly unrealistic, as it is essentially just a mean of the tip values weighted by branch length.
+We have added clarification to the text.
 
@@
 
-***
+We have added details to this Figure caption.
 

### 2015-06-26 Manuscript/Revision/response.tex
Few more reviewers comments dealt with

@@
 Yes this is correct.
-We meant to show that there have been lots of papers just within the last few years, and the figure was to show how these have increased over a slightly longer timeframe.
+We meant to show that there have been lots of papers just within the last few years, and the figure was to show how these have increased over a slightly longer time-frame.
 We have rephrased to make this distinction clearer.
@@
 
-In population genetics stabilising selection describes selection within a population of a species towards a fitness optimum.
+In population genetics stabilising selection describes selection within a species towards a fitness optimum.
 This is quite a different process biologically to the branching evolution of trait along a phylogeny and among different species towards an optimum trait value.
 We have added some explanation here to hopefully make this clearer.
+The population vs.
+clade level difference is key.
 
@@
 
-Because of the way that mu is estimated it will never fall outside the tip values or be wildly unrealistic, as it is essentially just a mean of the tip values weighted by branch length.
-We have added clarification to the text.
+Yes mu is the optimal value. We have clarified this in the text.
+
+mu can sometimes be wildly inaccurate, though this is generally in implementations where the root state is estimated too leading to strange values appearing optimal on the likelihood surface.
+This is generally fixed by not estimating the root state, and instead using mu as the root state or a root state that is distributed according to the stationary distribution of the OU process.
+My feeling is that this happens rarely (most implementations don't estimate the root state) and hopefully people would notice this and take any results with a large pinch of salt.
+We have added a brief statement to the discussion to alert people to this fact.
+We did not go into mu too much here as our focus was on alpha since this is the parameter people tend to interpret evolutionarily.
 
@@
 
-****
+Tree height is, probably as you suspected, the maximum height from the root to the tips of the tree.
+We have added this to the text.
+Alpha will scale negatively with tree height because the longer/taller the tree is, the more time there is for a trait to evolve back towards the optimum thus alpha - the strength of the force pulling the trait back towards the optimum, can be smaller.
+We have also added this to the text.
+
 \textcolor{blue}{L188 - Lol.}

### 2015-06-29 Manuscript/Revision/response.tex
Few more reviewer comments dealt with

@@
 
-Fair point. We mentioned what we mean by Type I error slightly above this point, so have replaced it with "and thus more prone to this problem" rather than repeating it in the abstract.
-
-Later we have clarified what we mean by Type I error.
+Fair point.
+We mentioned what we mean by Type I error slightly above this point, so have replaced it with "and thus more prone to this problem" rather than repeating it in the abstract.
+Later in the manuscript we have clarified what we mean by Type I error when we first mention it.
 
@@
 We have added this to the text.
-Alpha will scale negatively with tree height because the longer/taller the tree is, the more time there is for a trait to evolve back towards the optimum thus alpha - the strength of the force pulling the trait back towards the optimum, can be smaller.
+Alpha will scale negatively with tree height because the longer/taller the tree is, the more time there is for a trait to evolve back towards the optimum thus alpha - the strength of the force pulling the trait back towards the optimum - can be smaller.
 We have also added this to the text.
@@
 
+We have clarified what we mean by bounded (alpha must be 0 or greater) and reworded this section.
 ***
+
 \textcolor{blue}{L267 - Does this extend, then, to AIC(c)s? (As many people use AIC instead of the LRT and fit more than two models this could be important.)}
@@
 Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU $\alpha$ - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU.
-The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.}
+The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.
 
 Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs.
-BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I'm sure this one will be much read and cited.
+BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I'm sure this one will be much read and cited.}
 

### 2015-07-06 Manuscript/Revision/response.tex
Finished my reviewer editsas best I can

@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper and colleagues}
-\address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ ncooper@tcd.ie}
+\signature{Natalie Cooper, Gavin Thomas and colleagues}
+\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ nhcooper12@gmail.com}
 \longindentation=0pt
@@
 My feeling is that this happens rarely (most implementations don't estimate the root state) and hopefully people would notice this and take any results with a large pinch of salt.
-We have added a brief statement to the discussion to alert people to this fact.
-We did not go into mu too much here as our focus was on alpha since this is the parameter people tend to interpret evolutionarily.
+We have added a brief statement to alert people to this fact.
+We have also added more details on how mu is related to the state at the root as requested by the second reviewer.
+We did not go into mu too much here as our focus was on alpha since this is the parameter people tend to interpret.
 
@@
 
-****
+\textcolor{red}{Gavin?}
 
@@
 
-We have clarified what we mean by bounded (alpha must be 0 or greater) and reworded this section.
-***
+% We have clarified what we mean by bounded (alpha must be 0 or greater) and reworded this section to make it clearer and also in response to reviewer 2 below.
+\textcolor{red}{I think Rob wrote this originally - can you check I've reworded it correctly?}
 
@@
 
-***
+\textcolor{red}{Again Rob wrote this originally. Any ideas?}
 
@@
 
-****
+Yes that's correct, for example in GEIGER the default max is alpha = 150. We have added this information to the text.
 
@@
 
-***
+If t1/2 is very high then it suggests it will take twice that long to reach the optimal trait value for the species.
+Thus it suggests that selection for that optimum is extremely weak at best.
+For example if t1/2 is 100 million years, then it suggests that a species will take 200 million years to reach the trait optimum.
+This is hardly suggestive of a strong selective pressure towards that optimum.
+We have added this example to help clarify this point in the text.
 
@@
 
-***
+\textcolor{red}{Gavin?}
 
@@
 
-We have made this a little less informal in tone.
+We have made this less informal in tone.
 
@@
 
-We have added details to this Figure caption.
+% We have added details to this Figure caption.
+\textcolor{red}{Gavin?}
 
@@
 
-\textcolor{blue}{
-\textbf{Referee: 2 (Graham Slater)}\\
+\textcolor{blue}{\textbf{Referee: 2 (Graham Slater)}\\
 In this article, Cooper and colleagues explore through simulations one of phylogenetic comparative biologists worst kept but most ignored secrets - that the OU model (or, more specifically the single stationary peak version of the OU model) has poor statistical performance.
 Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU $\alpha$ - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU.
-The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.
+The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.}
 
-Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs.
+\textcolor{blue}{Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs.
 BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I'm sure this one will be much read and cited.}
@@
 
-****
-
 \textcolor{blue}{I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives.
@@
 
-****
+\textcolor{red}{Gavin? We may need to discuss this as I'm not entirely sure what Graham means?}
 
-\textcolor{blue}{As i thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works.
+\textcolor{blue}{As I thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works.
 I suspect (from my own dealings with authors during reviews) that many do not have an intuitive feel for how the model works still.
-Based on this, I’d suggest that the authors move their appendix to the main ms and integrate it with the OU model outline on lines 139-174.
+Based on this, I'd suggest that the authors move their appendix to the main ms and integrate it with the OU model outline on lines 139-174.
 It may additionally be useful to illustrate the process, both as a diffusion process and as its effects on the phylogenetic variance covariance structure.
@@
 
-***
+We moved the details of the model to the appendix for the same reasons that the reviewer mentions - that it was aimed at naive readers/students.
+There is evidence to suggest that rather than helping naive readers, too many equations actually impedes understanding and communication of concepts (see Fawcett and Higginson, 2012 PNAS; http://www.pnas.org/content/109/29/11735.short), thus we prefer to keep this as it is so as not to discourage people from reading further.
+\textcolor{red}{Not sure if you agree with me here, but I didn't really want to move the appendix back into the paper as I think it's confusing.}
+
+Illustrating the process however is a nice idea so we have added a figure to the paper that does this. 
+\textcolor{red}{Do you have code handy to make a figure for this? I'd have to write some from scratch and I'm meant to be on holiday...}
 
@@
 
-****
+This is a good point and we have clarified this in the text under where we first introduce mu.
 
@@
 
-***
+We have changed this and added the citation, though in brackets we have left the term "controlling for phylogeny"
+ as while it's not technically correct, it's how many people refer to this procedure.
 
@@
 
-Sorry for the error, we have removed the citation (but it's in there later anyway!)
+Sorry for the error, we have removed the citation (but yes it's in there later anyway!)
 
@@
 
-***
+\textcolor{blue}{Baele, G., Lemey, P., Bedford, T., Rambaut, A., Suchard, M.
+A., \& Alekseyenko, A.
+V.
+(2012).
+Improving the accuracy of demographic and molecular clock model comparison while accommodating phylogenetic uncertainty.
+Molecular biology and evolution, 29(9), 2157-2167.}
+
+\textcolor{red}{Gavin?}
 
@@
 
-***
+We have altered this paragraph to make it clearer, also in response to comments from reviewer 1.
+We have also now signposted that potential solutions are described below in the "Recommendations" section.
 
@@
 
-***
+We have added a comment in the text to say that measurement error can also favour BM over the true model.
 
@@
 
-***
+% We have added details to this Figure caption.
+\textcolor{red}{Gavin?}
 
-Baele, G., Lemey, P., Bedford, T., Rambaut, A., Suchard, M.
-A., & Alekseyenko, A.
-V.
-(2012).
-Improving the accuracy of demographic and molecular clock model comparison while accommodating phylogenetic uncertainty.
-Molecular biology and evolution, 29(9), 2157-2167.
-
-Benson, R. B., Frigot, R. A., Goswami, A., Andres, B., & Butler, R. J. (2014). Competition and constraint drove Cope's rule in the evolution of giant flying reptiles. Nature communications, 5.
-
-Hansen, T. F. (1997). Stabilizing selection and the comparative analysis of adaptation. Evolution, 1341-1351.
-
-Revell, L. J. (2010). Phylogenetic signal and linear regression on species data. Methods in Ecology and Evolution, 1(4), 319-329.
+Let us know if you require any further information,
 

### 2015-07-21 Manuscript/Revision/response.tex
GHT's edits

@@
 
-\textcolor{red}{Gavin?}
+\textcolor{red}{As noted above, all of the simulated trees were ultrametric. }
 
@@
 
+\textcolor{red}{Not sure this is quite right - isn't the half-life the time it takes to get halfway to the optimum?
+I don't think it is a linear process so the halfway point in trait value is not also the halfway point in time it will take to get the optimum}
+
 \textcolor{blue}{L382 - Could or should?
@@
 
-\textcolor{red}{Gavin?}
+\textcolor{red}{The reviewer is correct that we do not include fossil taxa in our simulations.
+We used birth-death trees as a means of assessing the impact of different tree shapes rather than explicitly testing the impact of fossil taxa.
+The issue of whether fossil taxa should or could be included is not straightforward.
+If fossil taxa can be reliably placed in the phylogeny then we anticipate that they will typically improve model accuracy, i.e.
+they should be included.
+On the other hand, if placement of fossil taxa is unreliable then they also have the potential to mislead analyses.
+GT - CHECK SANSOM and WILLS PAPER FOR EXAMPLE OF BAD PLACEMENT.
+}
 
@@
 
-\textcolor{red}{Gavin?}
+\textcolor{red}{The key point here is not that the tree itself is special but rather how the OU model is fitted.
+For ultrametric trees it is possible to fit the OU model simply by adjusting the branch lengths in the tree directly.
+This is not possible for non-ultrametric trees.
+The problem is that related species can effectively become more similar to one another than to themselves - an inherently not tree-like pattern.
+This means that the OU model can only be computed using the variance-covariance matrix, and not with the tree itself.
+The important point from an end-user perspective is that some software is not suitable for fitting the OU model to non-ultrametric trees.}
 
@@
 % We have added details to this Figure caption.
-\textcolor{red}{Gavin?}
+\textcolor{red}{This is Rob's figure originally - tippy etc is a hangover from earlier version.}
 
@@
 
-\textcolor{red}{Gavin? We may need to discuss this as I'm not entirely sure what Graham means?}
+\textcolor{red}{We agree that a half-life ~4 times the height of the tree is no different from Brownian motion in any meaningful biological sense.
+This is exactly the point.
+It is not uncommon to see interpretations of models based only on model fit and not on the the model parameters hence our emphasis is on problems with model fitting.
+However, the point about parameter interpretation is equally important and we have added a new paragraph (lines XX-XX) to discuss this explicitly.
+
+GT: we could add a figure or a couple of examples that for a given simulation set compared half lifes or alphas for simulations that favoured OU vs favoured BM. - maybe this is Graham's point?}
 
@@
 There is evidence to suggest that rather than helping naive readers, too many equations actually impedes understanding and communication of concepts (see Fawcett and Higginson, 2012 PNAS; http://www.pnas.org/content/109/29/11735.short), thus we prefer to keep this as it is so as not to discourage people from reading further.
-\textcolor{red}{Not sure if you agree with me here, but I didn't really want to move the appendix back into the paper as I think it's confusing.}
+\textcolor{red}{Not sure if you agree with me here, but I didn't really want to move the appendix back into the paper as I think it's confusing.
+
+GT: I agree, although wondered whether we could do anything like a TREE-style box.
+Maybe we can make the point that this is an in-print appendix, like what they did in the olden days, rather than online supplementary material.}
 
 Illustrating the process however is a nice idea so we have added a figure to the paper that does this. 
-\textcolor{red}{Do you have code handy to make a figure for this? I'd have to write some from scratch and I'm meant to be on holiday...}
+\textcolor{red}{Do you have code handy to make a figure for this? I'd have to write some from scratch and I'm meant to be on holiday...
+
+GT: I don't.
+I think we could flag figure 4 here - it already does what Graham is suggesting with respect to covariances.
+Perhaps combining this with a diffusion process for alpha = 0, 1, 10 would be sufficient.
+It could be a three column figure with the covariances, diffusion process and tree transform.}
 
@@
 
-\textcolor{red}{Gavin?}
+\textcolor{red}{One for Chris and Andrew to comment on I think.}
 
@@
 % We have added details to this Figure caption.
-\textcolor{red}{Gavin?}
+\textcolor{red}{Fig 2 is Rob's.
+I assume that the black line is just alpha = 0 and the red line is the critical value of the chi-square distribution to highlight the cases where the Brownian model would be rejected.}
 

### 2015-07-21 Manuscript/Revision/response.tex
Edits in response to GHT's changes

--- 2015-07-06 Manuscript/Revision/response.tex
@@
 \usepackage[usenames,dvipsnames]{xcolor}
-\usepackage[colorlinks = true, urlcolor = blue]{hyperref} %hyperlinks
 \usepackage[osf]{mathpazo}
@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper, Gavin Thomas and colleagues}
+\signature{Natalie Cooper \\ Gavin Thomas \\ Chris Venditti \\ Andrew Meade \\ Rob Freckleton}
 \address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ nhcooper12@gmail.com}
@@
 This is generally fixed by not estimating the root state, and instead using mu as the root state or a root state that is distributed according to the stationary distribution of the OU process.
-My feeling is that this happens rarely (most implementations don't estimate the root state) and hopefully people would notice this and take any results with a large pinch of salt.
+My feeling is that this happens rarely (most implementations don't estimate the root state by default) and hopefully people would notice this and take any results with a large pinch of salt.
 We have added a brief statement to alert people to this fact.
@@
 
-\textcolor{red}{Gavin?}
+A Yule model was chosen because time constraints meant we couldn't test every single different kind of tree as we did for the earlier analyses.
+As noted above, all of the simulated trees were ultrametric, so this should not influence our results.
 
@@
 
-% We have clarified what we mean by bounded (alpha must be 0 or greater) and reworded this section to make it clearer and also in response to reviewer 2 below.
-\textcolor{red}{I think Rob wrote this originally - can you check I've reworded it correctly?}
+We have clarified what we mean by bounded (alpha must be 0 or greater) and reworded this section to make it clearer and also in response to reviewer 2 below.
 
@@
 
-\textcolor{red}{Again Rob wrote this originally. Any ideas?}
+\textcolor{red}{Rob?}
 
@@
 Thus it suggests that selection for that optimum is extremely weak at best.
-For example if t1/2 is 100 million years, then it suggests that a species will take 200 million years to reach the trait optimum.
+For example if t1/2 is 100 million years, then it suggests that a species will take 100 million years to get halfway to the trait optimum.
 This is hardly suggestive of a strong selective pressure towards that optimum.
@@
 
-\textcolor{red}{Gavin?}
+The reviewer is correct that we do not include fossil taxa in our simulations.
+We used birth-death trees as a means of assessing the impact of different tree shapes rather than explicitly testing the impact of fossil taxa.
+The issue of whether fossil taxa should or could be included is not straightforward.
+If fossil taxa can be reliably placed in the phylogeny then we anticipate that they will typically improve model accuracy, i.e.
+they should be included.
+On the other hand, if placement of fossil taxa is unreliable then they also have the potential to mislead analyses.
+% GT - CHECK SANSOM and WILLS PAPER FOR EXAMPLE OF BAD PLACEMENT.
 
@@
 
-\textcolor{red}{Gavin?}
+The key point here is not that the tree itself is special but rather how the OU model is fitted.
+For ultrametric trees it is possible to fit the OU model simply by adjusting the branch lengths in the tree directly.
+This is not possible for non-ultrametric trees.
+The problem is that related species can effectively become more similar to one another than to themselves - an inherently not tree-like pattern.
+This means that the OU model can only be computed using the variance-covariance matrix, and not with the tree itself.
+The important point from an end-user perspective is that some software is not suitable for fitting the OU model to non-ultrametric trees.
+We have clarified this in the text.
 
@@
 
-% We have added details to this Figure caption.
-\textcolor{red}{Gavin?}
+We have added details to this Figure caption.
+Tippy trees are those with distributed disproportionately late in the clade history (i.e.
+nearer to the present).
+Rooty trees are those with branching events distributed disproportionately early in the clades history (nearer to the root).
+In all cases the "true" value of alpha is 0 (black dashed line).
+The red dashed line is ...
+% NC: Need to add this!
+
+\textcolor{red}{Rob?}
 
@@
 
-\textcolor{red}{Gavin? We may need to discuss this as I'm not entirely sure what Graham means?}
+We agree that a half-life ~4 times the height of the tree is no different from Brownian motion in any meaningful biological sense.
+This is exactly the point.
+It is not uncommon to see interpretations of models based only on model fit and not on the the model parameters hence our emphasis is on problems with model fitting.
+However, the point about parameter interpretation is equally important and we have added a new paragraph (lines XX-XX) % NC: ADD LINE NUMBERS
+to discuss this explicitly.
+
+% GT: we could add a figure or a couple of examples that for a given simulation set compared half lifes or alphas for simulations that favoured OU vs favoured BM. - maybe this is Graham's point?
 
@@
 There is evidence to suggest that rather than helping naive readers, too many equations actually impedes understanding and communication of concepts (see Fawcett and Higginson, 2012 PNAS; http://www.pnas.org/content/109/29/11735.short), thus we prefer to keep this as it is so as not to discourage people from reading further.
-\textcolor{red}{Not sure if you agree with me here, but I didn't really want to move the appendix back into the paper as I think it's confusing.}
+However, note that this is an in-print appendix, rather than online appendix, so will appear associated with the paper.
 
-Illustrating the process however is a nice idea so we have added a figure to the paper that does this. 
-\textcolor{red}{Do you have code handy to make a figure for this? I'd have to write some from scratch and I'm meant to be on holiday...}
+% NC: For now not mentioning the figure suggestion at all and hoping that this is just going to Anjali rather than back to Graham!
 
@@
 
-\textcolor{red}{Gavin?}
+\textcolor{red}{Chris and Andrew?}
 

### 2015-07-27 Manuscript/Revision/response.tex
Edits in response to GHT's changes
Added CV and AM edits

@@
 \textcolor{red}{Rob?}
+% NC: Need to add something here still...
 
@@
 they should be included.
-On the other hand, if placement of fossil taxa is unreliable then they also have the potential to mislead analyses.
-% GT - CHECK SANSOM and WILLS PAPER FOR EXAMPLE OF BAD PLACEMENT.
+On the other hand, if placement of fossil taxa is unreliable then they also have the potential to mislead analyses (see Sansom \& Wills 2012 for examples).
 
@@
 
-We agree that a half-life ~4 times the height of the tree is no different from Brownian motion in any meaningful biological sense.
+We agree that a half-life approximately 4 times the height of the tree is no different from Brownian motion in any meaningful biological sense.
 This is exactly the point.
@@
 
-% GT: we could add a figure or a couple of examples that for a given simulation set compared half lifes or alphas for simulations that favoured OU vs favoured BM. - maybe this is Graham's point?
+Also yes, HPDs = quantiles. We have changed this throughout.
 
@@
 However, note that this is an in-print appendix, rather than online appendix, so will appear associated with the paper.
-
-% NC: For now not mentioning the figure suggestion at all and hoping that this is just going to Anjali rather than back to Graham!
 
@@
 
-\textcolor{blue}{Baele, G., Lemey, P., Bedford, T., Rambaut, A., Suchard, M.
-A., \& Alekseyenko, A.
-V.
-(2012).
-Improving the accuracy of demographic and molecular clock model comparison while accommodating phylogenetic uncertainty.
-Molecular biology and evolution, 29(9), 2157-2167.}
-
-\textcolor{red}{Chris and Andrew?}
+We agree that Stepping Stones is better than the Harmonic Mean to estimate the marginal likelihood.
+And we agree that stepping stone sampling should be the method of choice for Bayesian comparative analyses.
+We have added a brief mention that Baele et al 2012 show that using stepping stone sampling is preferable to using harmonic means.
 

### 2015-07-27 Manuscript/ohyou.tex
Edits due to reviewer comments

@@
 \author{
-  Natalie Cooper$^{1,2,*,\dag}$, Gavin H. Thomas$^{3,\dag}$, Chris Venditti$^{4}$\\ Andrew Meade$^{4}$ and Rob P. Freckleton$^{3}$\\
+  Natalie Cooper$^{1,2,3,*,\dag}$, Gavin H. Thomas$^{4,\dag}$, Chris Venditti$^{5}$\\ Andrew Meade$^{5}$ and Rob P. Freckleton$^{4}$\\
 }
@@
   $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
-  $^4$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
-  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
-       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^3$ Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\
+  $^4$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
+  $^5$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
+  $^*$ Corresponding author: nhcooper12@gmail.com, Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\ Tel: +44 20 7942 5000.\\
   $^\dag$These authors contributed equally.
@@
   Phylogenetic comparative methods are increasingly used to give new insights into the dynamics of trait evolution in deep time. 
-  The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
+  For continuous traits the core of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
   However, the properties of these models are often poorly understood which can lead to the misinterpretation of results.
   Here we focus on one of these models - the Ornstein Uhlenbeck (OU) model.
-  We show that the OU model is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, and that many studies fitting this model use datasets that are small and prone to Type I error.
-  We also show that very small amounts of error in datasets can have profound effects on the performance of OU models.
+  We show that the OU model is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, and that many studies fitting this model use datasets that are small and prone to this problem.
+  We also show that very small amounts of error in datasets can have profound effects on the inferences derived from OU models.
   Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model. 
@@
   The popularity of the OU model has grown extensively in recent years (Fig.
-\ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supporting Information).
+\ref{figure.literature}); even just between 2012 and 2014 over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published (Google Scholar search 15th March 2015; see Supporting Information).
   This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
@@
   As formulated by \citet{hansen1997stabilizing}, a trait has a primary optimum that is the mean of individual species optimum for that trait. 
-  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive}. 
-  However, this is not an estimate of stabilising selection in the population genetics sense.
+  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central trait value \citep[the primary optimum;][]{hansen2012adaptive}. 
+  However, this is not an estimate of stabilising selection in the population genetics sense, where it is a measure of selection within a population towards a fitness optimum on an adaptive landscape (Lande, 1976).
+This is a qualitatively different process to trait evolution among species which is more akin to a trait tracking movement of the adaptive optima itself.
 
@@
   OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{christin2013anatomical}(Table \ref{table.uses})
-  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supporting Information).  
+  OU models are also commonly used model phylogenetically structured residual error in evolutionary correlations (\citealp{revell2010phylogenetic}; often colloquially referred to as ``controlling for phylogeny''; Table \ref{table.uses}; Supporting Information).
   Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
@@
   ; instead it improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
-  It is not clear that it was originally intended that the OU model would be used in such a context.
-  Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.
+  It is not clear that it was originally intended that the OU model would merely be used in such a context.
+  Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface,uyeda2014novel}.
 
@@
   The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
-  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.
+  Note that we focus on OU models with a single stationary optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.
 For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.
@@
 
-   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$.
+   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value, $\mu$ (also called $\theta$) with an attraction strength proportional to the parameter $\alpha$.
     The model has the following form:
@@
     Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
+    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean, $\mu$, if they evolve away from it.$\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$.
     The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
-    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
-    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see the Appendix.
+    For more details see the Appendix.
 
-    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
+    In many implementations of the OU model \citep[e.g. GEIGER][]{Harmon:2008aa}, $\mu$ is the same as the state of the root at time zero, $X(0)$. 
+    This is referred to as a ``single stationary peak'' or SSP model (\textit{sensu} \citealp{Harmon:2010aa}).
+Some implementations also allow users to estimate $X(0)$ \citep[e.g.
+OUwie;][]{beaulieu2012modeling}.
+    Where $X(0) \neq \mu$ this is referred to as a single peak OU process. 
+    However, estimating $X(0)$ can sometimes lead to nonsensical values of $\mu$, i.e.
+far outside the range of values for the trait, thus results should be interpreted with caution.
+$X(0)$ can also be defined \textit{a priori}, but this is only appropriate when fossil data, or other independent evidence, allow confident estimates of the root state.
+
+    When $\alpha$ is close to zero, evolution is approximately Brownian, (but note that in the special case of a single peak OU model with $X(0) \neq \mu$, when $\alpha$ is zero, evolution approximates Brownian motion with a trend; \citealp{hansen1997stabilizing,benson2014competition}), then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent.
     Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
-    Note that $\alpha$ scales with tree height, so it needs to be interpreted relative to this. 
-    Generally the simplest solution is to rescale tree heights to 1 \citep[e.g.,][simulations below]{ives2010phylogenetic}. 
+    Note that $\alpha$ scales with tree height (i.e.
+the maximum distance from the root of the tree to the tips); taller trees will have lower $\alpha$ values, all else being equal, because there is more time for traits to return to the optimum value thus the strength of the pull towards the optimum, $\alpha$, can be smaller.
+    Thus $\alpha$ values need to be interpreted relative to tree height.Generally the simplest solution is to rescale tree heights to 1 \citep[e.g.,][simulations below]{ives2010phylogenetic}. 
     \citet{ives2010phylogenetic} suggest interpreting $log( \alpha)$ (after rescaling trees to a height of 1) rather than raw $\alpha$.
@@
   \subsection{Performance of $\alpha$ and Likelihood ratio tests}
-    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT (\citealp{Thomas:2011aa}; \href{https://github.com/ghthomas/motmot}{github.com/ghthomas/motmot}).
+    To determine whether $\alpha$ is biased under conditions where it should be zero, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT (\citealp{Thomas:2011aa}; \href{https://github.com/ghthomas/motmot}{github.com/ghthomas/motmot}).
     This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model.
     We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
-    This gives us a measure of Type I error rate for the OU model.
+    We refer to this measure as the Type I error rate for the OU model.
 
@@
 Tree type refers to the extinction fraction for the birth-death trees.
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
         \bigskip
@@
         \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on trees simulated under time variable speciation rates.
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
         \bigskip
@@
       \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
-      \caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ``true'' value of $\alpha$ is 0.
+      \caption{Examples of profile likelihoods for selected simulated datasets for tippy (A), rooty (B) and Yule (C) simulated trees. 
+      Each solid black line represents one simulated dataset selected at random. 
+      Tippy trees are those with branching events distributed disproportionately late in the clade's history (i.e. nearer to the present).
+      Rooty trees are those with branching events distributed disproportionately early in the clade's history (i.e.
+nearer to the root).
+In all cases the ``true'', value of $\alpha$ is 0 (black dashed line).
+The red dashed line is...
+% FIX!!!
       }
@@
       \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure3.pdf}
-      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
+      \caption{The number of taxa in phylogenies used to fit Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
 Two studies with \textgreater 3000 taxa have been omitted for clarity.
@@
     We provide recommendations related to these findings below.
+
+    For small trees the confidence limits on parameter estimates are broad (Figure \ref{figure:ouprofiles}). 
+    However, the median is typically low irrespective of tree size (see Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}) and even when the OU model is favoured, scrutiny of the model parameters suggest that the favoured OU model is biologically indistinguishable from Brownian motion.
+    Indeed, examination of the effect of $\alpha$ on the expected covariances among taxa (Figure \ref{figure:ouhistory}) confirms that covariances are barely distinguishable when $\alpha$ \textless 1 for unit height trees.
+    This suggests that, notwithstanding the elevated Type I errors of OU models, their interpretation requires examination of parameters. 
+    We return to the issue of parameter interpretation below, specifically with reference to the increasingly commonly used ``phylogenetic half-life'', $t_\frac{1}{2}$.
 
@@
         \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.).
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
+The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
         \bigskip
@@
   \item \textbf{Likelihood ratio tests are untrustworthy.}
-  All Likelihood ratio tests depend on asymptotic properties, so there is no guarantee that any parameter such as $\alpha$ will behave appropriately when tested. 
-  In the case of $\alpha$ we would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
-  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as simulation and MCMC, should be considered.
+  All Likelihood ratio tests assume that the Likelihood ratio statistic is asymptotically $\chi^2$ distributed. 
+  This is unlikely to be the case here because $\alpha$ is bounded (i.e. it cannot be any smaller than zero) and has a non-linear effect on the expected variances. 
+  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as simulation and MCMC, should be considered (see Recommendations and potential solutions (i) and (ii) below).
   \item \textbf{Measurement error increases Type I error rates.} 
@@
       Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree \citep{FitzJohn:2012aa} and GEIGER \citep[the Single Stationary Peak model in \texttt{fitContinuousMCMC};][]{Harmon:2008aa} and stand alone software including BayesTraits \citep{pagel2013bayestraits}.
+      Below we explore Bayesian methods as a possible remedy to the limitations of fitting OU models in a likelihood framework.
 
-	    We (AM and CV) repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits \citep{pagel2013bayestraits}. 
+	    We repeated our simulations in a Bayesian framework implemented in BayesTraits \citep{pagel2013bayestraits}. 
       We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}. 
@@
       Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). 
+      We use stepping stone sampling as it has been shown to estimate the marginal Likelihood better than the harmonic mean \citep{baele2012improving}. 
       We treated Bayes factors \textgreater 2 as evidence favouring the OU model \citep{kass1995bayes}. 
@@
 
-      Table \ref{tables:bayesian} shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
-      The Bayesian approach results in highly conservative rejection rates regardless for all tree shapes in the absence of measurement error. 
+      Table \ref{tables:bayesian} shows the results from the exponential prior with mean = 10. 
+      This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
+      The Bayesian approach results in highly conservative rejection rates regardless of tree shape in the absence of measurement error. 
       While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power, although testing would be required to confirm this.
@@
       However, more error or larger trees result in the frequent rejection of the Brownian model. 
-      As noted above, this is not an issue of Type I and it is entirely correct that the Brownian model is rejected. 
+      As noted above, this is not an issue of Type I error and it is entirely correct that the Brownian model is rejected. 
       In such cases emphasis should shift to how the $\alpha$ parameter is interpreted. 
@@
 The value of $\alpha$ is the median across simulated data sets based on modal estimates from the posterior distribution.
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model based on Bayes factors \textgreater 2.}
+The rejection rate is the proportion of Ornstein Uhlenbeck  models favoured relative to a Brownian motion model based on Bayes factors \textgreater 2.}
         \bigskip
@@
 The value of $\alpha$ is the median across simulated data sets based on modal estimates from the posterior distribution.
-The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.} %NC: Might need to modify legend
+The rejection rate is the proportion of Ornstein Uhlenbeck  models favoured relative to a Brownian motion model.} %NC: Might need to modify legend
         \bigskip
@@
     \item \textbf{Consider plausible alternative hypotheses.}
-    Because the OU model was proposed with evolutionary process in mind, alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ are often overlooked.
+    Because the OU model was proposed with evolutionary processes in mind, alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ are often overlooked.
     The effects of measurement error in particular suggest that $\alpha$ must always be interpreted with caution. 
@@
     Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very small and biologically indistinguishable from Brownian (e.g., examples in \citealp{Harmon:2010aa}).
-    In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
+    In these circumstances, it is likely that measurement error, intraspecific variation, or phylogenetic uncertainty are generating noise that is more effectively modeled by the extra parameters in the OU model than by the Brownian model alone.
     Thus this does not reflect any kind of OU process underlying the data. 
@@
 	  \item \textbf{Calculate the phylogenetic half-life.}
-    The $\alpha$ parameter ranges from zero to infinity, thus recognising ``small'' or ``large'' values may not be intuitive. 
+    The $\alpha$ parameter ranges from zero to infinity (although in practice an upper bound is often set, for example in GEIGER this is $\alpha = 150$; \citealp{Harmon:2008aa}), , thus recognising ``small'' or ``large'' values may not be intuitive.
     $\alpha$ can sometimes be interpreted more easily by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:
@@
     $t_\frac{1}{2}$ equal to the height of the phylogeny is a moderate value \citep{hansen2012interpreting}.
-    We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
-    However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process. 
+    We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum'' \citep{hansen1997stabilizing}.
+    However, if $t_\frac{1}{2}$ is extremely large relative to tree height, it suggests that if an OU process is acting, it is extremely weak (e.g.
+for a clade that is 50 million years old, a $t_\frac{1}{2}$ of 100 million years suggests a species will not approach the optimum within the temporal range of the clade) thus should not be interpreted as evidence of any kind of process.
     As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would lead to similar biases in $t_\frac{1}{2}$.
@@
     \item \textbf{If possible, include ancillary data.}
-    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}, provided their placement is well known.
-    This is implemented in GEIGER \citep{Harmon:2008aa}.
+    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.
+    Indeed, if fossil taxa can be reliably placed in the phylogeny then they may improve model accuracy and a strong case can be made that fossils should be included. 
+    On the other hand, if there is substantial uncertainty in fossil placement then they should be treated cautiously.
+Placement of fossil taxa is implemented in GEIGER \citep{Harmon:2008aa} and BayesTraits \citep{pagel2013bayestraits}.
     A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances depend on both the shared distances between species and the distance of a node to the nearest tip (see equation \ref{equation:OUcov} in the Appendix).
-    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like \citep{slater2014correction}.
+    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like, for example, related species can effectively become more similar to one another than to themselves - an inherently non tree-like pattern \citep{slater2014correction}.
     Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
@@
   Indeed, the problems that we report are not limited to OU models. 
-  Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being misled by measurement error. 
+  Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being misled by measurement error, and in some scenarios measurement error can also incorrectly favour Brownian motion over the true model, e.g.
+if the true model is Early Burst.
   The fundamental problem is that rejection of the Brownian model in favour of another model does not necessarily say anything about process. 
-  This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}.
-  We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.
+  This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}, however, this does not appear to be possible for comparisons of the single stationary peak OU model with noisy Brownian processes.
+  We should therefore not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.
 
-\section{Epilogue: Open and (vaguely) reproducible science}
+\section{Epilogue: The challenges of open and reproducible science}
   At the symposium that generated this special issue, one of us (NC) gave a talk on Open Science and reproducibility. 
@@
   All simulated phylogenies, data, and R code are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
-However, although our R code is provided, it is messy and difficult to use.
+However, although our R code is provided, it is disorganised and thus difficult to use.
   We also do not provide an automated way of reproducing the data collection for our literature review. 
@@
   BayesTraits is free to download as a binary executable for various platforms but it is not Open Source. 
-  This means that while BayesTraits analyses are reproducible the software has limitations with respect to long-term development of the code. 
-  We included this epilogue to highlight that fully reproducible and Open Science is really hard, but we are trying to improve! 
+  This means that while BayesTraits analyses are reproducible, the software has limitations with respect to long-term development of the code. 
+  We included this epilogue to highlight that fully reproducible and Open Science is difficult, but we are trying to improve.
   With a little effort most people should be able to produce something vaguely reproducible, and to provide their data and code, moving us all slightly closer to truly reproducible and Open Science.
@@
 \section{Acknowledgments}
-  Thanks to Rich FitzJohn, Mark Pagel, Matt Pennell and Graham Slater for fruitful discussions about OU models, and three anonymous reviewers for helpful comments on a previous version of this paper. 
+  Thanks to Rich FitzJohn, Mark Pagel and Matt Pennell for fruitful discussions about OU models, and four anonymous reviewers and Graham Slater for helpful comments on previous versions of this paper.
   NC was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696. 
-  GHT was supported by a Royal Society University Research Fellowship. 
+  GHT was supported by a Royal Society University Research Fellowship, grant number: UF120016. 
   CV was supported by a Leverhulme Trust Research Project Grant RPG-2013-185. 

### 2015-07-27 Manuscript/ohyou.tex
Smalle edits to text

@@
   Note that we focus on OU models with a single stationary optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.
-For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.
+For discussions on the performance of multiple optima OU models we refer the reader to \citet{beaulieu2012ouwie}.
 
@@
     The model assumes the correlation structure among trait values is proportional to the extent of shared ancestry for pairs of species. 
-    This means that close relatives will be more similar in their trait values than more distant relatives. It also means that variance in the trait will increase (linearly) in proportion to time. 
+    This means that close relatives will be more similar in their trait values than more distant relatives. 
+    It also means that variance in the trait will increase (linearly) in proportion to time. 
     The model has two parameters, the Brownian rate parameter, $\sigma^2$ and the state of the root at time zero, $X(0)$.
@@
     Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
-    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean, $\mu$, if they evolve away from it.$\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$.
+    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean, $\mu$, if they evolve away from it.
+     $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$.
     The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
@@
     \citet{ives2010phylogenetic} suggest interpreting $log( \alpha)$ (after rescaling trees to a height of 1) rather than raw $\alpha$.
-They equate $-log( \alpha) = -4$ as a very low, almost Brownian value, and $-log( \alpha) = 4$ as a very high value.
+They equate $-log( \alpha) = 4$ as a very low, almost Brownian value, and $-log( \alpha) = -4$ as a very high value.
     Others \citep[e.g.][]{slater2015iterative,slater2013robust,hansen2012interpreting} prefer to use the ``phylogenetic half-life'': $t_\frac{1}{2}$ (see ``Recommendations for interpreting $\alpha$'' below).
@@
 \section{Epilogue: The challenges of open and reproducible science}
-  At the symposium that generated this special issue, one of us (NC) gave a talk on Open Science and reproducibility. 
+  At the symposium that generated the special issue this paper is part of (the Linnaean Society's ``Radiations and Extinctions in Deep Time''symposium in November 2014), one of us (NC) gave a talk on Open Science and reproducibility.
   We have therefore tried to make this paper as open and reproducible as possible. 
@@
 % Appendix
+\newpage
 \section{Appendix}
@@
 
-\newpage
-\section{Figure legends}
-
-\noindent
-\textbf{Figure 1}: The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year.
-See Supporting Information for details.\\
-
-\noindent
-\textbf{Figure 2}: Examples of profile likelihoods for selected simulated datasets. In all cases the ``true'' value of $\alpha$ is 0.
-
-\noindent
-\textbf{Figure 3}: The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014.
-Two studies with \textgreater 3000 taxa have been omitted for clarity.
-See Supporting Information for details.
-
-\noindent
-\textbf{Figure 4}: Scaling of expected trait similarity with time since evolutionary divergence predicted by the Ornstein Uhlenbeck model.
-The covariance between species' traits values is scaled by the intra-specific trait variance (i.e.
-equal to correlation between species' traits).
-This is plotted against the relative time of shared history (time at which species branched from each other, divided by the total tree height: $\frac{t_{ij}}{T}$).
-
-\newpage
-\section{Tables}
-
 \end{document}

### 2015-08-03 Manuscript/ohyou.tex
Update email address

@@
   $^5$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
-  $^*$ Corresponding author: nhcooper12@gmail.com, Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\ Tel: +44 20 7942 5000.\\
+  $^*$ Corresponding author: natalie.cooper@nhm.ac.uk, Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\ Tel: +44 20 7942 5000.\\
   $^\dag$These authors contributed equally.

### 2015-08-13 Manuscript/Revision/response.tex
Added RFs comments to response letter

@@
 
-\textcolor{red}{Rob?}
-% NC: Need to add something here still...
+AIC is a little tricky - in one sense there is no problem with calculating AIC itself.
+Comparisons of AICs however are probably dodgy - e.g.
+if you compare AIC of BM and OU model, the test is basically much the same as LRT anyway, so we are not sure that would be a sensible approach.
+For more complex models the AIC approach has problems that are unrelated to AIC, e.g.
+how do you calculate effective numbers of parameters?
+Unfortunately these problems are hard to pin down, and decisions are more related to personal opinions than clearly citable examples.
+We would tend to avoid using AIC for comparative questions, but not everyone agrees.
+So the short answer is yes, but we'd prefer not to open that particular can of worms here!
+We have however added a brief comment to the manuscript to mention this potential problem.
+
+% Note that these issues will also apply when using the Akaike Information Criteria (AIC) to compare models, as this is essentially the same as performing a Likelihood ratio test.
+AIC also presents other difficulties when used for complex models such as those use in phylogenetic comparative analyses, so we tend to avoid them it these reasons.
 
@@
 In all cases the "true" value of alpha is 0 (black dashed line).
-The red dashed line is ...
-% NC: Need to add this!
-
-\textcolor{red}{Rob?}
+The red dashed line is -1.92 log-likelihood units from the maximum: using log-likelihood ratio tests, values of $\alpha$ yielding values higher than this would be considered statistically indistinguishable from the maximum likelihood value.
 
@@
 In all cases the "true" value of alpha is 0 (black dashed line).
-The red dashed line is ...
-% NC: Need to add this!
-
-\textcolor{red}{Rob?}
+The red dashed line is -1.92 log-likelihood units from the maximum: using log-likelihood ratio tests, values of α yielding values higher than this would be considered statistically indistinguishable from the maximum likelihood value.
 

### 2015-08-13 Analyses/plotFig2.R
Code to recreate Figure 2

@@ -1,0 +1,53 @@
+rm( list = ls() )
+
+layout( matrix( c(1:3 ), ncol = 1))
+
+# rob.par™ - my trademark plotting parameters
+rob.par <- par(bty = "l",
+		cex.lab = 1.1,
+		tcl = 0.3,
+		las = 1,
+		pch = 1,
+		bty = "l")
+	
+# ---- Plot Tippy Tree -----
+alpha <- read.csv( "Data/ProfileDataTippyOU50.csv" )
+alpha <- alpha[1:500,]
+
+plot( alpha[,1], alpha[,2] - max( alpha[,2] ), xlim = c(0,5), ylim = c(-20, 0), type = "l", xlab = "", ylab = "", main = "Tippy" )
+
+for( i in 2:11) {
+	lines( alpha[,1], alpha[,i]- max( alpha[,i] ) )
+	}
+
+
+lines( c(0, 5), c(-1.92, -1.92), col = "red", lty = "22" ) 
+
+
+
+# ---- Plot Rooty Tree ----
+
+alpha <- read.csv( "Data/ProfileDataRootyOU50.csv" )
+alpha <- alpha[1:500,]
+
+plot( alpha[,1], alpha[,2] - max( alpha[,2] ), xlim = c(0,5), ylim = c(-20, 0), type = "l", xlab = "", ylab = "log Likelihood" , main = "Rooty" )
+
+for( i in 2:11) {
+	lines( alpha[,1], alpha[,i]- max( alpha[,i] ) )
+	}
+	
+lines( c(0, 5), c(-1.92, -1.92), col = "red", lty = "22" ) 
+
+# ---- Plot Yule Tree ----
+alpha <- read.csv( "Data/ProfileDataYuleOU50.csv" )
+alpha <- alpha[1:500,]
+
+plot( alpha[,1], alpha[,2] - max( alpha[,2] ), xlim = c(0,5), ylim = c(-20, 0), type = "l", xlab = expression(alpha), ylab = "" , main = "Yule")
+
+for( i in 2:11) {
+	lines( alpha[,1], alpha[,i]- max( alpha[,i] ) )
+	}
+
+lines( c(0, 5), c(-1.92, -1.92), col = "red", lty = "22" ) 
+
+

### 2015-08-13 Analyses/OhYouFigures.R
Specify figures recreated with OhYouFigures.R

@@ -1,4 +1,4 @@
-# Making figures 
+# Making figures 1 and 3
 
 # Read in data
 ds <- read.delim("Data/papers.txt")

### 2015-08-13 Manuscript/ohyou.tex
Edits from RPF

@@
   These findings will be applicable to other models of evolution, but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter.
-  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}. 
+  We are not the first to describe some of these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}. 
   However widespread use of the model is clear evidence that many are unaware of the potential problems.
@@
   For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.
+
   It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilising selection'', this is inaccurate and misleading. 
@@
   OU models are also commonly used model phylogenetically structured residual error in evolutionary correlations (\citealp{revell2010phylogenetic}; often colloquially referred to as ``controlling for phylogeny''; Table \ref{table.uses}; Supporting Information).
-  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
+  As an extension beyond modeling single traits, phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model.
   This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data.
@@
   ; instead it improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
-  It is not clear that it was originally intended that the OU model would merely be used in such a context.
+  However it is not clear that it was originally intended that the OU model would merely be used in such a context.
   Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface,uyeda2014novel}.
@@
       \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
-      \caption{Examples of profile likelihoods for selected simulated datasets for tippy (A), rooty (B) and Yule (C) simulated trees. 
+      \caption{Examples of profile likelihoods for selected simulated datasets for tippy (A), rooty (B) and Yule (C) simulated trees of 50 taxa. 
       Each solid black line represents one simulated dataset selected at random. 
@@
 In all cases the ``true'', value of $\alpha$ is 0 (black dashed line).
-The red dashed line is...
-% FIX!!!
+The red dashed line represents -1.92 log-likelihood units from the maximum: using log-likelihood ratio tests, values of α yielding values higher than this would be considered statistically indistinguishable from the maximum likelihood value.
       }
@@
   \item \textbf{Likelihood ratio tests are untrustworthy.}
-  All Likelihood ratio tests assume that the Likelihood ratio statistic is asymptotically $\chi^2$ distributed. 
-  This is unlikely to be the case here because $\alpha$ is bounded (i.e. it cannot be any smaller than zero) and has a non-linear effect on the expected variances. 
+  All Likelihood ratio tests assume that the Likelihood ratio statistic is asymptotically (i.e. as sample sizes become large) $\chi^2$ distributed. 
+  In the OU model $\alpha$ is bounded (i.e.
+it cannot be any smaller than zero) and has a non-linear effect on the expected variances, so the assumptions of the Likelihood ratio test are not likely to be upheld for small samples.
   Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as simulation and MCMC, should be considered (see Recommendations and potential solutions (i) and (ii) below).
+Note that these issues will also apply when using the Akaike Information Criteria (AIC) to compare models, as this is essentially the same as performing a Likelihood ratio test.
+  AIC also presents other difficulties when used for complex models such as those use in phylogenetic comparative analyses, so we recommend avoiding them it these reasons.
   \item \textbf{Measurement error increases Type I error rates.} 
@@
       Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree \citep{FitzJohn:2012aa} and GEIGER \citep[the Single Stationary Peak model in \texttt{fitContinuousMCMC};][]{Harmon:2008aa} and stand alone software including BayesTraits \citep{pagel2013bayestraits}.
-      Below we explore Bayesian methods as a possible remedy to the limitations of fitting OU models in a likelihood framework.
 
+      We explored Bayesian methods as a possible remedy to the limitations of fitting OU models in a likelihood framework.
       We repeated our simulations in a Bayesian framework implemented in BayesTraits \citep{pagel2013bayestraits}. 
-      We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}. 
+      We determined the fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}. 
       The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
@@
       Table \ref{tables:bayesian} shows the results from the exponential prior with mean = 10. 
-      This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
+      This is a broad, liberal prior, but our results are similar regardless of priors. 
       The Bayesian approach results in highly conservative rejection rates regardless of tree shape in the absence of measurement error. 
@@
 The value of $\alpha$ is the median across simulated data sets based on modal estimates from the posterior distribution.
-The rejection rate is the proportion of Ornstein Uhlenbeck  models favoured relative to a Brownian motion model.} %NC: Might need to modify legend
+The rejection rate is the proportion of Ornstein Uhlenbeck  models favoured relative to a Brownian motion model.}
         \bigskip

### 2015-08-13 Manuscript/Revision/response.tex
Final edits to response letter

@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper \\ Gavin Thomas \\ Chris Venditti \\ Andrew Meade \\ Rob Freckleton}
-\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ nhcooper12@gmail.com}
+\signature{Natalie Cooper, Gavin Thomas \\ Chris Venditti, Andrew Meade, \\ and Rob Freckleton}
+\address{}
 \longindentation=0pt
@@
 We have however added a brief comment to the manuscript to mention this potential problem.
-
-% Note that these issues will also apply when using the Akaike Information Criteria (AIC) to compare models, as this is essentially the same as performing a Likelihood ratio test.
-AIC also presents other difficulties when used for complex models such as those use in phylogenetic comparative analyses, so we tend to avoid them it these reasons.
 
@@
 It is not uncommon to see interpretations of models based only on model fit and not on the the model parameters hence our emphasis is on problems with model fitting.
-However, the point about parameter interpretation is equally important and we have added a new paragraph (lines XX-XX) % NC: ADD LINE NUMBERS
+However, the point about parameter interpretation is equally important and we have added a new paragraph (lines 273-282).
 to discuss this explicitly.
@@
 
-We have added details to this Figure caption.
-Tippy trees are those with distributed disproportionately late in the clade history (i.e.
-nearer to the present).
-Rooty trees are those with branching events distributed disproportionately early in the clades history (nearer to the root).
-In all cases the "true" value of alpha is 0 (black dashed line).
-The red dashed line is -1.92 log-likelihood units from the maximum: using log-likelihood ratio tests, values of α yielding values higher than this would be considered statistically indistinguishable from the maximum likelihood value.
+We have added details to this Figure caption. See our response to reviewer 1 above.
 
-Let us know if you require any further information,
+We hope we have responded to all comments appropriately,
 
